### 2015-07-08 manuscript/defs.tex
Updating manuscript

\documentclass{article}
\usepackage{longtable}
\pagestyle{myheadings}
\usepackage{color}

\setlength{\parindent}{0pt}
\setlength{\parskip}{1ex plus 0.5ex minus 0.2ex}

\newtheorem{definition}{Definition}
\newtheorem{observation}{Observation}
\newtheorem{lemma}{Lemma}

\begin{document}

\section{Basic definitions}

My first attempt at providing a coherent mathematical definition of a
repeat element family in the context of a spaced seed.

\begin{definition}
  A {\it repeat element family descriptor (refd)} is a
  string over the alphabet $\{A,C,G,T,*\}$, describing the contents
  of any string in the family.
\end{definition}

\begin{definition}
  Given refds $r$ and $r'$, we will say $r'$ is a {\it slack
    substring} of $r$ (denoted $r' \prec r$) if there is a
  substring $r''$ of $r$ such that (1) $|r'|=|r''|$, and (2) for all
  $0 \leq i < |r'|$, either $r'_i = r''_i$, $r'_i = *$, or $r''_i=*$.
\end{definition}
(In otherwords, its a substring, with a potentially different *
pattern.)
{\bf Note: } I'm not sure if we should be allowing $r''_i=*$.

\begin{definition}
  We say an refd $r$ {\it matches} a Genome $G$ at position $i$ if, for
  all $j$ such that $r_j \neq *$, $r_j = G_{i+j}$.
\end{definition}

Example: If $r = AA*TT$, and $G = AACTTGGAAGTT$, then $r$ matches $G$
at positions $i=0$ and $i=7$.  If $G=AAATTT$, then $r$ matches $G$ at
positions $i=0$ and $i=1$.

\begin{definition}
  A spaced seed $s$ {\it matches} an refd $r$ at position $i$
  if, for every $0 \leq j < |s|$, $s_j = 0$
  whenever $r_{i+j}=*$.
\end{definition}

Example: If $s=11011$ and $r = AAAAA*GGG$, then $s$ matches $r$
at positions $i=0$ and $i=3$, but not at any other $i$.

\begin{definition}
  Given a seed $s$ and refd $r$, let $M_s(r)$ be the set of values $i$
  such that $s$ matches $r$ at $i$.
\end{definition}

Let $s=11011$ and $r = AA*CC*GGGGGG$.  Then $M_s(r) = [0, 3, 6, 7]$.

\begin{definition}
  A spaced seed $s$ is consistent with an refd $r$ if
  for every $i$, $0 \leq i < |r|$, there is some
  $i - |s| \leq j \leq i$ such that $s$ matches $r$ at position $j$.
\end{definition}

In other words: for any position $i$ of the refd, we must be able to
match the seed to a position of $r$ such that it then covers position
$i$.

Example: The seed $11011$ is compatible with $r=AA*AA*AA$.  (The seed
matches at positions 0 and 3, and all positions are covered by these
two.)  But it is not consistent with $AAAAA**AAAAA$, and there is no
seed that can match this string at any position that can cover $i=5$
or $i=6$.

\begin{observation}
  Let $L$ be the sortest sequence of the values in $M_s(r)$.  Then $s$
  is consistent with $r$ if and only if $\max_{0 < j < |s|-1} L[j+1]-L[j] \leq |s|$.
\end{observation}
That is, in the sorted list, every pair of adjacent elements must be
within $|s|$ of each other.

\begin{definition}
  Given a fixed genome $G$, and fixed spaced seed
  $s$, and a fixed value $f$, we define a elementary repeat family as
  a set $S$ of genome coordinates, $|S| \geq f$, such that there
  exists an refd $r$ where:
  \begin{itemize}
  \item $r$ matches the sequence of length $|r|$ starting at each
    element of $S$.  ($S$ is the set of all instances.)
  \item $s$ is consistent with $r$.  ($r$ corresponds to the seed.)
  \item There does not exist an refd $r'$, $r' \prec r$,  such that $r'$
    is consistent with $s$ and $M_s(r') - M_s(r) \neq \emptyset$.  (You
    cannot have a proper substring of $r$ that describes sequences
    outside of the instances described by $r$ -- minimality.)
  \item There does not exist an refd $r'$, $r \prec r'$, such that $s$ is consistent
    with $r'$, such that the instances defined by $M_s(r')$ contain all the
    instances of $M_s(r)$.  (Maximality.)
  \end{itemize}
\end{definition}
 {\bf Comment:} I'm not sure if the $r \prec r'$ is the right
 relationship.  Maybe just straight substring?  Or perhaps the
 definition of $\prec$ isn't quite right?

\section{Random Lemmas}

The following are things I wanted to try to prove that may or may not
be useful.

\begin{lemma}
  Let $s$ be a seed, $r$ be a refd consistent with $s$,
  and $r'$ be an refd consistent with $s$ such that $|r'| < |r|$.  
\end{lemma}

\end{document}

### 2015-07-08 chromosome_simulator3.py

@@ -107,7 +107,8 @@ def generate_chromosome(seq, markov_list, coord_adjust, rpt_gen, mask = False, m
 
     tail_length = min(max_interval, length-current_coord)
     if tail_length > 0:
-        s.append(markov_gen.generate_sequence(markov_list, tail_length))
+        s.append("".join([random.choice("ACGT") for i in range(tail_length)]))
+    
 
     sim_seq = "".join(s)
     sim_seq_len = len(sim_seq)

### 2015-07-08 RAIDER_src_v2_options_new/src/LmerVector.h
New version

@@ -1,0 +1,84 @@
+#ifndef LMER_VECTOR
+#define LMER_VECTOR
+#include <vector>
+#include <stdexcept>
+#include <iostream>
+
+using namespace std;
+
+#ifndef uint
+typedef unsigned int uint;
+#endif
+
+class Family;
+
+class LmerVector {
+public:
+
+	LmerVector(uint position) {
+		family = nullptr;
+		push_back(position);
+	}
+    
+    ~LmerVector() {
+    }
+
+	void push_back(uint val) {
+        if(lmers.size() > 0){
+            previous = back();
+        }
+		lmers.push_back(val);
+	}
+
+	uint operator [](uint i) const {
+		return lmers[i];
+	}
+
+	uint & operator [](uint i) {
+		return lmers[i];
+	}
+
+    uint prev() const {
+        return previous;
+    }
+
+	uint back() const {
+		return lmers.back();
+	}
+
+	uint front() const {
+		return lmers.front();
+	}
+
+    uint getOff() const {
+        return off;
+    }
+    
+    void setOff(uint i) {
+        off = i;
+    }
+
+	uint size() const {
+		return lmers.size();
+	}
+
+	Family* getFamily() const {
+		return family;
+	}
+
+	void setFamily(Family *fam) {
+		family = fam;
+	}
+
+    friend ostream &operator<<( ostream &output, const LmerVector &L) { 
+        output << "(\tFront:\t" << L.front() << "\tBack:\t" << L.back() << "\tSize:\t" << L.size() << "\tOffset In Family\t" << L.getOff() << "\tFamily:\t" << L.getFamily() << "\t)";
+        return output;            
+    }
+
+	vector<uint> lmers;
+	Family* family;
+    uint previous;
+    uint off;
+};
+
+#endif //LMER_VECTOR

### 2015-07-08 manuscript/defs.tex

@@
 
-\section{Basic definitions}
-
-My first attempt at providing a coherent mathematical definition of a
-repeat element family in the context of a spaced seed.
-
 \begin{definition}
@@
 \begin{definition}
-  Given refds $r$ and $r'$, we will say $r'$ is a {\it slack
-    substring} of $r$ (denoted $r' \prec r$) if there is a
-  substring $r''$ of $r$ such that (1) $|r'|=|r''|$, and (2) for all
-  $0 \leq i < |r'|$, either $r'_i = r''_i$, $r'_i = *$, or $r''_i=*$.
-\end{definition}
-(In otherwords, its a substring, with a potentially different *
-pattern.)
-{\bf Note: } I'm not sure if we should be allowing $r''_i=*$.
-
-\begin{definition}
   We say an refd $r$ {\it matches} a Genome $G$ at position $i$ if, for
@@
 
-Example: If $r = AA*TT$, and $G = AACTTGGAAGTT$, then $r$ matches $G$
+Example If $r = AA*TT$, and $G = AACTTGGAAGTT$, then $r$ matches $G$
 at positions $i=0$ and $i=7$.  If $G=AAATTT$, then $r$ matches $G$ at
@@
   A spaced seed $s$ {\it matches} an refd $r$ at position $i$
-  if, for every $0 \leq j < |s|$, $s_j = 0$
-  whenever $r_{i+j}=*$.
+  if, for every $0 \leq j < |s|$, if $r_{i+j} = *$ then $s_j = 0$.
 \end{definition}
@@
 
-In other words: for any position $i$ of the refd, we must be able to
+In otherwords: for any position $i$ of the refd, $i$ must be able to
 match the seed to a position of $r$ such that it then covers position
@@
   Let $L$ be the sortest sequence of the values in $M_s(r)$.  Then $s$
-  is consistent with $r$ if and only if $\max_{0 < j < |s|-1} L[j+1]-L[j] \leq |s|$.
+  is consistent with $r$ if and only if $\max_{0 < j < |s|-1}
+  L[j+1]-L[j] \leq |s|$.  In the sorted list, every pair of adjacent
+  elements must be within $|s|$ of eachother.
 \end{observation}
-That is, in the sorted list, every pair of adjacent elements must be
-within $|s|$ of each other.
 
@@
   \item $s$ is consistent with $r$.  ($r$ corresponds to the seed.)
-  \item There does not exist an refd $r'$, $r' \prec r$,  such that $r'$
-    is consistent with $s$ and $M_s(r') - M_s(r) \neq \emptyset$.  (You
-    cannot have a proper substring of $r$ that describes sequences
-    outside of the instances described by $r$ -- minimality.)
-  \item There does not exist an refd $r'$, $r \prec r'$, such that $s$ is consistent
-    with $r'$, such that the instances defined by $M_s(r')$ contain all the
+  \item There does not exist a proper substring $r'$ of $r$ such that $r$
+    is consistent with $s$ and $M_s(r') - M_s(r) = \emptyset$.  You
+    can't have a proper substring of $r$ that describes sequences
+    outside of the instances described by $r$.  (Minimality.)
+  \item There does not exist a substring $r'$ propertly containing
+      $r$ such that the instances defined by $M_s(r')$ contain all the
     instances of $M_s(r)$.  (Maximality.)
@@
 \end{definition}
- {\bf Comment:} I'm not sure if the $r \prec r'$ is the right
- relationship.  Maybe just straight substring?  Or perhaps the
- definition of $\prec$ isn't quite right?
-
-\section{Random Lemmas}
-
-The following are things I wanted to try to prove that may or may not
-be useful.
-
-\begin{lemma}
-  Let $s$ be a seed, $r$ be a refd consistent with $s$,
-  and $r'$ be an refd consistent with $s$ such that $|r'| < |r|$.  
-\end{lemma}
 

### 2015-07-08 manuscript/defs.tex
Updating defs

--- 2015-07-08 manuscript/defs.tex
@@
 
+\begin{definition}
+Given the refd $r$ of an elementary repeat family with set $S$, we say
+that $r$ is {\it tight} if, for each $i$ such that $r_i = *$, there
+exists two sequences defined by $S$ that have different bases in
+position $i$.
+\end{definition}
+In other words: $r$ is tight if it only uses $*$ symbols where it must
+to match everything sequence defined by $S$.
+
+{\bf Comments:} I had it in mind that the algorithm would always
+return a tight refd.  But in Carly's defense she gave the example:
+s=11011, AAAAATCCCCC, AAGAATCCGCC, where ends up with AA*AA*AA*AA,
+which is not tight.  Interestingly, if we have AAAAATTCCCCC and
+AAGAATTCCGCC, then we get AA*AATTCC*CC -- to that extra T makes it
+tight.  Not sure if this is significant.
+
+\section{Algorithm}
+
+Let $s$ be a fixed spaced seed, $G$ be a fixed genomic sequence, and
+$f$ be a fixed integer $\geq 2$.  Let $f_s(i)$ be the sub-sequence at
+$G[i]$ inducted by the seed (with 0-matching characters removed).  So $s = 11011$ and
+the sequence at $G[i]$ is $AACGG$, the $f_s(i) = AAGG$.
+
+Let $H$ and $F$ be hash-tables mapping strings to integers:
+\begin{itemize}
+\item $H[f(i)]$ maps reduced sequences to counts.
+\item $F[f(i)]$ maps reduced sequences to family assignments.
+  (e.g. $f(i)$ is assigned to family $F[f(i)]$.
+
+Let $p(i)$ be the coorindate of $G$ the marks the beginning of the
+last instance of $f(i)$.  That is, if $s=11011$ and $G =
+ACCTCCAAACCTCC$, then $p(9) = 1$.  ($f(9) = CCCC$, and the largest
+$i < 9$ such that $f_s(i) = CCCC$ is $i=1$.  (Not all functions are
+pegged to a fixed $s$.)
+
+\begin{verbatim}
+num_fams <- 0 
+for i <- 0 to |G|:
+   seq = G[i:i+|s|]
+   H[seq] += 1
+   if H[seq] >= 2:
+      pseq =
+
+\end{verbatim}
+
 \section{Random Lemmas}
@@
 
-\begin{lemma}
-  Let $s$ be a seed, $r$ be a refd consistent with $s$,
-  and $r'$ be an refd consistent with $s$ such that $|r'| < |r|$.  
-\end{lemma}
-
 \end{document}

### 2015-07-09 manuscript/defs.tex

--- 2015-07-08 manuscript/defs.tex
@@
 \begin{definition}
+An $L$-mer is a string of length $l$.
+\end{definition}
+
+\begin{definition}
+Given a spaced seed $s$ of length $l$ and an $l$-mer $x$, the {\it seeded
+$l$-mer} $s'$ is created by removing all characters of $x$
+corresponding to $0$'s in $s$.  Let $\sigma_s(x)$ denote the seeded
+$l$-mer for $s$.
+\end{definition}
+Example: for $s=11011$ and $x = AACGG$, $\sigma_s(x) = AACC$.
+
+\begin{definition}
   A {\it repeat element family descriptor (refd)} is a
@@
 \begin{definition}
-  We say an refd $r$ {\it matches} a Genome $G$ at position $i$ if, for
+  We say an refd $r$ {\it matches} a genome $G$ at position $i$ if, for
   all $j$ such that $r_j \neq *$, $r_j = G_{i+j}$.
@@
 at positions $i=0$ and $i=3$, but not at any other $i$.
+
+\begin{observation}
+$s$ matches $r$ at $i$ if $\sigma_s(r[i:i+|s|])$ does not contain any
+  $*$ symbols.
+\end{observation}
+
+$\sigma(r[3:8]) = AAGG$ (hence a match at $i=3$), but $\sigma(r[2:7])
+= AA*G$ (hence no match at $i=2$).
 
@@
 
+\begin{definition}
+Given the refd $r$ of an elementary repeat family with set $S$, we say
+that $r$ is {\it tight} if, for each $i$ such that $r_i = *$, there
+exists two sequences defined by $S$ that have different bases in
+position $i$.
+\end{definition}
+In other words: $r$ is tight if it only uses $*$ symbols where it must
+to match everything sequence defined by $S$.
+
+{\bf Comments:} I had it in mind that the algorithm would always
+return a tight refd.  But in Carly's defense she gave the example:
+s=11011, AAAAATCCCCC, AAGAATCCGCC, where ends up with AA*AA*AA*AA,
+which is not tight.  Interestingly, if we have AAAAATTCCCCC and
+AAGAATTCCGCC, then we get AA*AATTCC*CC -- to that extra T makes it
+tight.  Not sure if this is significant.
+
+\section{From Thesis}
+
+In the following I'm going through definitions / parallels from Nate's
+thesis and seeing if I can create an analgy for spaced seeds.
+
+The following are things I wanted to try to prove that may or may not
+be useful.  In all of these I'm assuming a fixed genome $G$, a fixed
+frequence requriement $f$, and a fixed seed $s$ with length $l$ and
+weight $w$.
+
+\begin{lemma} 
+For any $l$-mer $x$, the seeded $l$-mer $\sigma_s(x)$ can be a member
+of at most $l$ different families.
+\end{lemma}
+My gut is that this is true, but not necessarily tight.  Maybe its
+$w$, or $l-w$?
+
+\begin{definition}
+Let $x$ and $y$ be two strings such that $x = a \dot b$, $y = b \dot
+c$, and $|b| = i$.  Then $x \circ_i y = a\dot b \dot c$.
+\end{definition}
+
+$AAACC \circ_2 CCGGG = AAACCGGG$.  $AAACC \circ_3 CCGGG$ is undefined.
+
+This is a modification of Nate's $x \circ y$ operator.  May need to be
+adapted for seeds.
+
 \section{Random Lemmas}
 
-The following are things I wanted to try to prove that may or may not
-be useful.
-
-\begin{lemma}
-  Let $s$ be a seed, $r$ be a refd consistent with $s$,
-  and $r'$ be an refd consistent with $s$ such that $|r'| < |r|$.  
-\end{lemma}
-
 \end{document}

### 2015-07-10 manuscript/defs2.tex
Adding defs2

\documentclass{article}
\usepackage{longtable}
\usepackage{ulem}
\pagestyle{myheadings}
\usepackage{ifthen}
\usepackage{color}

\setlength{\parindent}{0pt}
\setlength{\parskip}{1ex plus 0.5ex minus 0.2ex}

\newcommand{\red}[1]{{\color{red}#1}}
\newcommand{\blue}[1]{{\color{blue}#1}}
\newcommand{\cyan}[1]{{\color{cyan}#1}}
\newcommand{\magenta}[1]{{\color{magenta}#1}}

\newtheorem{definition}{Definition}
\newtheorem{observation}{Observation}
\newtheorem{lemma}{Lemma}

\begin{document}

\section{Definitions}

\subsection{REFDs}
In RAIDER, everything is defined in terms of $l$-mers.  hHence we need
to generlize slightly to allow the ambiguity character, so we will
create a description of a {\it repeat element family discriptior},
with an $l$-mer being a special case of that.

\begin{definition}
A \underline{repeat element family discriptor} (refd) is a sequence over
$\Sigma=\{A,C,G,T,*\}$, where $*$ denotes unspecified bases. 
\end{definition}

{\bf Note:} We will refer to sub-refds as consistent with the concept
of a substring, and denote then as consistent with Python.  (That is,
if $r =  AAAC*CGG'$, then the sub-refd $r_{3:6}$, or $r[3:6]$, is the refd
$C*C$).  Generally, we will define all standard string operators on
refds unless otherwise noted.  (e.g. $\cdot$ will be the concatenation
operator, $|r|$ will denote total length, etc...)

\begin{definition}
Given an refd, $|r|$ denotes the length of the string and $||r||$
denotes the number of non-$*$ characters in the string.
\end{definition}
For $r = A*A$, $|r| = 3$ and $||r|| = 2$.

\begin{definition}
An $l$-mer is an refd of length $l$.
\end{definition}

\begin{definition}
Let $r$ and $r'$ be two $l$-mers.
\begin{itemize}
\item $r$ and $r'$ are \underline{consistent} if,
  for any position $i$ in which they differ, one of them contains a
  $*$.  ($AAC*TT$ and $AA*GTT$ are consistent; $AAC*TT$ and $AAGGTT$
  are not.)
\item We will say $r$ \underline{generalizes} $r'$ if, wherever they differ,
  $r$ has a $*$.  Denote this as $r \vdash r'$.  ($AAC*TT \vdash AACGTT$, 
  $AAC*TT \not\vdash AA*CTT$.)
\item The \underline{merge} of $r$ and $r'$, denoted $r \oplus r'$, is
  the $l$-mer such that $s_i = r_i$ when $r_i=r'_i$,
  and $*$ otherwise. ($AAC*TT \oplus AA*GTT = AA**TT$.) 
\item The \underline{tight merge} of $r$ and $r'$, denoted $r \otimes r'$,
  is $r \oplus r'$ if $r$ and $r'$ are consistent, and undefined
  otherwise. ($AAC*TT \otimes AA*GTT = AA**GT$, but $AAC*TT \otimes AAAA*CGT$ is undefined.)
\item The \underline{compression} of consistent refds $r$ and $r'$, denoted $r\ominus r'$, 
  is the string $s$ where $s_i$ is: $*$ if both $r_i$ and $r'_i$ are
  $*$, other the non-$*$ character at $r_i$ or $r'_i$.  If $r$ is and
  $r'$ are not consistent, $r \ominus r'$ is undefined.  ($AA**C
  \ominus A**CC = AA*CC$.
\end{itemize}
\end{definition}

{\bf Comments:}
\begin{itemize}
\item $r \oplus r'$ is the string that finds the $s$ over all
$l$-mers $s$ that generalizes both $r$ and $r'$ which maximizes $||s||$.
\item $r \ominus r'$ is the string that finds the $s$ over all
  $l$-mers $s$ that can be gerlized t both $r$ and $r'$ which
  maximizes $||s||$.
\item It might makes sense to say $r' < r$ if $r$ generlizes $r$ --
  which would then indce a lattice.  I have no idea if this is
  actually interesting / useful, and am ignoring it for now.
\end{itemize}

\begin{definition}
Let $x$ and $y$ be REFDs. Define $x \circ y$ as follows.  Let $i$ be
the largest $i$ such that $x_{|x|-i:}$ and $y_{:i}$ are consistent, let
$b = x_{|x|-i:} \ominus y_{:i}$, let $a = x_{:|x|-i}$ and $c = y_{i:}$.  
$x \circ y = a \cdot b \cdot c$.
\end{definition}
{\bf Examples:}
\begin{itemize}
\item $AAACCC \circ C*CTTT = AACCC$.
\item $AAACCC \circ CTCTTT = AAATCCTCTTT$.
\item $AAACC* \circ CC*TTT = AAACC*TTT$.
\end{itemize}

\subsection{REFDs and Spaced Seeds}

\begin{definition}
Given an $l$-mer $r$ and a spaced seed $s$ of length $l$, The seeded
$l$-mer, denoted $\sigma_s(r)$, is created by removing all bases at
positions corresponding to 0s.
\end{definition}
Example: $\sigma_{11011}(AACGG) = AAGG$.

\begin{definition}
Given an $l$-mer $r$ and a spaced seed $s$ of length $l$, the
generlized $l$-mer (w.r.t. $s$), or $\gamma_s(r)$, is the refd created by replacing
all basses corresponding to 0s with * characters.
\end{definition}
Example: $\sigma_{110101}(AACGTG) = AA*G*G$.

\begin{definition}
Given a spaced seed $s$ of length $l$: 
\begin{itemize}
\item $s$ is {\it consistent} with an $l$-mer $r$ if, for all
$i$ such that $r_i = *$, $s_i = 0$.  
\item A spaced seed $s$ hits an refd $r$ at position $i$ if 
  $i <|r|-|s|$ and $s$ is consistent with $r_{i:i+|s|}$.
\end{itemize}
\end{definition}

\begin{definition} 
A spaced seed $s$ of length $l$ {\underline covers} and refd $r$ is, for every $j$,
there is an $i \leq j$ such that $s$ is consistent with the sub-refd
$r_{i:i+|l|}$.
\end{definition}
{\bf Examples:} Let $s = 11011$.
\begin{itemize}
\item $r = AAAAA*AA$. $s$ covers $r$ as we can place $s$ $i=0$ and
  $i=3$ and, between them, cover every character in $r$ without
  conclict between a 1 and a $*$.
\item $r = AAA*AA$.  $s$ does not cover $r$.  $s$ only hist at $i=1$,
  so there is no way to cover base 0.
\item $r = AAAAAAAAAA$.  $s$ covers $r$, since $s$ hits at every base.
\end{itemize}

\subsection{REFDs and Sequence Sets}
Looks like we will need to define the relationship between refds and
sets of sequences to get the elementary repeat sequence.

Assume a fixed genome $G$.

\begin{definition}
Let $r$ be an refd of length $l$ and let $S$ be a set of sequences,
each of length $l$.
\begin{itemize}
\item We say $r$ is \underline{consistent} with $S$ if $r$ is
  consistent with every sequence in $S$.
\item We say $r$ is \underline{tightly consistent} with $S$ if $r$ is
  consistent with $s$ and, for every $i$ such that $r_i=*$, there
  exists sequences $x,y \in S$ such that $x_i \neq y_i$.
\end{itemize}
That is: $r$ is tightly consistent with the set $S$ is $r$ only has
$*$ characters in those places it must.  (I thought this was going to
be necessary at one point, but am now doubting it.)

\subsection{Elementary Repeats}

\begin{definition}
Let $s$ be a fixed seed with length $l$, $f$ be a fixed integer, and $G$ be a fixed
genomic sequence.  An
\underline{Elementary Repeat Family} is a tuple $(r,S)$ where: 
\begin{enumerate}
\item $r$ is an refd such that $s$ covers $r$. 
\item $S$ is a set of genomic coordinates such that $|S| \geq f$.  Say
  that the sub-refds induced by $S$ are the substrings of $G$ starting
  at cordinates of $S$ and extended $|r|$ bases.
\item For every $i \in S$, $r$ is consistent with the string
  $G_{i:i+|r|}$.
\item Minimality condition.  Possibilities (I'm not sure which is
  right):
  \begin{itemize}
  \item There is no sub-refd $r'$ of $r$ that is covered by $s$ and is
     consistent some sequence $x$ that is sits outside the sequences
      induced by $S$.
  \item There is no $l$-mer $r'$ contained in $r$ that is consistence
    with $s$, such that $\sigma_s(r)'$ occurs anywhere in $G$ outside
    of the sequences induced by $S$.
 \end{itemize}
\item There is no refd $r'$ containing $r$ such that every sequence
  induced by $S$ is contained by some sequence consisted with $r'$.
\end{definition}

{\bf Comments:} For the last two points, I'm really not sure if the
substring relation is the correct relation between $r$ and $r'$.

\subsection{Lemmas}

Right now, I'm trying trying to think ofpossibly interesting
statements and prove them.

Assume a fixed seed $s$ of length $l$ and a fixed genome $G4.

\begin{lemma}
Let $r$ and $r'$ be two different $l$

\end{document}

### 2015-07-10 manuscript/defs2.tex
Updating defs2.tex

@@
 \end{itemize}
+\end{definition}
 That is: $r$ is tightly consistent with the set $S$ is $r$ only has
@@
   induced by $S$ is contained by some sequence consisted with $r'$.
+\end{enumerate}
 \end{definition}
@@
 
-Assume a fixed seed $s$ of length $l$ and a fixed genome $G4.
-
-\begin{lemma}
-Let $r$ and $r'$ be two different $l$
+Assume a fixed seed $s$ of length $l$ and a fixed genome $G$.
 

### 2015-07-13 manuscript/defs2.tex
Updating defs2

@@
 A \underline{repeat element family discriptor} (refd) is a sequence over
-$\Sigma=\{A,C,G,T,*\}$, where $*$ denotes unspecified bases. 
+$\Sigma=\{A,C,G,T,*\}$ (where $*$ denotes unspecified bases).
 \end{definition}
@@
 of a substring, and denote then as consistent with Python.  (That is,
-if $r =  AAAC*CGG'$, then the sub-refd $r_{3:6}$, or $r[3:6]$, is the refd
+if $r =  AAAC*CGG$, then the sub-refd $r_{3:6}$, or $r[3:6]$, is the refd
 $C*C$).  Generally, we will define all standard string operators on
@@
   $r'$ are not consistent, $r \ominus r'$ is undefined.  ($AA**C
-  \ominus A**CC = AA*CC$.
+  \ominus A**CC = AA*CC$.  But $AA**C \ominus AT**C$ is undefined
+  (because of the disagreement in the second position).
 \end{itemize}
@@
 \begin{itemize}
-\item $AAACCC \circ C*CTTT = AACCC$.
+\item $AAACCC \circ C*CTTT = AAACCCTTT$.
 \item $AAACCC \circ CTCTTT = AAATCCTCTTT$.
@@
 \begin{definition}
-Given an $l$-mer $r$ and a spaced seed $s$ of length $l$, The seeded
+Given an $l$-mer $r$ and a spaced seed $s$ of length $l$, the seeded
 $l$-mer, denoted $\sigma_s(r)$, is created by removing all bases at
@@
 \end{definition}
-Example: $\sigma_{110101}(AACGTG) = AA*G*G$.
+Example: $\gamma_{110101}(AACGTG) = AA*G*G$.
 
@@
 \begin{definition} 
-A spaced seed $s$ of length $l$ {\underline covers} and refd $r$ is, for every $j$,
+A spaced seed $s$ of length $l$ \underline{covers} and refd $r$ is, for every $j$,
 there is an $i \leq j$ such that $s$ is consistent with the sub-refd
@@
 \begin{itemize}
-\item $r = AAAAA*AA$. $s$ covers $r$ as we can place $s$ $i=0$ and
+\item $r = AAAAA*AA$. $s$ covers $r$ as we can place $s$ at $i=0$ and
   $i=3$ and, between them, cover every character in $r$ without
-  conclict between a 1 and a $*$.
-\item $r = AAA*AA$.  $s$ does not cover $r$.  $s$ only hist at $i=1$,
+  concflict between a 1 and a $*$.
+\item $r = AAA*AA$.  $s$ does not cover $r$.  $s$ only hits at $i=1$,
   so there is no way to cover base 0.
@@
 
+\subsection{REFD Decomposition}
+
+Given an REDF $r$ and a seed $s$ of length $l$ ($l \leq |r|$), we want
+to decompose $r$ into a set of overlapping $l$-mers.  These are the
+$l$-mers that cannot be repeated outside of the repeat.
+
+For example: if we have $G = \ldots AAAAACTTTTT\ldots AAAAAGTTTTT
+\ldots$ (notice: 6 As), with the seed $1101$, we get the refd $AAAAAA*TT*TT$, which
+deposes into $AA*AA$, $AA*TT$, and $TT*T$$.  If any string compatible
+with any of these three occurs anywhere else, it breaks the family.
+But if the string $AA*AC$ occurs somewhere else, it does not -- thats
+irrelevant.  (In this example the minimal decomposition is unique -- I'm not
+sure that this is always the case, or what multiple decompositions
+mean.)
+
+\begin{definition} 
+Given a spaced seed $s$ of length $l$ and a an refd $r$ that is
+covered by $s$, a \underline{decomposition} of $r$ w.r.t. $s$ is a set of coordinates $i$, $0 \leq i < |r|-l$, such
+that (a) for each $i \in S$, $s$ hits $r$ at $i$, and (b) $r$ is
+covered by $s$ using only those positions from $S$.
+\end{definition}
+
+Examples:
+\begin{itemize}
+\item $s=11011$, $r=AAAAA*CCCCC$, then the only decomposition is
+  ${0,3,6}$.
+\item $s=11011$, $r=AAAAAA*CCCCC$, then ${0,1,4,7}$ and ${0,4,7}$ are
+  both decompositions.
+\end{itemize}
+
+\begin{definition}
+A decomposition $S$ of $r$ with respect to $s$ is \underline{minimal}
+if no element of $S$ can be removed without breaking the decomposition
+properties. 
+\end{definition}
+In the above, the first set of the second example is not minimal.
+
 \subsection{Elementary Repeats}
+
+I want to get a definition of elementary repeats that paralleles the
+one we had.  (And, hopefully, is provably equivilent for spaced seeds
+without spaces.)  This is tricky: I'm not yet sure what the equivilent
+is for the minimality condition (there is no subsequence of an
+elementary repeat that occurs independently of the instances of the
+elementary repeat).  Lets look why.
+
+Suppose we have in the string out genome $AAAAATCCCCC$ and
+$AAAAAGCCCCC$.  Using the seed $s=11011$, we should get the refd
+$r=AA*AA*CC*CC$. Here are different ways of phrasing the minimality
+condition, and why they break:
+\begin{enumerate}
+\item No sub-refd can match any string in the genome outside of those
+  matching $r$.  Doesn't work: we can have an AAAAT occuring somewhere
+  else wihtout breaking this family, since the second $l$-mer is
+  irrelevant to construction.
+\item There cannot be any string $S'$, $|s'| < |r|$, occuring outside
+  the instances, such that $s'$ generalizes to a sub-refd $s$ or $r$.
+\end{enumerate}
 

### 2015-07-14 manuscript/defs2.tex
Updating defs2.txt

@@
 \subsection{REFDs}
-In RAIDER, everything is defined in terms of $l$-mers.  hHence we need
+In RAIDER, everything is defined in terms of $l$-mers.  Hence we need
 to generlize slightly to allow the ambiguity character, so we will
@@
 \begin{definition}
+\label{generalize}
 Given an $l$-mer $r$ and a spaced seed $s$ of length $l$, the
@@
 
+\begin{definition}
+\label{rmdecomp}
+Given an refd $r$ and a seed $s$ that covers $R4, the
+\underline{right-most} decomposition, denoted $\delta_s(r)$ is the decomposition $S$ created
+by:
+\begin{enuemrate}
+\item Set $i=0$.
+\item Let $S = S \cup \{i\}$.
+\item If $i == |r| - l$: quit.
+\item Let $j = max_{i < j < i+|l|}\{s \mbox{ hits } r \mbox{ at } j$.
+\item Goto (1).
+\end{enumerate}
+(In other words: Add $0$ to $S$, then keep taking then keep adding the
+right-most index you can where $s$ hits $r$ and doesn't leave a gap
+from the last hit.)
+
+{\bf Note:} I think later we should consder allowing it to leave a
+small gap to allow for insertions.
+
+{\bf Examples:}
+\begin{itemize}
+\item $\delta_{11011}(AA*AATT*TT) = \{0,5\}$.
+\item $\delta_{11011}(AA*AA*TT*TT) = \{0,3,6\}$.
+\end{itemize}
+
+\begin{observation}
+A few observations about the decompositions of $r$ w.r.t. to seed $s$:
+\begin{itemize}
+\item Any decomposition must containt $0$ and $|r|-|s|$ (as these are
+  the only way the first and last base can be covered).
+\item $\delta_s(r)$ exists if and only if there is some decompostion of
+  $r$ w.r.t $s$, which is true if and only if $s$ covers $r$.
+\item $\delta_s(r)$ is unique if it exists.
+\end{itemize}
+
 \subsection{Elementary Repeats}
-
-I want to get a definition of elementary repeats that paralleles the
-one we had.  (And, hopefully, is provably equivilent for spaced seeds
-without spaces.)  This is tricky: I'm not yet sure what the equivilent
-is for the minimality condition (there is no subsequence of an
-elementary repeat that occurs independently of the instances of the
-elementary repeat).  Lets look why.
-
-Suppose we have in the string out genome $AAAAATCCCCC$ and
-$AAAAAGCCCCC$.  Using the seed $s=11011$, we should get the refd
-$r=AA*AA*CC*CC$. Here are different ways of phrasing the minimality
-condition, and why they break:
-\begin{enumerate}
-\item No sub-refd can match any string in the genome outside of those
-  matching $r$.  Doesn't work: we can have an AAAAT occuring somewhere
-  else wihtout breaking this family, since the second $l$-mer is
-  irrelevant to construction.
-\item There cannot be any string $S'$, $|s'| < |r|$, occuring outside
-  the instances, such that $s'$ generalizes to a sub-refd $s$ or $r$.
-\end{enumerate}
 
@@
 genomic sequence.  An
-\underline{Elementary Repeat Family} is a tuple $(r,S)$ where: 
+\underline{Elementary Repeat Family} is a tuple $F=(r,s,S)$ where: 
 \begin{enumerate}
-\item $r$ is an refd such that $s$ covers $r$. 
+\item $s$ is a seed and $r$ is an refd such that $s$ covers $r$. 
 \item $S$ is a set of genomic coordinates such that $|S| \geq f$.  Say
-  that the sub-refds induced by $S$ are the substrings of $G$ starting
+  that the sequences induced by $S$ are the substrings of $G$ starting
   at cordinates of $S$ and extended $|r|$ bases.
@@
   $G_{i:i+|r|}$.
-\item Minimality condition.  Possibilities (I'm not sure which is
-  right):
+\item Consider the sub-refds of $r$ used for $\delta_r(s)$ (definition
+  \ref~{rmdecomp$}). None of these strings may generalize (definition
+  \ref~{generalize}) any substring of $G$ outside ofthose induced by
+  $S$.  (Minimiality condition.)
+\item There is no refd $r'$ such that all of the following are true:
 \begin{itemize}
-  \item There is no sub-refd $r'$ of $r$ that is covered by $s$ and is
-     consistent some sequence $x$ that is sits outside the sequences
-      induced by $S$.
-  \item There is no $l$-mer $r'$ contained in $r$ that is consistence
-    with $s$, such that $\sigma_s(r)'$ occurs anywhere in $G$ outside
-    of the sequences induced by $S$.
+  \item $r$ generalizes to some proper(?) substring of $v$ of $r'$.
+  \item The strings in $\delta_r(s)$ make up part of the cover of
+    $\delta_{s}(r')$.
+  \item Every string induced in $G$ by $S$ is covered by a string that
+    generalizes to $r'$.
 \end{itemize}
-\item There is no refd $r'$ containing $r$ such that every sequence
-  induced by $S$ is contained by some sequence consisted with $r'$.
 \end{enumerate}
@@
 
-{\bf Comments:} For the last two points, I'm really not sure if the
-substring relation is the correct relation between $r$ and $r'$.
+{\bf Comments:} 
+\begin{itemize}
+\item I'm not sure if the maximality condition is what we want.
+\item I haven't done anything to control the number of * symbols.
+  Currently, we can have multipled refds for the same family by
+  generalizing.  I'm not sure if this matters.
+\end{itemize}
+
+\begin{definition}
+Given a repeat family $F=(r,S)$, the length of the family $|F|=|r|$,
+and the size of the family $||F||=|S|$, the number of elements in the
+family.
+\end{definiton}
+
+Alternative definition of elementary repeats:
+\begin{definition}
+Let $F=(r,s,S)$ be a repeat family with refd $r$, spaced seed $s$, and
+coordinate set $S$ such that $s$ covers $r$. Let $C$ be the set of all
+sub-refds corresponding to the right-most decompostion of $S$
+w.r.t. $r$.  Then $F$ is a repeat element family (w.r.t. to $s$) if:
+\begin{enumerate}
+\item $|C| > 0$.
+\item For each $i \in S$, $G[i:i+|r|]$ generalizes to $r$.
+\item For each $w \in C$, the number of strings in $G$ that generlize
+  to $w$ is $|S|$.  (Minimality -- but only applies to the
+  decomposition.)
+\item $F$ is maximal w.r.t to this definition.  Specifically, we
+  cannot extend $r$ in either direction and have it covering exactly
+  the same sequences.  (Need to think about this.)
+\end{enumerate}
 
@@
 
-Right now, I'm trying trying to think ofpossibly interesting
-statements and prove them.
+Right now, I'm trying trying to think of possibly interesting
+statements and prove them.  Mostly just to get into the swing of this.
 
-Assume a fixed seed $s$ of length $l$ and a fixed genome $G$.
+\begin{defintion} 
+Conisider to elementary repeat families $F_1=(r_1,s,S_1)$ and
+$F_2=(r_2,s,S_2)$.  These families overlap if there exist
+some $i \in F_1$ and $j \in F_2$ such that either $0 \leq j-i < |r_1|$
+or $0 \leq i-j < |r_2|$.
+\end{definition}
+
+That is, you have one sequence induced by each family such that they
+actually overlap on $G$.
+
+\begin{lemma}
+Given repeat families $F_1$ and $F_2$ that overlap, and w.l.o.g. 
+pick the $i$ and $j$ such that $0 \leq j-i < |r_1|$.  Then there exist
+an offset vale $o = j-i$ such that, for any $i' \in S_1$, $j' \ in S_2$
+s.t. $j' \geq i'$, either $j' - i' > |r_1|$ or $j' - i' = o$.
+\end{lemma}
+(In otherwords, corresponding sequences are offset by the same amount
+every time.)
+
+\begin{lemma} 
+If repeat element families $F_1$ and $F_2$ overlap, then $||F_1|| \neq
+||F_2||$.  
+\end{lemma}
 

### 2015-07-20 manuscript/defs3.tex
defs3.txt

\documentclass{article}
\usepackage{longtable}
\usepackage{ulem}
\pagestyle{myheadings}
\usepackage{ifthen}
\usepackage{color}

\setlength{\parindent}{0pt}
\setlength{\parskip}{1ex plus 0.5ex minus 0.2ex}

\newcommand{\und}{\underline}

\newcommand{\red}[1]{{\color{red}#1}}
\newcommand{\blue}[1]{{\color{blue}#1}}
\newcommand{\cyan}[1]{{\color{cyan}#1}}
\newcommand{\magenta}[1]{{\color{magenta}#1}}

\newtheorem{theorem}{Theorem}
\newtheorem{notation}{Notation}
\newtheorem{definition}{Definition}
\newtheorem{observation}{Observation}
\newtheorem{lemma}{Lemma}

\begin{document}

\subsection{REFDs}

\begin{definition}
  A \underline{sequence descriptor} (sd) is a string over $\Sigma=\{A,C,G,T,*\}$.
\end{definition}

We will allow any normal string operator to apply to sequence descriptors.  In specific, for a s.d. $r$:
\begin{itemize}
\item $r_{i,i+l}$, or $r[i:i+l]$, is the length $l$ sub-sd (substring) of $r$ starting at $i$.
\item $|r|$ is the number of characters in $r$.
\item $||r||$ is the number of non-* characters in $r$.
\item We will use $l$-mer to denote a length $l$ sd.
\end{itemize}

\begin{definition}
  If $r_1$ and $r_2$ are sds of the same length:
\begin{enumerate}
\item $r_1$ and $r_2$ are \und{consistent} if, for any position $i$ in
  which they differ, one of them is a $*$. 
\item $r_2$ \und{generalizes} $r_1$ (and $r_1$ is \und{generlized} by $r_2$),
  if, for any position in which they are different, $r_2$ has a $*$.
  \end{enumerate}
\end{definition}

\begin{definition}
  The frequency of an s $r$ in genome sequence $G$ is the number of
  $i$ such that $G_{i:i+|r|}$ is compatible with $r$.  
\end{definition}

\begin{notation}
Let $r_1$ and $r_2$ be sds.  We say:
\begin{itemize}
\item $r_1 < r_2$ if $r_1$ is generalized by $r_2$.  Extended to
  $\leq$, $\geq$, and $>$ as normal. 
\item $r_1 || r_2$ ($r_1$ is ``parallel'' to $r_2$) if $r_1$ and $r_2$ are
  compatible, but neither generalizes the other.
\item $freq_G(r)$ is number of substrings in $G$ that are compatable with $r$.
\end{itemize}
\end{notation}

{\bf Comment:} This forms a lattice over the set of all
$l$-length sds.  I have no idea if this is important/useful.

\subsection{REFDs and Spaced Seeds}

\begin{definition}
Given a spaced seed of length $l$ and an $l$-mer $r$, we
\und{generlize} $r$
w.r.t $s$ by replacing $r_i$ with a * for all $i$ such that $s_i = 0$.
\end{definition}
Example: We generalize $AAAAAA$ w.r.t. $110101$ to $AA*A*A$.

\begin{definition}
  Let $s$ be a spaced seed of length $l$.

  \begin{itemize}
  \item We say $s$ and an $l$-mer $r$ are \und{consistent} if, when $s_i = 0$, $r_i = *$.
  \item We say seed $s$ \und{hits} sd $r$ at position $i$ such that
    $s$ is consistent with $r_{i:i+|s|}$ ($i \leq |r| - |s|$).
  \item We say seed $s$ \und{covers} sd $r$ if, for every $j$, $0 \leq
    j < |r|$, there is an $i$ such that $s$ hits $r$ at $i$ and covers
    position $j$ in doing so.
  \end{itemize}
\end{definition}

Example: For $s=11011$ and $r=AAAAA*TTTTT$, $s$ covers $r$, hitting at
$\{0,3,6\}$.  $s$ does not cover $AAAAA**TTTTT$, as there is no
position at which $s$ hits and extens over $i=5$.

\begin{definition}
  Let $r$ be an sd and $s$ be a seed that covers $r$.
\begin{itemize}
\item The \und{decompostion} of $r$ w.r.t $s$ is set of all $i$ such
  that $s$ hits $r$ at position $i$.
\item The \und{seqeunce decomposition} of $r$ w.r.t. $s$ is the set
  of all sds $r_{i:i+|s|}$ for each $i$ in the decomposition.
\item The \und{generalized sequence decomposition} of $r$ w.r.t. $s$
  is the set of generalizations of $w$ for all $w$ in the sequence decomposition.
\end{itemize}
\end{definition}

Examples: For $s=11011$ and $r = TAAAAA*CCCCCC$, the decomposition is
$\{1,2,5,8,9\}$, the sequence decomposition is $\{TAAAA, AAAAA, AA*CC,
CCCCC\}$ (note the two sequences a 8 and 9 are the same, so only
included once), and the generlized sequence decomposition is $\{TA*AA,
AA*AA, AA*CC, CC*CC\}$.

\begin{notation}
  \begin{itemize}
  \item Let $\gamma_s(r)$ denote the generalization of $r$ w.r.t. $s$
  (undefined if $|r| \neq |s|$).
  \item Let $\delta_s(r)$ denote the decomposition of $r$ w.r.t. $s$.
  \item Let $\Delta_s(r)$ denote the sequence decomposition of $r$ w.r.t. to
  $s$.
  \item Let $\Gamma_s(r)$ denote the generalized seqeunce decomposition of
  $r$ w.r.t. $s$.
  \end{itemize}
\end{notation}

{\bf Note:} I messed around with other decompositions (subsets of the
decomposition, minimal decompositions, right-most decompositions), but
this is the only one that seemed to be useful so far.

\subsection{Elementary repeats}

\begin{definition}
  \label{ERdef}
  Given a fixed seed $s$, genome $G$, and frequency $f$, an sd $r$ is
  a \underline{repeat element descriptor} (red) if the following
  holds true:
  \begin{enumerate}
  \item $s$ covers $r$.
  \item $freq_G(r) \geq f$.
  \item $freq_G(w) = freq_G(r)$ for all $w \in \Gamma_s(r)$.
  \item There does not exist an $r'$ with a substring $w$ such that
    $r \leq w$ and $r'$ meets conditions 1-3.
  \end{enumerate}
\end{definition}
{\bf Questions:} On (4): what if there exists a substring $w$ of $r'$
such that $r || w$?  What if $w$ is not a proper substring?  (That is,
$r = w$.)

\begin{definition}
  A {\it repeat element family} is a tuple $(r,S)$ where $r$ is a
  repeat element descriptor and $S$ is the set of coordinates of $G$
  where there is a sequence compatiable with $r$.
\end{definition}

\subsection{Equivilance proof}

We want to prove that our definition, when using a seed with no
spaces, is equivilent to the Z\&L definition.  Recall that in Nate's
there was an alternate definition proved equivilent to Z\&L.

\begin{definition}
\label{NateDef}
Given a genomic sequence $G$ and integers $f$ and $l$, a seqeunce $r$
is an {\it elementary repeat} if we can write $r = x_0 \cdot x_2 \cdot
\cdot x_k$ $where $|x_i|=l$, $k = |r|-|l|$, and:
\begin{enumerate}
\item $k \geq 0$.  (There is at least one $l$-mer substring).
\item $freq_G(r) \geq f$
\item $freq_G(x_i) = freq_G(r)$
\item There does not exist an $y$, $|y|=l$ such that either $y \cdot
  r$ or $r \cdot y$ meets conditions 1-3.
\end{enuemrate}
\end{definition}
Recall that $x \circ y$ is defined only when $x$ has an $|x|-1$ length suffix
that is a prefix of $y$, and represents their ``merging''.
(e.g. $ACCC \circ CCCG = ACCCG$.)

\begin{theorem}
  Consider a genome $G$, an integer $f$, and a seed $s$ such that $s$
  has no zero chartacters.  $r$ is an elementary
  repeat descriptor for $s$ (by Definition~\label{ERdef}) if and only if $r$
  is an elementary repeat for $l=|s|$ (by
  Definition~\label{NateDef})).
\end{theorem}

{\bf Proof:}

($Leftarrow$) First we observe that if $s$ has no zeros, and covers
$r$, then $r$ has no * characters.  From this it follows that
$\Gamma_s(r) = \{r_0,r_1,\cdtos,r_{k}$ where $|r_i|=l$ and $k=r-l$ --
every $l$-mer contained with $r$.  From this it follows that $r = r_0 \circ
r_1 \

Now we prove each of the four conditions of Nate's definiciton:
\begin{enumerate}
  \item Since $r$ is covered by $s$, $|r| \geq |s|$, so $k =
    |r|-|s|+1 > 0$.
  \item Since $r$ contains no * symbols, the only string it
    is compatible with is itself.  Meaning it has the sme frequency by
    either definition (sd or string).  Hence (2) of Nate's
    definition following directly from (2) of out definition.
  \item Follows immediately from the fact that every $r_i$is a
    member of $\Delta_G(r)$, and hence must have the right frequency.
   \item Argue by contradiction: suppose there were such a $y$ (assume
     w.l.g. it follow $r$, and let $r' = r \circ x$.  Since Nate's
     definition tells us $r'$ matches a string around every occurance
     of a match to $r$, this means $r'$ is (trivially) compatable to a
     substring around every occurance of a match to $r'$ -- to $r'$
     could not have been a repeat element descriptor.
\end{enumerate}

($Rightarrow$) This is the same argument backwards, and not worth
writing out.
\end{enumerate}

\subsection{Basic Lemmas}

\begin{definition}
  Given a repetitive element descriptor $r$ for seed $s$ of length
  $l$, we say a length $l$ s.d. $w$ belongs to $r$ is $w$ is
  consistent with so $w \in 
  \Gamma_s(r)$.  (That is: if there is a a string in the generlized
  decomposition of $r$ that is consistent with $r$.
\end{definition}

Using this definition of {\it belongs}, all of our lemmas and
observations still holds:
\begin{lemma}
  An $l$-mer belongs to at most one elementary repeat family.
\end{lemma}
If the $l$-mer belongs, then it subject to condition (3) of the
definition, which would be violated if the $l$-mer belonged to two
different families.

\begin{observation}
  Any $l$-mer that occurs $\geq f$ times either is an elementary
  repeat descriptor or belongs to one.
\end{observation}
If the $l$-mer occurs at least $f$ times is automatically meets
criteria 1-3.  So the only thing that could stop it from being an
elementary repeat is (4) -- meaning its contained within one.

\begin{lemma}
  Let $r$ be a sd that maximially matches $G$ in two locations.  Than
  $r$ may be an elementary repeat descriptior, or may contain an
  elementary repeat descriptor -- but it is not contained within
  another elementary repeat descriptor.  (That is, (4) must hold.)
\end{lemma}
Any elementary repeat descriptor containing it would have to hit at
both location, which we just said isn't possible.

{\bf COMMENT:} HMM...I'm worried about this one.  What if we extneded
by a whole sead length?  Could that matter?

\begin{lemma}
  If two $l$-mers appear blog to the same elementary repeat
  descriptor, then they have the same frequence.

\end{lemma}
Follows directly from the definition.

From here is should be very easy to prove that Carly's algorithm
returns an elementary repeat descriptors for each family.

\subsection{Variations}

Suppose $r_1$ and $r_2$ are repeat element descriptors for the same
set of instances.  What are their potential relationships?

Let $s=11011$ and $G = \cdots AAAAACGGGGG \cdots AAAAATGGGGG\cdots$.
Then possible repeat element descriptors are $11111*11111$,
$11*11*111111$, $11111*11*11$, and $11*11*11*11$.  So we can have $r_1
< r_2$ or $r_1 || r_2$.

For $11011$ and $G = \cdots ACCCCCGTTTTTA\cdots ACCCCCATTTTTA$, we can
have $AC*CCC*TT*TTA$ or $ACC*CC*TTT*TA$$, among others.

What do we know about the r.e.d. that Carly's algorithm will pick?

\end{document}

### 2015-07-21 manuscript/BMC/bmc_article.tex
Adding BMC directorY

%% BioMed_Central_Tex_Template_v1.06
%%                                      %
%  bmc_article.tex            ver: 1.06 %
%                                       %

%%IMPORTANT: do not delete the first line of this template
%%It must be present to enable the BMC Submission system to
%%recognise this template!!

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                     %%
%%  LaTeX template for BioMed Central  %%
%%     journal article submissions     %%
%%                                     %%
%%          <8 June 2012>              %%
%%                                     %%
%%                                     %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                                                 %%
%% For instructions on how to fill out this Tex template           %%
%% document please refer to Readme.html and the instructions for   %%
%% authors page on the biomed central website                      %%
%% http://www.biomedcentral.com/info/authors/                      %%
%%                                                                 %%
%% Please do not use \input{...} to include other tex files.       %%
%% Submit your LaTeX manuscript as one .tex document.              %%
%%                                                                 %%
%% All additional figures and files should be attached             %%
%% separately and not embedded in the \TeX\ document itself.       %%
%%                                                                 %%
%% BioMed Central currently use the MikTex distribution of         %%
%% TeX for Windows) of TeX and LaTeX.  This is available from      %%
%% http://www.miktex.org                                           %%
%%                                                                 %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%% additional documentclass options:
%  [doublespacing]
%  [linenumbers]   - put the line numbers on margins

%%% loading packages, author definitions

%\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
\documentclass{bmcart}

%%% Load packages
%\usepackage{amsthm,amsmath}
%\RequirePackage{natbib}
%\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
%\RequirePackage{hyperref}
\usepackage[utf8]{inputenc} %unicode support
%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                             %%
%%  If you wish to display your graphics for   %%
%%  your own use using includegraphic or       %%
%%  includegraphics, then comment out the      %%
%%  following two lines of code.               %%
%%  NB: These line *must* be included when     %%
%%  submitting to BMC.                         %%
%%  All figure files must be submitted as      %%
%%  separate graphics through the BMC          %%
%%  submission process, not included in the    %%
%%  submitted article.                         %%
%%                                             %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\def\includegraphic{}
\def\includegraphics{}

%%% Put your definitions there:
\startlocaldefs
\endlocaldefs

%%% Begin ...
\begin{document}

%%% Start of article front matter
\begin{frontmatter}

\begin{fmbox}
\dochead{Research}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the title of your article here     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\title{Fast de novo transpoible element annotation}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors here                   %%
%%                                          %%
%% Specify information, if available,       %%
%% in the form:                             %%
%%   <key>={<id1>,<id2>}                    %%
%%   <key>=                                 %%
%% Comment or delete the keys which are     %%
%% not used. Repeat \author command as much %%
%% as required.                             %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\author[
   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
   noteref={n1},                        % id's of article notes, if any
   email={figuernd@miamiOH.edu}   % email address
]{\inits{ND}\fnm{Nathaniel D} \snm{Figueroa}}
\author[
   addressref={aff1},
   email={schaefce@miamiOH.edu}
]{\inits{SE}\fnm{Carly E} \snm{Schaeffer}}
\author[
   addressref={aff1,aff2},
   email={liux17@miamiOH.edu}
]{\inits{X}\fnm{Xiaolin} \snm{Liu}}
\author[
  addressref={aff1,aff2,aff3,aff4},
  corref={aff1},
  noteref={n2},
  email={karroje@miamiOH.edu}
]{\inits{JE}\fnm{John E} \snm{Karro}}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter the authors' addresses here        %%
%%                                          %%
%% Repeat \address commands as much as      %%
%% required.                                %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\address[id=aff1]{%                           % unique id
  \orgname{Department of Computer Science and Software Engineering}, % university, etc
  %\street{Miami University},                     %
  %\postcode{}                                % post or zip code
  %\city{Oxford},                              % city
  %\cny{USA}                                    % country
}
\address[id=aff2]{%                           % unique id
  \orgname{Center for  Molecular and Structural Biology},
  %\street{Miami University},                     %
  %\postcode{}                                % post or zip code
  %\city{Oxford},                              % city
  %\cny{USA}                                    % country
}
\address[id=aff3]{%                           % unique id
  \orgname{Department of Statistics}, % university, etc
  %\street{Miami University},                     %
  %\postcode{}                                % post or zip code
  %\city{Oxford},                              % city
  %\cny{USA}                                    % country
}
\address[id=aff4]{%                           % unique id
  \orgname{Department of Microbiology}, % university, etc
  \street{Miami University},                     %
  %\postcode{}                                % post or zip code
  \city{Oxford},                              % city
  \cny{USA}                                    % country
}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Enter short notes here                   %%
%%                                          %%
%% Short notes will be after addresses      %%
%% on first page.                           %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{artnotes}
%\note{Sample of title note}     % note to the article
  \note[id=n1]{Equal contributor} % note, connected to author
  \note[id=n2]{Corresponding Author}
\end{artnotes}

\end{fmbox}% comment this for two column layout

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Abstract begins here                 %%
%%                                          %%
%% Please refer to the Instructions for     %%
%% authors on http://www.biomedcentral.com  %%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{abstractbox}

\begin{abstract} % abstract
\parttitle{First part title} %if any
Text for this section.

\parttitle{Second part title} %if any
Text for this section.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The keywords begin here                  %%
%%                                          %%
%% Put each keyword in separate \kwd{}.     %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{keyword}
\kwd{sample}
\kwd{article}
\kwd{author}
\end{keyword}

% MSC classifications codes, if any
%\begin{keyword}[class=AMS]
%\kwd[Primary ]{}
%\kwd{}
%\kwd[; secondary ]{}
%\end{keyword}

\end{abstractbox}
%
%\end{fmbox}% uncomment this for twcolumn layout

\end{frontmatter}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% The Main Body begins here                %%
%%                                          %%
%% Please refer to the instructions for     %%
%% authors on:                              %%
%% http://www.biomedcentral.com/info/authors%%
%% and include the section headings         %%
%% accordingly for your article type.       %%
%%                                          %%
%% See the Results and Discussion section   %%
%% for details on how to create sub-sections%%
%%                                          %%
%% use \cite{...} to cite references        %%
%%  \cite{koon} and                         %%
%%  \cite{oreg,khar,zvai,xjon,schn,pond}    %%
%%  \nocite{smith,marg,hunn,advi,koha,mouse}%%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%%%%%%%%%%%%%%%%%%%%%%%% start of article main body
% <put your article body there>

%%%%%%%%%%%%%%%%
%% Background %%
%%
\section*{Content}
Text and results for this section, as per the individual journal's instructions for authors. %\cite{koon,oreg,khar,zvai,xjon,schn,pond,smith,marg,hunn,advi,koha,mouse}

\section*{Section title}
Text for this section \ldots
\subsection*{Sub-heading for section}
Text for this sub-heading \ldots
\subsubsection*{Sub-sub heading for section}
Text for this sub-sub-heading \ldots
\paragraph*{Sub-sub-sub heading for section}
Text for this sub-sub-sub-heading \ldots
In this section we examine the growth rate of the mean of $Z_0$, $Z_1$ and $Z_2$. In
addition, we examine a common modeling assumption and note the
importance of considering the tails of the extinction time $T_x$ in
studies of escape dynamics.
We will first consider the expected resistant population at $vT_x$ for
some $v>0$, (and temporarily assume $\alpha=0$)
%
\[
 E \bigl[Z_1(vT_x) \bigr]= E
\biggl[\mu T_x\int_0^{v\wedge
1}Z_0(uT_x)
\exp \bigl(\lambda_1T_x(v-u) \bigr)\,du \biggr].
\]
%
If we assume that sensitive cells follow a deterministic decay
$Z_0(t)=xe^{\lambda_0 t}$ and approximate their extinction time as
$T_x\approx-\frac{1}{\lambda_0}\log x$, then we can heuristically
estimate the expected value as
%
\begin{eqnarray}\label{eqexpmuts}
E\bigl[Z_1(vT_x)\bigr] &=& \frac{\mu}{r}\log x
\int_0^{v\wedge1}x^{1-u}x^{({\lambda_1}/{r})(v-u)}\,du
\nonumber\\
&=& \frac{\mu}{r}x^{1-{\lambda_1}/{\lambda_0}v}\log x\int_0^{v\wedge
1}x^{-u(1+{\lambda_1}/{r})}\,du
\nonumber\\
&=& \frac{\mu}{\lambda_1-\lambda_0}x^{1+{\lambda_1}/{r}v} \biggl(1-\exp \biggl[-(v\wedge1) \biggl(1+
\frac{\lambda_1}{r}\biggr)\log x \biggr] \biggr).
\end{eqnarray}
%
Thus we observe that this expected value is finite for all $v>0$ (also see \cite{koon,khar,zvai,xjon,marg}).
%\nocite{oreg,schn,pond,smith,marg,hunn,advi,koha,mouse}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                                          %%
%% Backmatter begins here                   %%
%%                                          %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\begin{backmatter}

\section*{Competing interests}
  The authors declare that they have no competing interests.

\section*{Author's contributions}
    Text for this section \ldots

\section*{Acknowledgements}
  Text for this section \ldots
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                  The Bibliography                       %%
%%                                                         %%
%%  Bmc_mathpys.bst  will be used to                       %%
%%  create a .BBL file for submission.                     %%
%%  After submission of the .TEX file,                     %%
%%  you will be prompted to submit your .BBL file.         %%
%%                                                         %%
%%                                                         %%
%%  Note that the displayed Bibliography will not          %%
%%  necessarily be rendered by Latex exactly as specified  %%
%%  in the online Instructions for Authors.                %%
%%                                                         %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% if your bibliography is in bibtex format, use those commands:
\bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
\bibliography{bmc_article}      % Bibliography file (usually '*.bib' )
% for author-year bibliography (bmc-mathphys or spbasic)
% a) write to bib file (bmc-mathphys only)
% @settings{label, options="nameyear"}
% b) uncomment next line
%\nocite{label}

% or include bibliography directly:
% \begin{thebibliography}
% \bibitem{b1}
% \end{thebibliography}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Figures                       %%
%%                               %%
%% NB: this is for captions and  %%
%% Titles. All graphics must be  %%
%% submitted separately and NOT  %%
%% included in the Tex document  %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%%
%% Do not use \listoffigures as most will included as separate files

\section*{Figures}
  \begin{figure}[h!]
  \caption{\csentence{Sample figure title.}
      A short description of the figure content
      should go here.}
      \end{figure}

\begin{figure}[h!]
  \caption{\csentence{Sample figure title.}
      Figure legend text.}
      \end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Tables                        %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

%% Use of \listoftables is discouraged.
%%
\section*{Tables}
\begin{table}[h!]
\caption{Sample table title. This is where the description of the table should go.}
      \begin{tabular}{cccc}
        \hline
           & B1  &B2   & B3\\ \hline
        A1 & 0.1 & 0.2 & 0.3\\
        A2 & ... & ..  & .\\
        A3 & ..  & .   & .\\ \hline
      \end{tabular}
\end{table}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%%                               %%
%% Additional Files              %%
%%                               %%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\section*{Additional Files}
  \subsection*{Additional file 1 --- Sample additional file title}
    Additional file descriptions text (including details of how to
    view the file, if it is in a non-standard format or the file extension).  This might
    refer to a multi-page table or a figure.

  \subsection*{Additional file 2 --- Sample additional file title}
    Additional file descriptions text.

\end{backmatter}
\end{document}

### 2015-07-21 manuscript/BMC/bmc_article.tex
bmc article

@@
 %%%%%%%%%%%%%%%%%%%%%%%%% start of article main body
-% <put your article body there>
+\section({Introduction}
+Here we introduce RAIDER (Rapid Ab Initio Detection of Elementary
+Repeat), designed to enable the quick annoation of transposible
+elements and repetitive DNA without requiring a pre-compiled template
+library.  Transposible elements, or TEs, are present in almost every
+higher order genome, covering as much as 45\% of the human genome and
+90\% of the maize genome [CHECK THIS] [CITE].  Given their prevelence,
+identiication of TEs is important not only for those who study them, but for
+anyone trying to understand a geome.  If not filtered, TEs can trigger
+hug numbers of false positives in the course of applying automated
+gene finding techniques, as well as blow up runtime.  [CITE: Jiang]
+So identifying and masking them in a must for any genome being
+studied.
 
-%%%%%%%%%%%%%%%%
-%% Background %%
-%%
-\section*{Content}
-Text and results for this section, as per the individual journal's instructions for authors. %\cite{koon,oreg,khar,zvai,xjon,schn,pond,smith,marg,hunn,advi,koha,mouse}
-
-\section*{Section title}
-Text for this section \ldots
-\subsection*{Sub-heading for section}
-Text for this sub-heading \ldots
-\subsubsection*{Sub-sub heading for section}
-Text for this sub-sub-heading \ldots
-\paragraph*{Sub-sub-sub heading for section}
-Text for this sub-sub-sub-heading \ldots
-In this section we examine the growth rate of the mean of $Z_0$, $Z_1$ and $Z_2$. In
-addition, we examine a common modeling assumption and note the
-importance of considering the tails of the extinction time $T_x$ in
-studies of escape dynamics.
-We will first consider the expected resistant population at $vT_x$ for
-some $v>0$, (and temporarily assume $\alpha=0$)
-%
-\[
- E \bigl[Z_1(vT_x) \bigr]= E
-\biggl[\mu T_x\int_0^{v\wedge
-1}Z_0(uT_x)
-\exp \bigl(\lambda_1T_x(v-u) \bigr)\,du \biggr].
-\]
-%
-If we assume that sensitive cells follow a deterministic decay
-$Z_0(t)=xe^{\lambda_0 t}$ and approximate their extinction time as
-$T_x\approx-\frac{1}{\lambda_0}\log x$, then we can heuristically
-estimate the expected value as
-%
-\begin{eqnarray}\label{eqexpmuts}
-E\bigl[Z_1(vT_x)\bigr] &=& \frac{\mu}{r}\log x
-\int_0^{v\wedge1}x^{1-u}x^{({\lambda_1}/{r})(v-u)}\,du
-\nonumber\\
-&=& \frac{\mu}{r}x^{1-{\lambda_1}/{\lambda_0}v}\log x\int_0^{v\wedge
-1}x^{-u(1+{\lambda_1}/{r})}\,du
-\nonumber\\
-&=& \frac{\mu}{\lambda_1-\lambda_0}x^{1+{\lambda_1}/{r}v} \biggl(1-\exp \biggl[-(v\wedge1) \biggl(1+
-\frac{\lambda_1}{r}\biggr)\log x \biggr] \biggr).
-\end{eqnarray}
-%
-Thus we observe that this expected value is finite for all $v>0$ (also see \cite{koon,khar,zvai,xjon,marg}).
-%\nocite{oreg,schn,pond,smith,marg,hunn,advi,koha,mouse}
+The most well known tools for repeat identification ar RepeatMasker
+and [Eddy's tool] [CITE], both of which work on the principle of a
+library-based search.  That is, to final transposable elements of a
+given family,
 

### 2015-07-22 manuscript/BMC/bmc_article.tex
Updating defs / manuscript

@@
 and [Eddy's tool] [CITE], both of which work on the principle of a
-library-based search.  That is, to final transposable elements of a
-given family,
+library-based search.  That is, to find transposable elements of a
+given family, the tools require some description of sequence in the
+family (in the form of a model sequence for blasting, or a pHMM) that
+is used as the basis for identification.  This begs the obvious
+question: much like we ask how the snow plow driver gets to work, or
+how dictinary authors look up the spelling of a word [CITE], we must ask how the
+tools compile their libraries.  How do we compile the libraries of
+transposible elements, or discover new transpoable element families.
+
+Within mammalian speices we can, to some extent, rely on homology
+relationships to port libraries across species.  A library description
+of a human ALU sequence can likely be used for the initial annotation
+of a newly sequence primate sequence, and many LINEs will extend to
+rodent, or even more distantn mammalian species.  This does not hold
+so well in plants.  In many cases TE composition of a given organism
+is species-specific.  For example, a rice-based TE libarary will only identify 25\%
+of the TEs in the maize genome. [CITE Jiang]
+
+To solve this problem we turn to {\it de novo} TE identification
+tools: tools identifying TEs though a genome using only the genome
+sequence information.  A number of such tls have been publised in the
+literature, operating on a number of different principles.  Tools such
+as RECON and PILER are based on self-alignment, using WU-BLAST and
+PALS for the alignemnt [CITE].  RECON show good sensitivity but is
+computationally intesive and infeasible for use on whole genomes (requiring 60 hours for 18Mb rice genome in a 2013
+report), while PILER achieves a good runtime with very low
+sensitivey. [CITE JIANG].   ReAS and RepeatScout are based on $k$-mer
+searches [CITE], with the earlier showing less sensitivity than
+RECON.  Finally RepeatGluer is based on a vairation od DeBrujin
+grauphs [CITE], which allows for a decomposition of TE families into
+domains, but is very, very computationally expensive.  In Saha {\it et
+  al.} the authors perform an extenive comparison of the tools, and
+conclude that RepeatScout is the best tool overall for assembled
+genomes and ReAS the best when dealing with unassembled sequeunce
+fragments [CITE].
+
+Another line of developement, first proposed by Zheng and Lonardi,
+involves the use of {\it elementary repeats} [CITE].  Similar in concept to
+the motifes identified by RepeatGluer [CITE], an elementary repeat is
+foramlly defined sequence pattern that could be considered a buiding
+block (or motif) of a transposable element.  Zheng and Lonardi defined
+the concept and proposed an idenficiation algorith, but that algorithm
+suffered from a runtime that was quadratic in the size of the genome.
+Two other works, He [CITE] and Huo {\it et al.} employed variations of
+suffix trees to get linear identification algorithms, but lost
+sensitivity in the resence of genomic substitions -- making them less
+useful on real data.
+
+Here we describe RAIDER (Rapid Ab Initio Detection of Elementary
+Repeats),
 

### 2015-07-22 manuscript/defs2.tex
Updating defs / manuscript

--- 2015-07-13 manuscript/defs2.tex
@@
 \newtheorem{lemma}{Lemma}
+\newtheorem{theorem}{Theorem}
 
@@
 \subsection{REFDs}
-In RAIDER, everything is defined in terms of $l$-mers.  hHence we need
-to generlize slightly to allow the ambiguity character, so we will
-create a description of a {\it repeat element family discriptior},
+In RAIDER, everything is defined in terms of $l$-mers.  Hence we need
+to generalize slightly to allow the ambiguity character, so we will
+create a description of a {\it repeat element family description},
 with an $l$-mer being a special case of that.
@@
 
+\begin{definition}
+Given aan refd and a sequence $G$, let $freq_G(r)$ be the number of
+positions $i$ at which $r$ generalizes $G_{i:i+|r|}$.
+\end{definition}
+
+{\bf Example:} Let $R=AA*AA$ and $G=AAAAACCAAGAAC$.  Then $freq_G(r)
+= 2$.
+
 \subsection{REFDs and Spaced Seeds}
@@
 \ldots$ (notice: 6 As), with the seed $1101$, we get the refd $AAAAAA*TT*TT$, which
-deposes into $AA*AA$, $AA*TT$, and $TT*T$$.  If any string compatible
+deposes into $AA*AA$, $AA*TT$, and $TT*T$.  If any string compatible
 with any of these three occurs anywhere else, it breaks the family.
@@
 
+\begin{definition}
+A decomposition $S$ of $r$ w.r.t. $s$ is \underline{maximal} it
+countains every $i$ such that $s$ hits $r$ at $i$.
+\end{definition}
+(Clearly this is unique.)
+
+Eample: 
+\begin{itemize}
+\item If $s=11011$ and $r=AAAAAA*CCCCCC$, then $D=\{0,1,4,7,8\}$.
+\item If $s$ has no 0s, then $D = \{i \; : \l 0 \leq i \leq
+  |r|-|s|\}$.
+\end{itemize}
+
+\begin{definition}
+Given an refd $r$, a length $l$ covering seed $s$, and a decomposition $D$, the
+\underline{sequence decomposion} $Q = \{\gamma(r_{i:i+l}) \; : \; i \in D}$ 
+\end{definition}
+
+{\bf Examples:}
+\begin{itemize}
+\item $s=11011$, $r=AAAAA*CCCCC$, the only minimal decomposition is
+  $D=\{0,4,6\}$, and the corresponding sequence decomposition is
+  $\{AA*AA, AA*CC, CC*CC\}$.
+\item $s=11011$, $r=AAAAATCCCCC$, there are five different minimal
+  decompositions.  For $\{0,1,6\}$ the sequence decomposition would be
+  $AA*AA, AA*AT$, and $TT*TT$.
+\end{itemize}
+
+\begin{definition}
+The {\it representitive sequence} of a decomposition is
+created by compressing the overlapping portions of the elements of the
+decompresion.
+\end{definition}
+
+{\bf Examples:}
+\begin{itemize}
+\item $s=11011$, $r=AAAAA*CCCCC$, $D={0, 3, 6}$,
+  $Q=\{AA*AA,AA*CC,CC*CC$}.  The representative sequence would be
+  $\{AA*AA*CC*CC\}$.
+\item $s=11011$, $r=AAAAATCCCCC$, $D={0,2,6}$, $Q=\{AA*AA, AA*TC, CC*CC\}$.  
+ The representative sequence would be $\{AAAAATCC*CC\}$.  We have to
+ compress the overlap at $r=2:5$, which endsup removing both *
+ symbols.
+\end{itemize}
+
+{\bf TO CONSIDER: What is the relationship between $r$ and the
+  representative sequence of one of its decompostion?  Always a
+  generlization of $r$?}
+
 \subsection{Elementary Repeats}
 
-I want to get a definition of elementary repeats that paralleles the
-one we had.  (And, hopefully, is provably equivilent for spaced seeds
-without spaces.)  This is tricky: I'm not yet sure what the equivilent
-is for the minimality condition (there is no subsequence of an
-elementary repeat that occurs independently of the instances of the
-elementary repeat).  Lets look why.
-
-Suppose we have in the string out genome $AAAAATCCCCC$ and
-$AAAAAGCCCCC$.  Using the seed $s=11011$, we should get the refd
-$r=AA*AA*CC*CC$. Here are different ways of phrasing the minimality
-condition, and why they break:
+\begin{definition}
+For a ferquenct $f$, a genome $G$ and seed $s$, an refd $r$ is an \undetline{elementary
+  repeat} if the following holds.  
 \begin{enumerate}
-\item No sub-refd can match any string in the genome outside of those
-  matching $r$.  Doesn't work: we can have an AAAAT occuring somewhere
-  else wihtout breaking this family, since the second $l$-mer is
-  irrelevant to construction.
-\item There cannot be any string $S'$, $|s'| < |r|$, occuring outside
-  the instances, such that $s'$ generalizes to a sub-refd $s$ or $r$.
-\end{enumerate}
-
-\begin{definition}
-Let $s$ be a fixed seed with length $l$, $f$ be a fixed integer, and $G$ be a fixed
-genomic sequence.  An
-\underline{Elementary Repeat Family} is a tuple $(r,S)$ where: 
-\begin{enumerate}
-\item $r$ is an refd such that $s$ covers $r$. 
-\item $S$ is a set of genomic coordinates such that $|S| \geq f$.  Say
-  that the sub-refds induced by $S$ are the substrings of $G$ starting
-  at cordinates of $S$ and extended $|r|$ bases.
-\item For every $i \in S$, $r$ is consistent with the string
-  $G_{i:i+|r|}$.
-\item Minimality condition.  Possibilities (I'm not sure which is
-  right):
-  \begin{itemize}
-  \item There is no sub-refd $r'$ of $r$ that is covered by $s$ and is
-     consistent some sequence $x$ that is sits outside the sequences
-      induced by $S$.
-  \item There is no $l$-mer $r'$ contained in $r$ that is consistence
-    with $s$, such that $\sigma_s(r)'$ occurs anywhere in $G$ outside
-    of the sequences induced by $S$.
- \end{itemize}
-\item There is no refd $r'$ containing $r$ such that every sequence
-  induced by $S$ is contained by some sequence consisted with $r'$.
+\item $s$ covers $r$.
+\item $freq_G(r) \geq f$.
+\item For the maximimal sequence decomposition $Q$, 
+  $freq_G(q) = freq_G(r)$ for all $q \in Q$.
+\item There does not exist an $r'$ that is covered by $s$ such that
+  (1) $r$ is compatable(2) with a sub-refd of $r'$, and (2) $r'$ meets
+  conditions (1)-(3) of this definition.
 \end{enumerate}
@@
 
-{\bf Comments:} For the last two points, I'm really not sure if the
-substring relation is the correct relation between $r$ and $r'$.
+\begin{theorm}
+If $s$ is a contiguous spaced seed (that is, no 0s) and $r$ is a
+repeat element w.r.t to $s$, the $r$ is a repeat element by Definition
+5 of Nate's thesis.
+\end{theorem}
+{\bf Proof:} Let $s$ be a spaces seed of length $L$ with no 0 symbols,
+and $r$ be an elementary repeat.  Note that $r$ has no $*$
+symbols, as otherwise $s$ could not cover it.  Now consider the
+conditions for Definition 5:
+\begin{itemize}
+\item The repeat element is at least length $L$: since $r$ is covered
+  by seed $s$ of length $L$, this is clearly true.\footnote{There
+    appears to be a mistake in Nate's thesis with condition 1 -- this
+    is what it was supposed to mean.}
+\item $freq_G(r) \geq f$: Follows directly from (2).
+\item Every substring of $r$ has the same frequence as $r$: Since $s$
+  has no 0s, $Q$ is the set of all substrings of $r$. Hence this
+  condition applies to every substring of $r$, which is what we nee.
+\item $r$ is maximal w.r.t to Nate's defintion: Suppose this were not
+  true: we could add 1 character to $r$  to form $r'$, and have an element that
+  meats conditions Nate's definition.  (Note that, with no spaces, if
+  we can add one or more characters we can add only 1 character -- we
+  can assume are only adding 1.)  That means that every subsequence of
+  $r'$ has the same frequency in $G$ (by condition (3) of Nate's
+  defintion), meaning that they fufill condition (3) of our
+  definition.  The other two conditions of our definition our
+  immediate, meaning $r'$ contradicts (4) and hence $r$ is not an elementary
+  repeat (by our definition).
+\end{itemize}
 
 \subsection{Lemmas}
-
-Right now, I'm trying trying to think ofpossibly interesting
-statements and prove them.
-
 Assume a fixed seed $s$ of length $l$ and a fixed genome $G$.
 
+\begin{definition}
+An $l$-mer {\it belongs} to a repeat element $r$ if the element is
+compatable with an element in the maximal sequence decomposition of $r$.
+
+\begin{lemma}
+An $l$-mer can belong to at most one elementary repeat family.
+\end{lemma}
+Using our definition of ``belongs'', violating this would vilate
+condition (3) of the r.e. definition.
+
+\begin{observation}
+For any string $l$-mer $w$, if $freq(\gamma_s(w)) \geq w$ than
+$\gamma_s(w)$ is either is an elementary
+repeat or is compatable with some substring of an elementary repeat.
+\end{observation}
+$\gamma_s(w)$ trivially satsfies the first three conditions, so only
+the fourth condition could break it.
+
+\begin{lemma}
+Let $s_1$ and $s_2$ be two sequences in $G$ such that refd generalizes
+both but there is no refd that generalizes generlizes superstrings
+properly containing each, then $r$ is not compatable with a sub-refd
+of any repeat element family.
+\end{lemma}
+Violation of this would imply a violation of (2) in the definition.
+
+\begin{lemma}
+If two $l$-mers belong to the same elementary repeat, then they have
+the same frequency.
+\end{lemma}
+Given our definition of ``belongs'', this follow immediately.
+
 \end{document}
+
+%%  LocalWords:  REFDs mers

### 2015-07-22 manuscript/defs3.tex
Updating defs / manuscript

@@
 
-\subsection{REFDs}
+\section{REFDs}
 
@@
 
-We will allow any normal string operator to apply to sequence descriptors.  In specific, for a s.d. $r$:
+We will allow any normal string operator to apply to sequence descriptors.  In specific, for an s.d. $r$:
 \begin{itemize}
-\item $r_{i,i+l}$, or $r[i:i+l]$, is the length $l$ sub-sd (substring) of $r$ starting at $i$.
+\item $r_{i,i+l}$, or $r[i:i+l]$, is the length $l$ sub-sd (substring) of $r$ starting at $i$.  (Python-like notation.)
 \item $|r|$ is the number of characters in $r$.
@@
 \item $r_1$ and $r_2$ are \und{consistent} if, for any position $i$ in
-  which they differ, one of them is a $*$. 
+  which they differ, one of them has a $*$ in that position.
 \item $r_2$ \und{generalizes} $r_1$ (and $r_1$ is \und{generlized} by $r_2$),
@@
 \begin{definition}
-  The frequency of an s $r$ in genome sequence $G$ is the number of
+  The frequency of an sd $r$ in genome sequence $G$ is the number of
   $i$ such that $G_{i:i+|r|}$ is compatible with $r$.  
 \end{definition}
+{\bf Note:} Debated whether this should be ``compatible'' or ``generalized'' -- then realized that since $G$ is a string (as opposed to an sd) they are equivilent.
 
@@
 
-\subsection{REFDs and Spaced Seeds}
+\section{REFDs and Spaced Seeds}
 
@@
   \begin{itemize}
-  \item We say $s$ and an $l$-mer $r$ are \und{consistent} if, when $s_i = 0$, $r_i = *$.
+  \item We say $s$ and an $l$-mer $r$ are \und{consistent} if, when $r_i = *$, $s_i=0$.
   \item We say seed $s$ \und{hits} sd $r$ at position $i$ such that
@@
 $\{0,3,6\}$.  $s$ does not cover $AAAAA**TTTTT$, as there is no
-position at which $s$ hits and extens over $i=5$.
+position at which $s$ hits and extends over $i=5$.
 
@@
 \begin{notation}
+Notation for the concepts from this section:
   \begin{itemize}
@@
 
-\subsection{Elementary repeats}
+\section{Elementary repeats}
 
@@
 \begin{definition}
-  A {\it repeat element family} is a tuple $(r,S)$ where $r$ is a
-  repeat element descriptor and $S$ is the set of coordinates of $G$
-  where there is a sequence compatiable with $r$.
+  A {\it repeat element family} is a set $S$ of genome coordinates
+  such that there exists an sd $r$ where:
+\begin{itemize}
+\item $r$ generlizes the string $G_{i:i+|r|}$ for every $i \inS4.
+\item $r$ does not generalize any other substring of $G$.
 \end{definition}
+Notice that there can be multiple such $r$ for a given set.  Any two
+such $r$ for the same $S$ will be compatible, but they don't
+necessarily generlize each other.
 
-\subsection{Equivilance proof}
+\section{Equivilance proof}
 
@@
 Given a genomic sequence $G$ and integers $f$ and $l$, a seqeunce $r$
-is an {\it elementary repeat} if we can write $r = x_0 \cdot x_2 \cdot
-\cdot x_k$ $where $|x_i|=l$, $k = |r|-|l|$, and:
+is an {\it elementary repeat} if we can write $r = x_0 \circ x_2 \circ
+\cdots \circ x_k$ where $|x_i|=l$, $k = |r|-|l|$, and:
 \begin{enumerate}
@@
 \item $freq_G(x_i) = freq_G(r)$
-\item There does not exist an $y$, $|y|=l$ such that either $y \cdot
-  r$ or $r \cdot y$ meets conditions 1-3.
-\end{enuemrate}
+\item There does not exist an $y$, $|y|=l$ such that either $y \circ
+  r$ or $r \circ y$ meets conditions 1-3.
+\end{enumerate}
 \end{definition}
@@
   has no zero chartacters.  $r$ is an elementary
-  repeat descriptor for $s$ (by Definition~\label{ERdef}) if and only if $r$
+  repeat descriptor for $s$ (by Definition~\ref{ERdef}) if and only if $r$
   is an elementary repeat for $l=|s|$ (by
-  Definition~\label{NateDef})).
+  Definition~\ref{NateDef})).
 \end{theorem}
@@
 
-($Leftarrow$) First we observe that if $s$ has no zeros, and covers
+($\Leftarrow$) First we observe that if $s$ has no zeros, and covers
 $r$, then $r$ has no * characters.  From this it follows that
-$\Gamma_s(r) = \{r_0,r_1,\cdtos,r_{k}$ where $|r_i|=l$ and $k=r-l$ --
-every $l$-mer contained with $r$.  From this it follows that $r = r_0 \circ
-r_1 \
+$\Gamma_s(r) = \{r_0,r_1,\cdots,r_{k}\}$ where $|r_i|=l$ and $k=r-l$.
+That is, every $l$-mer contained with $r$.  From this it follows that $r = r_0 \circ
+r_1 \circ \cdots \circ r_k$.
 
@@
   \item Since $r$ contains no * symbols, the only string it
-    is compatible with is itself.  Meaning it has the sme frequency by
+    is compatible with is itself.  Meaning it has the same frequency by
     either definition (sd or string).  Hence (2) of Nate's
-    definition following directly from (2) of out definition.
+    definition follows directly from (2) of our definition.
   \item Follows immediately from the fact that every $r_i$is a
@@
    \item Argue by contradiction: suppose there were such a $y$ (assume
-     w.l.g. it follow $r$, and let $r' = r \circ x$.  Since Nate's
+     w.l.g. it follow $r$), and let $r' = r \circ x$.  Since Nate's
      definition tells us $r'$ matches a string around every occurance
      of a match to $r$, this means $r'$ is (trivially) compatable to a
-     substring around every occurance of a match to $r'$ -- to $r'$
+     substring around every occurance of a match to $r'$ -- so $r'$
      could not have been a repeat element descriptor.
@@
 
-($Rightarrow$) This is the same argument backwards, and not worth
+($\Rightarrow$) This is the same argument backwards, and not worth
 writing out.
-\end{enumerate}
 
-\subsection{Basic Lemmas}
+\section{Basic Lemmas}
 
@@
 {\bf COMMENT:} HMM...I'm worried about this one.  What if we extneded
-by a whole sead length?  Could that matter?
+by a whole se1d length?  Could that matter?
 
 \begin{lemma}
-  If two $l$-mers appear blog to the same elementary repeat
-  descriptor, then they have the same frequence.
+  If two $l$-mers belong to the same elementary repeat
+  descriptor, then they have the same frequency.
 
@@
 
-\subsection{Variations}
+\section{Variations}
 
@@
 For $11011$ and $G = \cdots ACCCCCGTTTTTA\cdots ACCCCCATTTTTA$, we can
-have $AC*CCC*TT*TTA$ or $ACC*CC*TTT*TA$$, among others.
+have $AC*CCC*TT*TTA$ or $ACC*CC*TTT*TA$, among others.
 
-What do we know about the r.e.d. that Carly's algorithm will pick?
+\subsection{Carly's output}
+
+\begin{definition}
+Given a repeat element family $S$ and a descriptor $r$, we say:
+\begin{itemize}
+\item We say position $i$ is variable in $S$ if there are two elements
+  of $S$ whose sequences vary at positin $i$.
+\item Position $i$ of $r$ is \und{tight} if either $r_i \neq *$ or
+  $S$ variable at position $i$.  (That is, $r$ must have a *.)
+\item We say $r$ is tight if $r_i$ is tight for all $0 \leq i < |r|$.
+\end{itemize}
+\end{definition}
+
+Facts about Carly's output:
+\begin{itemize}
+\item If $S$ is invarient in all positions, then the results $r$ has
+  no * characters.  For every $i$, $s$ hits at either $i$ or
+  $i-|s|+1$, meaning that there is a $w \in \Gamme_s(r)$ that starts
+  (or ends) at position $i$.  Since a seed starts (ends) with a 1,
+  there canot be a * at this position.
+\item The output descriptor is not necessarily tight.  For example:
+  $s=11011$, where $G$ contains $AAAAACGGGGG$ and $AAAAATGGGGG$.  Then
+  $r = AA*AA*AA*AA$, hence is not tight at position 2.
+\end{itemize}
 

### 2015-07-24 RAIDER_eval.py
modifications to shell scripts. small change to REVAL allowing for PRA output in summary file

@@ -78,7 +78,6 @@ RedhawkLocations = {'build_lmer_table':'./build_lmer_table',
                     'araider':'./araider',
                     'raider2': './raiderv2_options'}#,'raider2_old':'./raiderv2_old','raider2_oldest':'./raiderv2_oldest'}
 Locations = None;    # This will be set to one of the above two, and references to find exectuable locations.
-raider2_ages = [0,1]
 
 
 #########
@@ -105,7 +104,7 @@ def file_dir(file):
 def parse_redhawk_time(time_str):
     """Parse time limit string for redhawk (format HH:MM:SS) into seconds amount"""
     secs = sum(int(x) * 60 ** i for i,x in enumerate(reversed(time_str.split(":"))))
-    print(time_str, '/t', secs)
+    #print(time_str, '/t', secs)
     return secs
 
 def convert_seed(seed):
@@ -177,12 +176,15 @@ def parse_params(args):
     raider2_argument = parser.add_argument_group("RAIDER2 parameters")
     raider2_argument.add_argument('--age', type = int, help="Use older version of raider2", default=1) 
     raider2_argument.add_argument('--aa', '--all_ages', dest="all_ages", action="store_true", help="Run all ages of raider2", default=False) # type = int, help="Use older version of raider", default=0)
+    raider2_argument.add_argument('--multi', '--multi_seed', dest="multi_seed", action="store_true", help="Run all seeds in seed file concurrently",default=False)
 
     # REPSCOUT ARGUMENTS
     repscout_argument = parser.add_argument_group("REPSCOUT parameters")
     raider_argument.add_argument('--repscout_min', type = int, help = "Minimum repeat length for repscout.", default = 10)
-    raider_argument.add_argument('--uff', '--use_first_filter', dest = "use_first_filter", action = "store_true", help = "Use the first RepScout filter", default = False)
-    raider_argument.add_argument('--usf', '--use_second_filter', dest = "use_second_filter", action = "store_true", help = "Use the second RepScout filter", default = False)
+    raider_argument.add_argument('--rs_min_freq', type = int, help = "Minimum repeat length for repscout.", default = 3)
[…]

### 2015-07-24 aseed_files/programs/create_all.py

@@ -1,0 +1,24 @@
+#!/software/python/3.3.3/bin/python3.3
+
+import itertools
+import argparse
+
+
+def generate_seed_file(seed_file, length, weight): 
+    f = open(seed_file, 'w')
+    for bits in itertools.combinations(range(length - 2), weight - 2):
+        s = ['0'] * length
+        s[0] = '1'
+        s[length - 1] = '1'
+        for bit in bits:
+            s[bit + 1] = '1'
+        f.write(''.join(s) + '\n')
+
+
+if __name__ == "__main__":
+    parser = argparse.ArgumentParser(description = "Generate All Seeds With Specified Length and Weight.")
+    parser.add_argument('-l', '--length', type = int, help = "Seed length", default = 23)
+    parser.add_argument('-w', '--weight', type = int, help = "Seed weight", default = 13)
+    parser.add_argument("seed_file", help = "Seed output file")
+    args = parser.parse_args()
+    generate_seed_file(args.seed_file, args.length, args.weight)

### 2015-07-24 aseed_files/programs/create_all_palls.sh

@@ -1,0 +1,26 @@
+#!/bin/bash
+
+
+#pass in minw maxw minl maxl
+
+aseed_dir='../aseeds/'
+dir=${aseed_dir}L_${minl}_${maxl}_W_${minw}_${maxw}
+currl=${minl}
+currw=${minw}
+
+mkdir ${dir}
+
+echo "min weight= ${minw} max weight= ${maxw}"
+echo "min length= ${minl} max length= ${maxl}"
+
+while [[ $currl -le $maxl ]]; do
+    currw=${minw}
+    echo "currl= ${currl}"
+    while [[ $currw -le $maxw && $currw -le $currl ]]; do
+        echo "currw= ${currw}"
+        python3 create_all_pals.py -l ${currl} -w ${currw} ${dir}/L${currl}_W${currw}
+        let currw=currw+1
+    done
+    let currl=currl+1
+done
+

### 2015-07-24 aseed_files/programs/create_all_pals.py

@@ -1,0 +1,25 @@
+#!/software/python/3.3.3/bin/python3.3
+
+import itertools
+import argparse
+
+
+def generate_seed_file(seed_file, length, weight): 
+    f = open(seed_file, 'w')
+    for bits in itertools.combinations(range((length - 2)//2), (weight - 2)//2):
+        s = ['0'] * length
+        s[0] = '1'
+        s[length - 1] = '1'
+        for bit in bits:
+            s[bit + 1] = '1'
+            s[length - bit - 2] = '1'
+        f.write(''.join(s) + '\n')
+
+
+if __name__ == "__main__":
+    parser = argparse.ArgumentParser(description = "Generate All Palindromic Seeds With Specified Length and Weight.")
+    parser.add_argument('-l', '--length', type = int, help = "Seed length", default = 23)
+    parser.add_argument('-w', '--weight', type = int, help = "Seed weight", default = 13)
+    parser.add_argument("seed_file", help = "Seed output file")
+    args = parser.parse_args()
+    generate_seed_file(args.seed_file, args.length, args.weight)

### 2015-07-24 aseed_files/programs/create_pals.sh

@@ -1,0 +1,15 @@
+#!/bin/bash
+
+#rep is the string to be repeated
+str=${rep}
+until [ "${#str}" -ge "${min}" ]; do
+    str=${str}${rep}
+done
+
+until [ ${#str} -ge ${max} ]; do
+    strcur=${str}1
+    echo "${strcur}" >> $fname
+    str=${str}${rep}
+done
+
+

### 2015-07-24 cleanup.sh

@@ -1,7 +1,7 @@
 #!/bin/bash
 rm -r $dir/RAIDER/
 rm -r $dir/ARAIDER/
-rm -r $dir/RAIDERV2/
-rm -r $dir/RPT_SCT/
+rm -r $dir/RAIDERV2*/
+rm -r $dir/RPT_SCT*/
 rm -r $dir/SOURCE_DATA/
 #rm *.batch.*

### 2015-07-24 sorted_analyze.R

@@ -1,7 +1,6 @@
 #!/software/R/3.0.2-gcc/bin/Rscript
 
 suppressPackageStartupMessages(require(optparse))
-
 suppressPackageStartupMessages(require(gdata))
 getDf <- function(fname) {
     #dat <- readLines(fname,n=-1)
@@ -35,9 +34,10 @@ getMatches <- function(expr, seed){
 
 numberOf <- function(seed, val){
     match.string <- paste("(?<=", val, "\\^\\{)[[:digit:]]+(?=\\})", sep="")
-    numOnes <- sumOverMatches(getMatches(match.string, seed))
+    numOnes <- sumOverMatches(getMatches(match.string, as.character(seed)))
     match.string <- paste(val, "(?!\\^)", sep="")
     numOnes + length(unlist(getMatches(match.string, seed)))
+    
 }
 
 toSeedString <- function(seed){
@@ -46,12 +46,20 @@ toSeedString <- function(seed){
     match.indices <- gregexpr(match.string,seed,perl=TRUE)[[1]]
     match.values <- regmatches(seed,gregexpr(match.string,seed,perl=TRUE))[[1]]
     j = 1
+    #print("here")
+    #print(match.indices)
+    if (length(match.indices) == 1 && match.indices[1]==-1){
+        #print("seed")
+        seed
+        #return as.string(seed)
+    }
+    else{
     for( i in 1:length(match.indices)){
         next.index <- match.indices[i]
         next.val <- match.values[i]
         seed.val.index = next.index - 3
         if(seed.val.index > j + 1){
-            print( substring(seed, j, seed.val.index - 1))
+            #print( substring(seed, j, seed.val.index - 1))
             seed.string <- paste(seed.string, substring(seed, j, seed.val.index - 1), sep="")
         }
         j = next.index + length(as.character(next.val)) + 1
@@ -59,7 +67,7 @@ toSeedString <- function(seed){
         seed.string <- paste(seed.string, paste(replicate(next.val,as.character(seed.val)),collapse=""), sep="")
     }
     seed.string
-    
+    }
 
 }
 
@@ -67,11 +75,11 @@ toSeedString <- function(seed){
 addInfo <- function(fname, include.seeds, include.type, include.chrom, include.dir){
     dat <- getDf(fname) #read.table(fname,header=TRUE,comment.char="")
     if(include.type) {
-        type <- basename(dirname(dirname(dirname(fname))))
[…]

### 2015-07-24 manuscript/BMC/bmc_article.tex

@@
 
+\newtheorem{definition}{Definition}
+\newtheorem{observation}{Observation}
+\newtheorem{lemma}{Lemma}
+
 %%%%%%%%%%%%%%%%%%%%%%%%% start of article main body
-\section({Introduction}
-Here we introduce RAIDER (Rapid Ab Initio Detection of Elementary
+\section*{Background}
+Here we introduce phRAIDER (Rapid Ab Initio Detection of Elementary
 Repeat), designed to enable the quick annoation of transposible
@@
 
+\subsection*{Elementary Repeats}
 Another line of developement, first proposed by Zheng and Lonardi,
-involves the use of {\it elementary repeats} [CITE].  Similar in concept to
-the motifes identified by RepeatGluer [CITE], an elementary repeat is
-foramlly defined sequence pattern that could be considered a buiding
-block (or motif) of a transposable element.  Zheng and Lonardi defined
-the concept and proposed an idenficiation algorith, but that algorithm
-suffered from a runtime that was quadratic in the size of the genome.
-Two other works, He [CITE] and Huo {\it et al.} employed variations of
-suffix trees to get linear identification algorithms, but lost
-sensitivity in the resence of genomic substitions -- making them less
-useful on real data.
+involves the use of {\it elementary repeats} [CITE].  
+Zheng and Lonardi formulated their defintion with the explicit intention of factoring minimum length and
+frequency into the definition: to be an elementary repeat a sequence
+must have some minimum length $l$ and occur at least $f$ times.  More
+importantly, it must be minimal (that is, no section of an elementary
+repeat can itself be an elementary repeat), and it must have appear
+independently of any other sequence meeting definition of an
+elementary repeat.  [CITE]  Formally:
+\begin{definition}
+Give a length criteria $l$ and a frequence criteria $f$, a sequence
+$r$ is an elementary repeat in a genome if:
+\begin{enumerate}
+\label{ZLDef}
+\item $r$ is of length at least $l$.
+\item There are at least $f$ copies of $r$ in the genome.
+\item There is no proper subsequence $r'$ of $r$ such that $r'$ is of
+  length $l$ or grater and $r'$ occurs independently of $r$.  (That
+  is, every length $>l$ subsequence of $r'$ occurs only where $r'$
+  occurs in the genome.  (The minimality condition.)
+\item There is no sequence $r'$ that properly contains $r$ and appears
+  around every occurence of $r$ in the genome.
+\end{enumerate}
+\end{definition}
+The third condition insures minimality (that an elementary repeat
+cannot contain elementary repeats), while the fourth condition that
+each elementary repeat is maximal in length.  Having proposed this
+definition, Zheng and Lonardi developed an identification algorithm
+that had a runtime quadratic in the genome size -- so not of practical
+use on whole-genome analysis.  This was refined to linear time by He [CITE],
+and also by Huo {\it et al.} [CITE] based on variations suffix tree
+approaches, but because of that use of suffix trees they are limited
+in their ability to handle sequence variation.  As transposible
+elements naturallly accumulate instance-specific base substitutions
+over time, this is a significant limitation.
 
-Here we describe RAIDER (Rapid Ab Initio Detection of Elementary
-Repeats),
+\subsection*{RAIDER}
+It was with the objective of creating a linear time identification
+algorithm that could handle variation through use of
+PatternHunter-like spaced seeds [CITE].  A rough implementatin was
+first presentd in Figuera {\it et al.}, with more details in the
+Figuera masters thesis [CITE].  Not itself fully able to handle the
+spaced seed component, RAIDER was built along an alternate,
+but equivilent definition of elementary repeats based on $l$-mers
+(sequences of length $l$).  Specifically, it was observed that
+condition (3) of Definition~\ref{ZLDef} could be rewritten as: There
+is no $l$-mer contained with $r$ that appears in the genome
+independelty of $r$.
+
+From this it follows that:
+\begin{enumerate}
+\item An $l$-mer cannot belong to two different elementary repeats.
+\item Any $l$-mer in the genome that occurs $f$ or more times is
+  either an elementary repeat or is contained on one.
+\item Any two $l$-ers contained within an elementary repeat must occur
+  the same number of times in the genome.
+\item Given two sequences in the genome that are {\it maximally
+  idenitcal} (that is, cannot be extended in either direction and stil
+  be identical), these sequences cannot be contained with in
+  elementart repeat.
+\end{enumerate}
+(For dicussion and proof, see the Figueroa Thesis [CITE]].)
+
+Based on these observations, we discover we can find all elementary
+repeats in a single scan of the genome.  Specifically, as we scan from
+left to right, we track $l$-mer occurances, and look for repeated
+$l$-mer sequences.  When we find the same sequence of $l$-mers
+occuring multile times in a row, we can mark it as a tentative family,
+then brack it down later if we discover violations of the minimality
+condition.  The alogrithm is summerized in Figueroa {\it et al.} [CITE], and
+discussed in detail in the Figueroa Thesis [CITE].
+
+Results of the preliminary implementation were promising.  Based on
+the Saha {\it et al} analysis putting RepeatMasker as the top {\it de
+  novo} search tool, comparisone were aginst that.  RAIDER result
+quality was close, but slightly behind -- with runtime
+order of magnitudes better.  On human chromosome 22 we saw a
+$12\times$ speedup (2344 seconds to 192 seconds), while coverage
+actually improved (77\% to 84\%), while on mouse chromsome 19 we saw 
+the same speedup with a sinifcant drop in coverage (53\% to 30\%).  On
+the full human genome RAIDER ran in $6.3$ hours, while RepeatScout was
+unable to complete it run.  For details, see Figueroa {\it et al.} [CITE]
+
+\subsection*{Spaced Seeds}
+
+PatternHunter, first proposed as a very succesful augmentation to
+BLAST [CITE PatternHunter,BLAST], is based on the notion of {\it
+  spaced seeds}.  In the context of BLAST, instead of looking for a
+cntiguous substring match to triger the alignment exention,
+PatternHunter uses a more refised match pattern.  For example, the
+{\it seed} 1111110111111 would specify that we were looking for an exact
+match of two length-six substrings seperated by a single base (which
+might or might not match).  In the toy seed $s=110101$, we would say the
+strings $AACACA$ and $AAGACA$ because all positions corresponding to a
+1 match.  The string $AACACA$ does not math $TACACA$ (with respect to
+$s$), as the single miss-match does not correspond to the 1.  Nor does
+$AACACA$ match ant substring of $TTAACTCA$.
+
+The incorperations of spaced seeds into BLAST lead to significant
+improvements at vertual no cost in runtime [CITE], so it made sence to
+incorportat them here.  Specifcially, instead of looking at repeat
+elements as a set of identical strings, we could require that they
+match with respect to a PatternHunter spaced seed.  The preliminary
+version of RAIDER from Figueroa {\it et al.} made some attempt to
+incorperate this -- was succesful enough to provide a
+proof-of-concept, but a significant amount of work was left to be done
+for the developement of phRAIDER (PatternHunter-based RAIDER) -- the
+main method in this paper.
+
+\section{Methods}
 

### 2015-07-27 manuscript/BMC/bmc_article.tex
updaing paper; adding bmc.bib

--- 2015-07-21 manuscript/BMC/bmc_article.tex
@@
 
-\title{Fast de novo transpoible element annotation}
+\title{Fast de novo transposable element annotation}
 
@@
 \author[
+   addressref={aff1},
+   noteref={n1},
+   email={schaefce@miamiOH.edu}
+]{\inits{SE}\fnm{Carly E} \snm{Schaeffer}}
+\author[
    addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
@@
 ]{\inits{ND}\fnm{Nathaniel D} \snm{Figueroa}}
-\author[
-   addressref={aff1},
-   email={schaefce@miamiOH.edu}
-]{\inits{SE}\fnm{Carly E} \snm{Schaeffer}}
 \author[
@@
 \begin{abstract} % abstract
-\parttitle{First part title} %if any
-Text for this section.
+\parttitle{Background} The problem of {\it de novo} identification of
+transposable elements (the discovery and annotation of transposable
+elements without the use of a pre-complied library describing sequence
+or structural features of the family) has been addressed by a number
+of tools -- all of which are either based on computationally complex
+algorithms that cannot scale to whole-genome use, or whose sensitivity
+suffers significantly from the presence of the sequence variation
+present in all sets.  Here we present phRAIDER (Pattern Hunter based
+Rapid Ab Initio Detection of Elementary Repeats) for quickly masking
+a newly sequence, or poorly annotated, genome of transposable elements
+that is robust to the sequence variation present in real data.
+
+\parttitle{Results}
+
+\parttitle{Conclusions}
+
+Here we introduce phRAIDER
 
@@
 \begin{keyword}
-\kwd{sample}
-\kwd{article}
-\kwd{author}
+\kwd{transposable elements}
+\kwd{elementary repeats}
+\kwd{pattern hunter}
+\kwd{genomic masking}
 \end{keyword}
@@
 
+\newtheorem{definition}{Definition}
+\newtheorem{observation}{Observation}
+\newtheorem{lemma}{Lemma}
+\newtheorem{theorem}{Theorem}
+
 %%%%%%%%%%%%%%%%%%%%%%%%% start of article main body
-\section({Introduction}
-Here we introduce RAIDER (Rapid Ab Initio Detection of Elementary
-Repeat), designed to enable the quick annoation of transposible
-elements and repetitive DNA without requiring a pre-compiled template
-library.  Transposible elements, or TEs, are present in almost every
-higher order genome, covering as much as 45\% of the human genome and
-90\% of the maize genome [CHECK THIS] [CITE].  Given their prevelence,
-identiication of TEs is important not only for those who study them, but for
-anyone trying to understand a geome.  If not filtered, TEs can trigger
-hug numbers of false positives in the course of applying automated
-gene finding techniques, as well as blow up runtime.  [CITE: Jiang]
-So identifying and masking them in a must for any genome being
-studied.
+\section*{Background}
+Transposable elements, or TEs, are genomic sequences that have at some
+point had the capacity to insert copies of themselves into other
+genomic locations, resulting in homologous families of sequences
+spread across the genome.  TEs are present in almost every higher
+order genome, covering as much as 45\% of the human genome and 90\% of
+the maize genome [CHECK THIS] \cite{Venter:2001p92} [CITE MAIZE].  TEs have
+proved an importent sorce of data in a number of studies \cite{Arndt:2005p279,Karro:2008p123,Mugal:2009p581,Hardison:2003p97}, but
+given their prevalence, it is just important for those not interested
+in them to have them masked out -- their bases replaced by N or case
+changed for easy identification and filtering.  For example. if not
+filtered, TEs can trigger huge numbers of false positives automated
+gene finding tool \cite{Jiang:2013jt}, as well as inflate tool runtime.
+Hence the 09
 
-The most well known tools for repeat identification ar RepeatMasker
-and [Eddy's tool] [CITE], both of which work on the principle of a
-library-based search.  That is, to final transposable elements of a
-given family,
+The most well known tools for repeat identification are RepeatMasker
+and nHMMER \cite{RepeatMaskerOpen:XkNxXSd7,Wheeler:2013gj}, both of
+which work on the principle of a library-based search.  That is, to
+identify transposable elements to a known family, the tools require
+some description of sequences in the family (e.g. a ancestral sequence
+for BLASTing, or a profile HMM description) that is used as the basis
+for identification.  But, much like we ask how the snow plow driver
+gets to work, or how dictionary authors look up the spelling of a word
+\cite{Pratchett:uw}, we must ask how the libraries on which the tools are based are
+initially compiled?  How are families initially discovered, and what
+do we do with a newly sequenced genome?
+
+Within mammalian species we can, to some extent, rely on homology
+relationships to port libraries across species.  A library description
+of a human ALU sequence can likely be used for the initial annotation
+of a newly sequenced primate genome, and many LINEs will extend to
+rodent, or even more distant, mammalian species.  This does not hold
+so well in plants: in many cases TE composition of a given plant organism
+is species-specific.  For example, a rice-based TE library will only identify 25\%
+of the TEs in the maize genome \cite{Jiang:2013jt}.
+
+To solve this problem we turn to {\it de novo} TE identification
+tools: tools identifying TEs though a genome using only the genome
+sequence information.  A number of such tools are discussed in the
+literature, operating on a number of different principles.  Tools such
+as RECON \cite{Bao:2002} and PILER \cite{Edgar:2005p2365} are based on
+self-alignment, using WU-BLAST and LASTZ for the alignment
+\cite{Lopez:2003td,Harris:2007uf}.  RECON show good sensitivity but is
+computationally intensive and infeasible for use on whole genomes
+(requiring 60 hours for 18Mb rice genome in a 2013 report), while
+PILER achieves a good runtime with very low sensitivity
+\cite{Jiang:2013jt}.  ReAS \cite{Li:2005p3138} and RepeatScout
+\cite{Price:2005p1247} are based on $k$-mer searches, with the
+earlier showing less sensitivity than RECON \cite{Jiang:2013jt}.
+Finally RepeatGluer \cite{Pevzner:2004p3157,Zhi:2006p3199} is
+based on a variation of DeBrujin graphs, which allows for a
+decomposition of TE families into domains, but is very, very
+computationally expensive.  In Saha {\it et al.} the authors perform
+an extensive comparison of the tools, and conclude that RepeatScout is
+the best tool overall for assembled genomes, while ReAS the best when
+dealing with unassembled sequence fragments \cite{Saha:2008dm}.
+
+\subsection*{Elementary Repeats}
+Another line of development, first proposed by Zheng and Lonardi
+\cite{Zheng:2005bl}, involves the use of {\it elementary repeats}.
+Similar to the RepeatGluer domains, elementary repeats are
+decompositions of TEs into basic building blocks.  Identification of
+these building blocks are be sufficient of their own for the purposed
+of masking, and can be assembled into Transposible Elements for those
+interested in studying those elements.
+
+Zheng and Lonardi formulated their definition with the explicit
+intention of factoring minimum length and 
+frequency into the definition: to be an elementary repeat a sequence
+must have some minimum length $l$ and occur at least $f$ times.  More
+importantly, it must be minimal (that is, no section of an elementary
+repeat can itself be an elementary repeat), and it must have appear
+independently of any other sequence meeting definition of an
+elementary repeat  Formally:
+\begin{definition}
+Give a length criteria $l$ and a frequency criteria $f$, a sequence
+$r$ is an elementary repeat in a genome if:
+\begin{enumerate}
+\label{ZLDef}
+\item $r$ is of length at least $l$.  (The length condition.)
+\item There are at least $f$ copies of $r$ in the genome.  (The
+  frequency condition.)
+\item There is no proper subsequence $r'$ of $r$ such that $r'$ is of
+  length $l$ or greater and $r'$ occurs independently of $r$.  That
+  is, every length $>l$ subsequence of $r'$ occurs only where $r'$
+  occurs in the genome.  (The minimality condition.)
+\item There is no sequence $r'$ that properly contains $r$ and appears
+  around every occurrence of $r$ in the genome.  (The maximality condition.)
+\end{enumerate}
+\end{definition}
+Having proposed this definition, Zheng and Lonardi developed an
+identification algorithm that had a runtime quadratic in the genome
+size -- so not of practical use on whole-genome analysis.  This was
+refined to linear time by He and also by Huo {\it et al.}
+\cite{He:2006gpa,Huo:2009hoa} based on variations of suffix tree
+approaches, but because of that use of suffix trees they are limited
+in their ability to handle sequence variation.  As transposable
+elements naturally suffer from copy mistakes and accumulate
+instance-specific base substitutions over time, this is a significant
+limitation.
+
+\subsection*{RAIDER}
+\label{RAIDERSec}
+It was with the objective of creating a linear time identification
+algorithm that could handle variation through use of
+PatternHunter-like spaced seeds that we developed the prototype RIADER
+\cite{Li:2004wl}.  A rough implementation was first presented in
+Figueroa {\it et al.}, with more details in the Figueroa masters
+thesis \cite{Figueroa:2014uk,Figueroa:2013cz}.  Not itself fully able
+to handle the spaced seed component, RAIDER was built along an
+alternate, but equivalent definition of elementary repeats based on
+$l$-mers (sequences of length $l$).  Specifically, it was observed
+that condition (3) of Definition~\ref{ZLDef} could be rewritten as:
+There is no $l$-mer contained with $r$ that appears in the genome
+independently of $r$.
+
+\begin{observation}
+  \label{RAIDERobs}
+  From the definition, we can now observe:
+\begin{enumerate}
+\item An $l$-mer cannot belong to two different elementary repeats.
+\item Any $l$-mer in the genome that occurs $f$ or more times is
+  either an elementary repeat or belong to one.
+\item Any two $l$-ers belonging to an elementary repeat must appear
+  the same number of times in the genome. 
+\item Given two sequences in the genome that are {\it maximally
+  identical} (that is, cannot be extended in either direction and still
+  be identical), these sequences cannot belong to a larger
+  elementary repeat.
+\end{enumerate}
+\end{observation}
+By ``belong'' we mean ``is a substring of'' -- the reason for the
+terminology will be explained shortly.  For discussion and proof, see
+the Figueroa Thesis \cite{Figueroa:2013cz}.
+
+Based on these observations, we discover we can find all elementary
+repeats in a single scan of the genome.  Specifically, as we scan from
+left to right, we track $l$-mer occurrences, and look for repeated
+$l$-mer sequences.  When we find the same sequence of $l$-mers
+occurring multiple times in a row, we can mark it as a tentative family,
+then break it down later if we discover violations of the minimality
+condition.  The algorithm is summarized in Figueroa {\it et al.}, and
+discussed in detail in the Figueroa Thesis \cite{Figueroa:2014uk,Figueroa:2013cz}.
+
+Results of the preliminary implementation were promising.  Based on
+the Saha {\it et al} analysis putting RepeatMasker as the top {\it de
+  novo} search tool, comparisons were against that.  RAIDER result
+quality was close, but slightly behind -- with runtime
+order of magnitudes better.  On human chromosome 22 we saw a
+$12\times$ speedup (2344 seconds to 192 seconds), while coverage
+actually improved (77\% to 84\%), while on mouse chromosome 19 we saw 
+the same speedup with a significant drop in coverage (53\% to 30\%).  On
+the full human genome RAIDER ran in $6.3$ hours, while RepeatScout was
+unable to complete it run.  For details, see Figueroa {\it et al.} \cite{Figueroa:2014uk}.
+
+\subsection*{Spaced Seeds}
+
+PatternHunter, first proposed as a very successful augmentation to
+BLAST \cite{Altschul:1997p843,Li:2004wl}, is based on the notion of {\it
+  spaced seeds}.  In the context of BLAST, instead of looking for a
+contiguous substring match to trigger the alignment extension,
+PatternHunter uses a more refined match pattern.  For example, the
+{\it seed} 1111110111111 would specify that we were looking for an exact
+match of two length-six substrings separated by a single base (which
+might or might not match).  In the toy seed $s=110101$, we would say the
+strings $AACACA$ and $AAGACA$ because all positions corresponding to a
+1 match.  The string $AACACA$ does not math $TACACA$ (with respect to
+$s$), as the single miss-match does not correspond to the 1.  Nor does
+$AACACA$ match ant substring of $TTAACTCA$.
+
+The incorporation of spaced seeds into BLAST lead to significant
+improvements at virtually no cost in runtime \cite{Li:2004wl}, so it made sense to
+incorporate them here.  Specifically, instead of looking at repeat
+elements as a set of identical strings, we could require that they
+match with respect to a PatternHunter spaced seed.  The preliminary
+version of RAIDER from Figueroa {\it et al.} made some attempt to
+incorporate this -- was successful enough to provide a
+proof-of-concept, but a significant amount of work was left to be done
+for the development of phRAIDER (PatternHunter-based RAIDER) -- the
+main method in this paper.
+
+\section*{Methods}
+
+In order to allow the use of a spaced seed strategy, we will need to
+define a new model of elementary repeats that will accommodate it.  In
+the following we will first outline our theoretical framework
+(described in detail in the Supplementary Materials), and then review
+the phRAIDER algorithm.
+
+\subsection*{Seeded Elementary Repeats}
+In order to present our new model of elementary repeats, we first need
+to define a few terms:
+\begin{enumerate}
+  \item A spaced seed is a binary string that starts and ends with 1
+    symbols (as defined in Li {\it et al.} \cite{Li:2004wl}.
+  \item A {\it sequence descriptor} is a DNA string that allows *
+    symbols (indicating that a position where base content does not
+    matter).
+  \item We say the frequency of $r$ in a genomic sequence $G$ (denoted
+    $\nu_G(r)$) if the number of sequences in $G$ that match $r$.
+    (That is, are the same length, and are the same as $r$ in all
+    non-* positions.
+  \item A spaced seed $s$ {\it hits} a substring of sequence
+    descriptor $r$ if that substring is the same length of $s$ and,
+    when aligned, every $*$ in the substring matches a 0 in the
+    string.  (Example: if $s=11011$ and $r=AAA*TTTTT$, then $s$
+    hits both $AA*CC$ and $TTTTT$, but not $A*TTT$.)
+  \item The {\it decomposition} of $r$ (with respect to $s$) is the set of all substrings of
+    $r$ that are hit by $s$.  For the {\it generalized decomposition},
+    we take each member of the decomposition and replace by $*$ any
+    base that aligned to a 0 in the seed.  (Example: if $s=11011$ and
+    $r=AAATTT$, then $AAATT$ is in the decomposition, so $AA*TT$ is in
+    the generalized decomposition.)
+  \item We say a string $s$ {\it covers} a sequence descriptor $r$ if
+  every base in $r$ is contained within at least one substring from
+  the decomposition of $r$.
+\end{enumerate}
+
+The Zheng and Lonardi elementary repeats are all identical, hence can
+be described by a single string.  Because we allow variation, we
+required the flexibility of the sequences descriptor.  Leading us to
+our definition:
+
+\begin{definition}
+  \label{REDDef}
+  Given a fixed genomic sequence $G$, a integer $f$, and spaced seed $s$, a sequence
+  descriptor $r$ is a {\it repeat element descriptor} if:
+  \begin{enumerate}
+  \item $s$ covers $r$.  (Minimum length requirement.)
+  \item $\nu_G(r) \geq f$.  (Frequency requirement.)
+  \item For every string $w$ in the generalized decomposition of $r$
+    (w.r.t. $s$), $\nu_G(w) = \nu_G(r)$.  (Minimality requirement.)
+  \item There is no sequence descriptor $r'$ such that (1) $r$ is
+    a substring of $r'$, (2) $\nu_G(r') = \nu_G(r)$, and (3) $r'$
+    satisfies conditions 1-3.  (Maximality requirement.)
+  \end{enumerate}
+\end{definition}
+
+We will refer to the set of sequences in the genome that match the
+repeat family descriptor to be the repeat element family of the
+descriptor.
+
+\begin{theorem}
+Definition~\ref{REDDef} is equivalent to the Z\&L
+definition of elementary repeats when $s$ contains no 0 symbols.
+\end{theorem}
+In the Supplementary materials we prove our definition equivalent to
+the Figueroa definition (Definition 5 from the thesis) -- which is
+proved to be equivalent to the Z\&L definition in that work.
+
+It is instructive to look at a specific example.  Consider the toy
+seed $11011$, let $f=2$, and assume the genome contains a unique
+instance of $g_1 = AAAAA{\mathbf C}TTTTT$ and a unique instance of
+$g_2 = AAAAA{\mathbf G}TTTTT$.  Consider the sequence descriptor
+$r_1 = AAAAA*TTTTT$.  A simple inspection shows us that conditions (1)-(3)
+of Definition~\ref{REDDef} hold, so if we assume the surrounding
+sequence is such that (4) holds as well, this is an elementary repeat
+descriptor.  (As is $AA*AA*TT*TT$ -- the elementary repeat descriptor
+is not unique.)  But if we add a a second instance of $g_3 = AAAAAGTTTTT$,
+something happens that was not possible.  While $r_1$ continues to
+meet the definition, the $g_2/g_3$ subsequence $r_1 = AAAAGTTTT$ also
+induces an elementary repeat family $r_2 = AAAA*TTTT$. (Note that we can not
+extend this by, for example, one bast to the left, as the $AAAAA$
+substring would cause a violation of condition (3)).  So $r_2$
+corresponds to strings in $G$ that are substrings of those that $r_1$
+matches -- a condition not possible with the seedless elementary
+repeats.  This also then violates one of our core observation on which
+RAIDER is broken: we now
+have an $l$-mer ($AAAA*$ that belongs to multiple descriptors
+
+A second complicating factor is that our repeat element descriptor is
+not a sequence of consecutive $l$-mers. In the seedless version, an
+elementary repeat of length $n$ can be viewed as a sequence of $n-l+1$
+$l$-mers, each overlapping the last by exactly $l-1$ bases. In this
+definition we use members of the generalize decomposition, and find
+they are not necessarily offset by just one base: given the $g_1$ and
+$g_2$ example above, we see that the $r_1 = AAAAA*TTTTT$ descriptor is
+composed of $AA*AA$, $AA*TT$, and $TT*TT$ -- each offset from the last
+by 3.  As RAIDER assumed that it need a look-back of only one, this
+too breaks the algorithm.
+
+\subsection*{phRAIDER}
+
+In order to adapt the RAIDER algorithm to the new definition, we
+require observations parallel to those on which RAIDER was built.  As
+it turns out, its easy to adopt those.  Consider, for example, the
+first observation: ``An $l$-mer cannot belong to two different
+elementary repeats.''  We have already seen a violation, but can
+eliminate this by changing the definition of {\it belongs}:
+\begin{definition}
+  Given a elementary repeat descriptor $r$ for seed $s$ with length
+  $l$, we say a length $w$ substring \underline{belongs} to $r$ is it
+  matches a member of the generalized decomposition of $r$ (w.r.t. to
+  $s$.)
+\end{definition}
+We notice in the above example, the string $AAAAA$ belongs to the
+generalized decomposition, hence could not appear in both elementary
+repeats.  But the string $AAAAT$ does not match any string in the
+generalized decomposition of $r$, and hence was free to belong to
+the descriptor of a ``distinct'' family.
+
+\subsection*{phRAIDER Algorithm}
 
@@
 \bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
-\bibliography{bmc_article}      % Bibliography file (usually '*.bib' )
+\bibliography{bmc}      % Bibliography file (usually '*.bib' )
 % for author-year bibliography (bmc-mathphys or spbasic)
@@
 
-\section*{Figures}
-  \begin{figure}[h!]
-  \caption{\csentence{Sample figure title.}
-      A short description of the figure content
-      should go here.}
-      \end{figure}
+%% \section*{Figures}
+%%   \begin{figure}[h!]
+%%   \caption{\csentence{Sample figure title.}
+%%       A short description of the figure content
+%%       should go here.}
+%%       \end{figure}
 
-\begin{figure}[h!]
-  \caption{\csentence{Sample figure title.}
-      Figure legend text.}
-      \end{figure}
+%% \begin{figure}[h!]
+%%   \caption{\csentence{Sample figure title.}
+%%       Figure legend text.}
+%%       \end{figure}
 
@@
 %%
-\section*{Tables}
-\begin{table}[h!]
-\caption{Sample table title. This is where the description of the table should go.}
-      \begin{tabular}{cccc}
-        \hline
-           & B1  &B2   & B3\\ \hline
-        A1 & 0.1 & 0.2 & 0.3\\
-        A2 & ... & ..  & .\\
-        A3 & ..  & .   & .\\ \hline
-      \end{tabular}
-\end{table}
+%% \section*{Tables}
+%% \begin{table}[h!]
+%% \caption{Sample table title. This is where the description of the table should go.}
+%%       \begin{tabular}{cccc}
+%%         \hline
+%%            & B1  &B2   & B3\\ \hline
+%%         A1 & 0.1 & 0.2 & 0.3\\
+%%         A2 & ... & ..  & .\\
+%%         A3 & ..  & .   & .\\ \hline
+%%       \end{tabular}
+%% \end{table}
 
@@
 
-\section*{Additional Files}
-  \subsection*{Additional file 1 --- Sample additional file title}
-    Additional file descriptions text (including details of how to
-    view the file, if it is in a non-standard format or the file extension).  This might
-    refer to a multi-page table or a figure.
+%% \section*{Additional Files}
+%%   \subsection*{Additional file 1 --- Sample additional file title}
+%%     Additional file descriptions text (including details of how to
+%%     view the file, if it is in a non-standard format or the file extension).  This might
+%%     refer to a multi-page table or a figure.
 
-  \subsection*{Additional file 2 --- Sample additional file title}
-    Additional file descriptions text.
+%%   \subsection*{Additional file 2 --- Sample additional file title}
+%%     Additional file descriptions text.
 
@@
 \end{document}
+
+%%  LocalWords:  de addressref aff neteref Schaeffer noteref Xiaolin
+%%  LocalWords:  Liu corref JE Karro Initio phRAIDER genomic TEs Ns
+%%  LocalWords:  JIANG RepeatMasker nHMMER BLASTing homology ALU ReAS
+%%  LocalWords:  LINEs Jiang PILER RepeatScout mer RepeatGluer Saha
+%%  LocalWords:  infeasible DeBrujin et al Zheng Lonardi pre Huo mers
+%%  LocalWords:  PatternHunter decompositions subsequence minimality
+%%  LocalWords:  maximality RIADER ers substring AACACA AAGACA TACACA
+%%  LocalWords:  TTAACTCA TTTTT TTT substrings AAATTT AAATT TT AAAAA
+%%  LocalWords:  FIGUREOA AAAAAGTTTTT AAAAGTTTT AAAA TTTT AAAAT
+%%  LocalWords:  Acknowledgements

### 2015-07-27 manuscript/BMC/bmc_article.tex
Updating paper

@@
 transposable elements (the discovery and annotation of transposable
-elements without the use of a pre-complied library describing sequence
-or structural features of the family) has been addressed by a number
-of tools -- all of which are either based on computationally complex
-algorithms that cannot scale to whole-genome use, or whose sensitivity
-suffers significantly from the presence of the sequence variation
-present in all sets.  Here we present phRAIDER (Pattern Hunter based
-Rapid Ab Initio Detection of Elementary Repeats) for quickly masking
-a newly sequence, or poorly annotated, genome of transposable elements
+elements without the use of a pre-complied profile) has been addressed
+by a number of tools -- all of which are either based on
+computationally complex algorithms that cannot scale to whole-genome
+use, or whose sensitivity suffers significantly from the presence of
+sequence variation.  Here we present phRAIDER (Pattern Hunter based
+Rapid Ab Initio Detection of Elementary Repeats) for quickly masking a
+newly sequence, or poorly annotated, genome of transposable elements
 that is robust to the sequence variation present in real data.
@@
 
-Here we introduce phRAIDER
-
-\parttitle{Second part title} %if any
-Text for this section.
 \end{abstract}
@@
 order genome, covering as much as 45\% of the human genome and 90\% of
-the maize genome [CHECK THIS] \cite{Venter:2001p92} [CITE MAIZE].  TEs have
-proved an importent sorce of data in a number of studies \cite{Arndt:2005p279,Karro:2008p123,Mugal:2009p581,Hardison:2003p97}, but
-given their prevalence, it is just important for those not interested
-in them to have them masked out -- their bases replaced by N or case
-changed for easy identification and filtering.  For example. if not
+the maize genome [CHECK THIS] \cite{Venter:2001p92} [CITE MAIZE].  TEs
+have proved an importent sorce of data in a number of studies
+\cite{Arndt:2005p279,Karro:2008p123,Mugal:2009p581,Hardison:2003p97},
+but given their prevalence, it is just important for those studying
+the genome in them to have them masked out -- their bases replaced by
+N to allow for easy identification and filtering.  For example. if not
 filtered, TEs can trigger huge numbers of false positives automated
-gene finding tool \cite{Jiang:2013jt}, as well as inflate tool runtime.
-Hence the 09
+gene finding tool \cite{Jiang:2013jt}, as well as inflate tool
+runtime.
 
@@
 \cite{Pratchett:uw}, we must ask how the libraries on which the tools are based are
-initially compiled?  How are families initially discovered, and what
-do we do with a newly sequenced genome?
+initially compiled.  Such tools are useful for the discovery of new
+families, and cannot be applied to a newly sequenced genome for which
+such a library has not been compiled.
 
@@
 of a human ALU sequence can likely be used for the initial annotation
-of a newly sequenced primate genome, and many LINEs will extend to
+of a newly sequenced primate genome, and many human LINEs will extend to
 rodent, or even more distant, mammalian species.  This does not hold
@@
 literature, operating on a number of different principles.  Tools such
-as RECON \cite{Bao:2002} and PILER \cite{Edgar:2005p2365} are based on
+as RECON and PILER are based on
 self-alignment, using WU-BLAST and LASTZ for the alignment
-\cite{Lopez:2003td,Harris:2007uf}.  RECON show good sensitivity but is
+\cite{Bao:2002,Edgar:2005p2365,Lopez:2003td,Harris:2007uf}.  RECON show good sensitivity but is
 computationally intensive and infeasible for use on whole genomes
-(requiring 60 hours for 18Mb rice genome in a 2013 report), while
+(requiring 60 hours for 18Mb rice genome int 2013 \cite{Jiang:2013jt}), while
 PILER achieves a good runtime with very low sensitivity
@@
 
+[TO DO: Make sure we mention somewhere that RAIDER should work for
+  unassembled genomes.]
+
 \subsection*{Elementary Repeats}
@@
 decompositions of TEs into basic building blocks.  Identification of
-these building blocks are be sufficient of their own for the purposed
+these building blocks are sufficient for the purpose
 of masking, and can be assembled into Transposible Elements for those
@@
 
-Zheng and Lonardi formulated their definition with the explicit
-intention of factoring minimum length and 
-frequency into the definition: to be an elementary repeat a sequence
-must have some minimum length $l$ and occur at least $f$ times.  More
-importantly, it must be minimal (that is, no section of an elementary
-repeat can itself be an elementary repeat), and it must have appear
-independently of any other sequence meeting definition of an
-elementary repeat  Formally:
-\begin{definition}
-Give a length criteria $l$ and a frequency criteria $f$, a sequence
-$r$ is an elementary repeat in a genome if:
+While it is noteriously difficult to state a mathematical definition
+of elementary repeats \cite{Bao:2002}, elemenary repeats are a bit
+easier to model.  For a given genome, a nucleotide sequence $r$ is an
+elementary repeat if:
 \begin{enumerate}
-\label{ZLDef}
-\item $r$ is of length at least $l$.  (The length condition.)
-\item There are at least $f$ copies of $r$ in the genome.  (The
-  frequency condition.)
-\item There is no proper subsequence $r'$ of $r$ such that $r'$ is of
-  length $l$ or greater and $r'$ occurs independently of $r$.  That
-  is, every length $>l$ subsequence of $r'$ occurs only where $r'$
-  occurs in the genome.  (The minimality condition.)
-\item There is no sequence $r'$ that properly contains $r$ and appears
-  around every occurrence of $r$ in the genome.  (The maximality condition.)
+\item It is of at least length $l$.  (The length requirement.)
+\item At least $f$ exact copies of $r$ appear in the genome.  (The
+  frequency requirement.)
+\item There is no proper substring of $r$ that is of at least length
+  $l$ and appears in the genome independently of $r$. (The minimality
+  requirement.)
+\item There is no string that contains $r$, appears in the genome
+  the same number of times as $r$, and satisfies conditions 1-3.  (The
+  maximality requirement.)
 \end{enumerate}
-\end{definition}
+The first two criteria are straight-foward, allowing us to designate a
+threshold length and frequency for our model.  The minimality
+requirement ensures that an elementary repeat does not contain
+anything that could be viewed as an independent elementary repeat --
+that it is the ``smallest building block'' that we can find.  While
+the maximality requirement ensures that it is not tied to some other, larger,
+elementary repeat.
+
 Having proposed this definition, Zheng and Lonardi developed an
@@
 $l$-mers (sequences of length $l$).  Specifically, it was observed
-that condition (3) of Definition~\ref{ZLDef} could be rewritten as:
+that the minimality condition could be rewritten as: 
 There is no $l$-mer contained with $r$ that appears in the genome
-independently of $r$.
-
-\begin{observation}
-  \label{RAIDERobs}
-  From the definition, we can now observe:
+more times than $r$.  From there we make our four core observations that form
+the basis for the RAIDER algorithm:
 \begin{enumerate}
@@
 \item Any $l$-mer in the genome that occurs $f$ or more times is
-  either an elementary repeat or belong to one.
+  either an elementary repeat or belongs to one.
 \item Any two $l$-ers belonging to an elementary repeat must appear
@@
 \end{observation}
-By ``belong'' we mean ``is a substring of'' -- the reason for the
+By ``belong'' we mean ``is a substring of''; the reason for the
 terminology will be explained shortly.  For discussion and proof, see
@@
 repeats in a single scan of the genome.  Specifically, as we scan from
-left to right, we track $l$-mer occurrences, and look for repeated
-$l$-mer sequences.  When we find the same sequence of $l$-mers
+left to right, we track $l$-mer occurrences and identify multiple
+copies of the same $l$-mer.  When we find the same sequence of $l$-mers
 occurring multiple times in a row, we can mark it as a tentative family,
@@
 order of magnitudes better.  On human chromosome 22 we saw a
-$12\times$ speedup (2344 seconds to 192 seconds), while coverage
-actually improved (77\% to 84\%), while on mouse chromosome 19 we saw 
+$12\times$ speedup from RepeatScout to RAIDER (2344 seconds to 192 seconds), while coverage
+of the RepBase ancesteral sequecne \cite{Jurka:2005bl} actually improved (77\% to 84\%), while on mouse chromosome 19 we saw 
 the same speedup with a significant drop in coverage (53\% to 30\%).  On
@@
 PatternHunter, first proposed as a very successful augmentation to
-BLAST \cite{Altschul:1997p843,Li:2004wl}, is based on the notion of {\it
+BLAST \cite{Li:2004wl,Altschul:1997p843}, is based on the notion of {\it
   spaced seeds}.  In the context of BLAST, instead of looking for a
@@
 PatternHunter uses a more refined match pattern.  For example, the
-{\it seed} 1111110111111 would specify that we were looking for an exact
+seed 1111110111111 would specify that we were looking for an exact
 match of two length-six substrings separated by a single base (which
@@
 $s$), as the single miss-match does not correspond to the 1.  Nor does
-$AACACA$ match ant substring of $TTAACTCA$.
+$AACACA$ match any substring of $TTAACTCA$, though it does match a
+substring of $T{\mathbf AAGATA}C$.
 
 The incorporation of spaced seeds into BLAST lead to significant
-improvements at virtually no cost in runtime \cite{Li:2004wl}, so it made sense to
-incorporate them here.  Specifically, instead of looking at repeat
-elements as a set of identical strings, we could require that they
-match with respect to a PatternHunter spaced seed.  The preliminary
-version of RAIDER from Figueroa {\it et al.} made some attempt to
-incorporate this -- was successful enough to provide a
-proof-of-concept, but a significant amount of work was left to be done
-for the development of phRAIDER (PatternHunter-based RAIDER) -- the
-main method in this paper.
+improvements at virtually no cost in runtime \cite{Li:2004wl}.
+SpecificallyIn our context, we will use them to allow limited
+miss-matches between instances of ``the same'' elementary repeat.  The
+preliminary version of RAIDER from Figueroa {\it et al.} made some
+attempt to incorporate this -- was successful enough to provide a
+proof-of-concept, but a significant amount of work was left to be
+done.
+
+In this paper we present phRAIDER (PatternHunter-based RAIDER),
+implemented as open source code that may be obtained [NEED WEB
+  ADDRESS].  In our Methods section we present the new model that
+allows us to extend RAIDER to correctly use spaced seeds, and briefly
+outline the algorithm (with more details provided in the supplmentary
+materials), with a analsis of phRAIDER performance in our Results.
 
@@
   \item A spaced seed is a binary string that starts and ends with 1
-    symbols (as defined in Li {\it et al.} \cite{Li:2004wl}.
+    symbols (as defined in Li {\it et al.} \cite{Li:2004wl}).
   \item A {\it sequence descriptor} is a DNA string that allows *
     symbols (indicating that a position where base content does not
-    matter).
+    matter).  A string matches a sequence decriptor of equal length if
+    it has the same base in all non-* positions.
   \item We say the frequency of $r$ in a genomic sequence $G$ (denoted
-    $\nu_G(r)$) if the number of sequences in $G$ that match $r$.
-    (That is, are the same length, and are the same as $r$ in all
-    non-* positions.
+    $\nu_G(r)$) is the number of subsequences in $G$ that match $r$.
   \item A spaced seed $s$ {\it hits} a substring of sequence
@@
 the Figueroa definition (Definition 5 from the thesis) -- which is
-proved to be equivalent to the Z\&L definition in that work.
+proved to be equivalent to the Z\&L definition in that work \cite{Figueroa:2014uk}.
 
@@
 instance of $g_1 = AAAAA{\mathbf C}TTTTT$ and a unique instance of
-$g_2 = AAAAA{\mathbf G}TTTTT$.  Consider the sequence descriptor
-$r_1 = AAAAA*TTTTT$.  A simple inspection shows us that conditions (1)-(3)
+$g_2 = AAAAA{\mathbf G}TTTTT$.  Consider the sequence descriptor $r_1
+= AAAAA*TTTTT$.  A simple inspection shows us that conditions (1)-(3)
 of Definition~\ref{REDDef} hold, so if we assume the surrounding
 sequence is such that (4) holds as well, this is an elementary repeat
-descriptor.  (As is $AA*AA*TT*TT$ -- the elementary repeat descriptor
-is not unique.)  But if we add a a second instance of $g_3 = AAAAAGTTTTT$,
-something happens that was not possible.  While $r_1$ continues to
-meet the definition, the $g_2/g_3$ subsequence $r_1 = AAAAGTTTT$ also
-induces an elementary repeat family $r_2 = AAAA*TTTT$. (Note that we can not
-extend this by, for example, one bast to the left, as the $AAAAA$
-substring would cause a violation of condition (3)).  So $r_2$
-corresponds to strings in $G$ that are substrings of those that $r_1$
-matches -- a condition not possible with the seedless elementary
-repeats.  This also then violates one of our core observation on which
-RAIDER is broken: we now
-have an $l$-mer ($AAAA*$ that belongs to multiple descriptors
+descriptor for this family of sequences.  (Note $AA*AA*TT*TT$ -- the
+elementary repeat descriptor is not unique.)  But if we add a a second
+instance of $g_3 = AAAAAGTTTTT$, something happens that was not
+possible in the unseeded version.  While $r_1$ continues to meet the definition, the $g_2/g_3$
+subsequence $AAAAGTTTT$ also induces an elementary repeat family
+$r_2 = AAAA*TTTT$. (Note that we cannot extend this by, for example,
+one base to the left, as the $AAAAA$ substring would cause a violation
+of condition (3)).  So $r_2$ corresponds to strings in $G$ that are
+substrings of those that $r_1$ matches -- a condition not possible
+with the seedless elementary repeats.  This also then violates one of
+our core observation on which RAIDER is broken: we now have an $l$-mer
+($AAAA*$) that belongs to multiple descriptors.
 
@@
   Given a elementary repeat descriptor $r$ for seed $s$ with length
-  $l$, we say a length $w$ substring \underline{belongs} to $r$ is it
+  $l$, we say a length $w$ substring \underline{belongs} to $r$ if it
   matches a member of the generalized decomposition of $r$ (w.r.t. to

### 2015-07-27 RAIDER_eval.py
updated options

@@ -357,11 +357,9 @@ def simulate_chromosome(chromosome_file, rng_seed, length, neg_strand, fam_file,
     batch_name = data_dir + "/" + output_file + ".sim.batch"
     job_name = "simulation.%d" % (get_job_index("simulation"))
     
-    #print("Sim batch: %s\n" % (batch_name))
     p = pbsJobHandler(batch_file = batch_name, executable = cmd, job_name = job_name,
                       stdout_file = output_file + ".stdout", stderr_file = output_file + ".stderr", 
                       output_location = data_dir, walltime = time_limit, arch_type = ["n09","bigmem"])
-    #p.submit(preserve=True)
     if not timing_jobs:
         p.submit(preserve=True)
     else:
@@ -373,12 +371,7 @@ def simulate_chromosome(chromosome_file, rng_seed, length, neg_strand, fam_file,
     p.index = file_index
     return p
 
-# def run_raider_chrom(p, seed, f, m, output_dir):
-#    """Given the pbs object used to start simulated chromosome job as well as 
-#    raider parameters, wait until the job is done and then call run_raider 
-#    on the newly generated chromosome output file"""
-#    p.wait()
-#    return run_raider(seed, f, m, p.chrom_output, output_dir, p.curr_dir)
+
 def run_raider(seed, seed_num, f, m, input_file, raider_dir, mem):
     """Given raider parameters and an input file, run RAIDER and put the output into
     the directory specified in output_dir (creating a random name is none is
@@ -405,12 +398,10 @@ def run_raider(seed, seed_num, f, m, input_file, raider_dir, mem):
 
     batch_name = raider_dir + "/" + input_base + ".raider.batch"
     job_name = "R.{input}.{seed}.{num}".format( num = get_job_index("raider") , input=re.sub("hg18.","",input_base), seed=seed_num)
-    #progress_fp.write("Sim batch: " + batch_name + "\n")
     p = pbsJobHandler(batch_file = batch_name, executable = cmd1 + "; " + cmd2, job_name = job_name,
                       stdout_file = input_base + ".raider.stdout", stderr_file = input_base + ".raider.stderr",
                       output_location = output_dir, walltime = time_limit, mem = mem, #ppn = 8 if mem else 1, 
                       arch_type = ['n09'])
-    #p.submit(preserve=True)
[…]

### 2015-07-27 RAIDER_src_v2_options_new/src/Family.h

@@ -9,150 +9,150 @@ using namespace std;
 
 class Family {
 public:
-
-	void adopt(LmerVector *v) {
-		v->setFamily(this);
-		vectors.push_back(v);
-		setLast(v);
-		setExpectedEnd(v->back() + 1);
-	}
-
-    void adopt(LmerVector *v, uint L){
-        v->setFamily(this);
-        //cout << L << endl;
-        if (vectors.size() > 0){
-            uint off = v->front() - vectors.back()->front();
-            addToOffset(off - 1);
-            setRepeatLength(getRepeatLength() + off);
-            v->setOff(off - 1);
-        }
-        else{
-            setRepeatLength(L);
-            setOffset(0);
-            v->setOff(0);
-        }
-        vectors.push_back(v);
-        setLast(v);
-        setExpectedEnd(v->back() + L);
+  
+  void adopt(LmerVector *v) {
+    v->setFamily(this);
+    vectors.push_back(v);
+    setLast(v);
+    setExpectedEnd(v->back() + 1);
+  }
+  
+  void adopt(LmerVector *v, uint L){
+    v->setFamily(this);
+    //cout << L << endl;
+    if (vectors.size() > 0){
+      uint off = v->front() - vectors.back()->front();
+      addToOffset(off - 1);
+      setRepeatLength(getRepeatLength() + off);
+      v->setOff(off - 1);
     }
-
-	uint size() const {
-		if (vectors.size() == 0) {
-			return 0;
-		}
-		return vectors.front()->size();
-	}
-
-    void addToOffset(uint off){
-        offset += off;
+    else{
+      setRepeatLength(L);
+      setOffset(0);
+      v->setOff(0);
     }
-
-    void setOffset(uint off){
-        offset = off;
+    vectors.push_back(v);
+    setLast(v);
+    setExpectedEnd(v->back() + L);
+  }
+  
+  uint size() const {
+    if (vectors.size() == 0) {
+      return 0;
     }
-
-    uint getOffset() const {
-        return offset;
+    return vectors.front()->size();
+  }
+  
+  void addToOffset(uint off){
+    offset += off;
+  }
+  
+  void setOffset(uint off){
+    offset = off;
+  }
+  
+  uint getOffset() const {
+    return offset;
+  }
+  
+  uint getRepeatLength() const {
+    return repeatlength;
+  }
+  
+  void setRepeatLength(uint L){
[…]

### 2015-07-27 RAIDER_src_v2_options_new/src/LmerVector.h

@@ -14,71 +14,71 @@ class Family;
 
 class LmerVector {
 public:
-
-	LmerVector(uint position) {
-		family = nullptr;
-		push_back(position);
-	}
-    
-    ~LmerVector() {
+  
+  LmerVector(uint position) {
+    family = nullptr;
+    push_back(position);
+  }
+  
+  ~LmerVector() {
+  }
+  
+  void push_back(uint val) {
+    if(lmers.size() > 0){
+      previous = back();
     }
-
-	void push_back(uint val) {
-        if(lmers.size() > 0){
-            previous = back();
-        }
-		lmers.push_back(val);
-	}
-
-	uint operator [](uint i) const {
-		return lmers[i];
-	}
-
-	uint & operator [](uint i) {
-		return lmers[i];
-	}
-
-    uint prev() const {
-        return previous;
-    }
-
-	uint back() const {
-		return lmers.back();
-	}
-
-	uint front() const {
-		return lmers.front();
-	}
-
-    uint getOff() const {
-        return off;
-    }
-    
-    void setOff(uint i) {
-        off = i;
-    }
-
-	uint size() const {
-		return lmers.size();
-	}
-
-	Family* getFamily() const {
-		return family;
-	}
-
-	void setFamily(Family *fam) {
-		family = fam;
-	}
-
-    friend ostream &operator<<( ostream &output, const LmerVector &L) { 
-        output << "(\tFront:\t" << L.front() << "\tBack:\t" << L.back() << "\tSize:\t" << L.size() << "\tOffset In Family\t" << L.getOff() << "\tFamily:\t" << L.getFamily() << "\t)";
-        return output;            
-    }
-
-	vector<uint> lmers;
-	Family* family;
-    uint previous;
-    uint off;
+    lmers.push_back(val);
+  }
+  
+  uint operator [](uint i) const {
+    return lmers[i];
+  }
+  
+  uint & operator [](uint i) {
+    return lmers[i];
+  }
+  
+  uint prev() const {
+    return previous;
+  }
+  
+  uint back() const {
+    return lmers.back();
+  }
+  
+  uint front() const {
+    return lmers.front();
+  }
+  
+  uint getOff() const {
+    return off;
+  }
+  
+  void setOff(uint i) {
+    off = i;
+  }
+  
+  uint size() const {
+    return lmers.size();
+  }
+  
+  Family* getFamily() const {
+    return family;
+  }
+  
[…]

### 2015-07-27 RAIDER_src_v2_options_new/src/SeedChain.h

@@ -1,5 +1,5 @@
-#ifndef SCANER_SEEDCHAIN
-#define SCANER_SEEDCHAIN
+#ifndef RAIDER2_SEEDCHAIN_H
+#define RAIDER2_SEEDCHAIN_H
 
 #include <seqan/sequence.h>
 #include <seqan/seq_io.h>
@@ -8,6 +8,7 @@
 #include "Family.h"
 #include <unordered_map>
 #include <ctime>
+#include <string>
 
 using namespace std;
 
@@ -16,484 +17,453 @@ typedef unordered_map<size_t, LmerVector*> LmerMap;
 int p = getpid();
 
 void prettyPrintTabbing(uint t){
-    for(uint i = 0; i < t; i++){
-        cout << '\t';
-    }
+  for(uint i = 0; i < t; i++){
+    cout << '\t';
+  }
 }
 
 void prettyPrintLmer(int t, LmerVector* v){
-    prettyPrintTabbing(t);
-    cout << "Lmer:\t" << *v << endl << endl;
+  prettyPrintTabbing(t);
+  cout << "Lmer:\t" << *v << endl << endl;
 }
 
 
 void prettyPrintFamily(uint t, Family* fam, bool isNew, bool isMod){
-    prettyPrintTabbing(t);
-    if (isNew) {
-        cout << "New family:\t" << *fam << endl << endl;
-    }
-    else if (isMod) {
-        cout << "Modified family:\t" << *fam << endl;
-    }
-    else{
-        cout << "Family:\t" << *fam << endl;
-    }
+  prettyPrintTabbing(t);
+  if (isNew) {
+    cout << "New family:\t" << *fam << endl << endl;
+  }
+  else if (isMod) {
+    cout << "Modified family:\t" << *fam << endl;
+  }
+  else{
+    cout << "Family:\t" << *fam << endl;
+  }
 }
 
+void prettyPrintMethodState(uint t, string methodname, LmerVector* v, Family* fam, bool isNew, bool isMod){
+  prettyPrintTabbing(t);
+  cout << "--- " << methodname << " ---" << endl;
+  prettyPrintLmer(t + 1, v);
+  prettyPrintFamily(t + 1, fam, isNew, isMod);
+}
+
+
+// Determines numeric representation of base. Only works for meaningful bases
+// ('A', 'C', 'G', 'T'). Helper for hashing function.
 size_t baseToInt(char b) {
-	switch (toupper(b)) {
[…]

### 2015-07-27 RAIDER_src_v2_options_new/src/main.cpp

@@ -36,21 +36,29 @@ const string OUTPUT_SUMMARY_FILENAME = "summary_info";
 
 // This struct stores the options from the command line.
 struct AppOptions {
-	// Age.  0 -- curr, 1 -- old, 2 -- oldest, 
-	int age;
-	// Verbosity level.  0 -- quiet, 1 -- normal, 2 -- verbose, 3 -- carly verbose
-	int verbosity;
-	// Minimum repeat length
-	uint min;
-	// Minimum number of repeats to be significant
-	uint count;
-
-	seqan::CharString seed;
-	seqan::CharString sequence_file;
-	seqan::CharString output_directory;
-
-	AppOptions() :
-			verbosity(1) {}
+  // Age. Presets for familyarray, excising, overlaps, and tieup
+  int age;
+  // Verbosity level.  0 -- quiet, 1 -- normal, 2 -- verbose, 3 -- carly verbose
+  int verbosity;
+  // Minimum repeat length
+  uint min;
+  // Minimum number of repeats to be significant
+  uint count;
+  // Whether or not to use family array
+  bool family_array;
+  // Whether or not to allow for excising
+  bool excising;
+  // Whether or not to require overlaps
+  bool overlaps;
+  // Whether or not to allow for modified tie up
+  bool tieup;
+  
+  seqan::CharString seed;
+  seqan::CharString sequence_file;
+  seqan::CharString output_directory;
+  
+  AppOptions() :
+  verbosity(1), age(1), family_array(true), excising(false), overlaps(true), tieup(false){}
 };
 
 // ==========================================================================
@@ -61,121 +69,147 @@ struct AppOptions {
 // Function parseCommandLine()
 // --------------------------------------------------------------------------
 seqan::ArgumentParser::ParseResult parseCommandLine(AppOptions & options, int argc, char const ** argv) {
-	// Setup ArgumentParser.
-	seqan::ArgumentParser parser("RAIDER2");
-	// Set short description, version, and date.
-	setShortDescription(parser, "RAIDER - Rapid Ab Initio Detection of Elementary Repeats");
-	setVersion(parser, "2.0");
-	setDate(parser, "June 2015");
+  // Setup ArgumentParser.
+  seqan::ArgumentParser parser("RAIDER2");
+  // Set short description, version, and date.
[…]

### 2015-07-27 RAIDER_src_v2_options_new/src/Family.h
Fixing make

@@ -1,158 +1,52 @@
-#ifndef SCANER_FAMILY
-#define SCANER_FAMILY
+#ifndef FAMILY_H
+#define FAMILY_H
 #include <vector>
 #include "LmerVector.h"
 #include <iostream>
 #include <assert.h>
 
-using namespace std;
-
 class Family {
 public:
-  
-  void adopt(LmerVector *v) {
-    v->setFamily(this);
-    vectors.push_back(v);
-    setLast(v);
-    setExpectedEnd(v->back() + 1);
-  }
-  
-  void adopt(LmerVector *v, uint L){
-    v->setFamily(this);
-    //cout << L << endl;
-    if (vectors.size() > 0){
-      uint off = v->front() - vectors.back()->front();
-      addToOffset(off - 1);
-      setRepeatLength(getRepeatLength() + off);
-      v->setOff(off - 1);
-    }
-    else{
-      setRepeatLength(L);
-      setOffset(0);
-      v->setOff(0);
-    }
-    vectors.push_back(v);
-    setLast(v);
-    setExpectedEnd(v->back() + L);
-  }
-  
-  uint size() const {
-    if (vectors.size() == 0) {
-      return 0;
-    }
-    return vectors.front()->size();
-  }
-  
-  void addToOffset(uint off){
-    offset += off;
-  }
-  
-  void setOffset(uint off){
-    offset = off;
-  }
-  
-  uint getOffset() const {
-    return offset;
-  }
-  
-  uint getRepeatLength() const {
-    return repeatlength;
-  }
-  
-  void setRepeatLength(uint L){
-    repeatlength = L;
-  }
-  
-  uint repeatLength(uint L) const {
-    return repeatlength + L - L;
-    //return vectors.size() + L - 1;
-  }
-  
-  vector<LmerVector*>* getLmers() {
-    return &vectors;
-  }
-  
-  LmerVector* findByPos(uint pos){
-    uint sumOff = 0;
-    if (pos == 0){
-      return vectors.front();
-    }
-    for (LmerVector* v : vectors){
-      sumOff += v->getOff();
-      if (pos == sumOff){
-        return v;
-      }
-      sumOff++;
-    }
-    return vectors.front();
-  }
-  
-  
-  LmerVector* at(uint index) {
-    return vectors.at(index);
-  }
-  
-  LmerVector* see(uint index) const {
-    return vectors.at(index);
-  }
-  
-  void push_back(LmerVector* v) {
-    vectors.push_back(v);
-  }
-  
[…]

### 2015-07-27 RAIDER_src_v2_options_new/src/LmerVector.cpp

@@ -1,0 +1,22 @@
+#include "LmerVector.h"
+
+LmerVector::LmerVector(uint position) {
+  family = nullptr;
+  push_back(position);
+}
+
+void LmerVector::push_back(uint val) {
+  if(lmers.size() > 0){
+    previous = back();
+  }
+  lmers.push_back(val);
+}
+
+std::ostream &operator<<( std::ostream &output, const LmerVector &L) {
+  output  << "(\tFront:\t" << L.front()
+          << "\tBack:\t" << L.back()
+          << "\tSize:\t" << L.size()
+          << "\tOffset In Family\t" << L.getOff()
+          << "\tFamily:\t" << L.getFamily() << "\t)";
+  return output;
+}

### 2015-07-27 RAIDER_src_v2_options_new/src/LmerVector.h

@@ -1,10 +1,8 @@
-#ifndef LMER_VECTOR
-#define LMER_VECTOR
+#ifndef LMER_VECTOR_H
+#define LMER_VECTOR_H
 #include <vector>
 #include <stdexcept>
 #include <iostream>
-
-using namespace std;
 
 #ifndef uint
 typedef unsigned int uint;
@@ -15,70 +13,36 @@ class Family;
 class LmerVector {
 public:
   
-  LmerVector(uint position) {
-    family = nullptr;
-    push_back(position);
-  }
+  LmerVector(uint position);
+  ~LmerVector() { }
   
-  ~LmerVector() {
-  }
+  void push_back(uint val);
   
-  void push_back(uint val) {
-    if(lmers.size() > 0){
-      previous = back();
-    }
-    lmers.push_back(val);
-  }
+  uint operator [](uint i) const { return lmers[i]; }
+  uint & operator [](uint i) { return lmers[i]; }
   
-  uint operator [](uint i) const {
-    return lmers[i];
-  }
+  uint prev() const { return previous; }
+  uint back() const { return lmers.back(); }
+  uint front() const { return lmers.front(); }
   
-  uint & operator [](uint i) {
-    return lmers[i];
-  }
+  uint getPosition() const { return position; }
+  void setPosition(uint i) { position = i; }
   
-  uint prev() const {
-    return previous;
-  }
+  uint getOff() const { return off; }
+  void setOff(uint i) { off = i; }
   
-  uint back() const {
-    return lmers.back();
-  }
+  uint size() const { return lmers.size(); }
   
-  uint front() const {
-    return lmers.front();
-  }
+  Family* getFamily() const { return family; }
+  void setFamily(Family *fam) { family = fam; }
   
-  uint getOff() const {
-    return off;
-  }
+  friend std::ostream &operator<<( std::ostream &output, const LmerVector &L);
   
-  void setOff(uint i) {
-    off = i;
-  }
-  
-  uint size() const {
-    return lmers.size();
-  }
-  
-  Family* getFamily() const {
-    return family;
-  }
-  
-  void setFamily(Family *fam) {
-    family = fam;
-  }
-  
-  friend ostream &operator<<( ostream &output, const LmerVector &L) {
[…]

### 2015-07-27 RAIDER_src_v2_options_new/src/SeedChain.h

@@ -211,8 +211,8 @@ Family* exciseRepeatsByLmer(Family* fam, LmerVector *v, uint L, uint verbosity) 
   newFam->setExpectedEnd(v->back() + newFam->repeatLength(L));
   
   if (verbosity > 2){
-    prettyPrintFamily(tabbing+1, fam, false, true);
-    prettyPrintFamily(tabbing+1, newFam, true, false);
+    prettyPrintFamily(2, fam, false, true);
+    prettyPrintFamily(2, newFam, true, false);
   }
   return newFam;
   
@@ -266,8 +266,8 @@ Family* splitRepeatsByLmer(Family* fam, LmerVector *v, bool keepV, uint L, uint 
   }
   
   if (verbosity > 2){
-    prettyPrintFamily(tabbing+1, fam, false, true);
-    prettyPrintFamily(tabbing+1, newFam, true, false);
+    prettyPrintFamily(2, fam, false, true);
+    prettyPrintFamily(2, newFam, true, false);
   }
   
   return newFam;
@@ -344,7 +344,7 @@ bool fragmentSplit(LmerVector* v, uint L, vector<Family*> &families, uint verbos
   // if it was previously skipped in forming a family instance
   else if (excising && v->prev() < fam->getLast()->prev() && v->size() < fam->size() ) {
     if (verbosity > 2){
-      prettyPrintTabbing(tabbing + 1);
+      prettyPrintTabbing(1);
       cout << "The lmer was skipped in a previous iteration" << endl;
     }
     // OLDEST (1):

### 2015-07-27 RAIDER_eval.py
switches for raider2 added

@@ -137,11 +137,11 @@ def parse_params(args):
     parser_tools.add_argument('-R', '--raider_on', dest = 'run_raider', action = 'store_true', help = 'Turn RAIDER on', default = False)
     parser_tools.add_argument('--R2', '--raider2_on', dest = 'run_raider2', action = 'store_true', help = 'Turn RAIDERV2 on', default = False)
     parser_tools.add_argument('--AR', '--araider_on', dest = 'run_araider', action = 'store_true', help = 'Turn ARAIDER on', default = False)
-    parser_tools.add_argument('--RS', '--repscout_on', dest = 'run_repscout', action = 'store_true', help = 'Turn RAINDER on', default = False)
+    parser_tools.add_argument('--RS', '--repscout_on', dest = 'run_repscout', action = 'store_true', help = 'Turn RAIDER on', default = False)
     parser_tools.add_argument('-B', '--bigfoot_on', dest = 'run_bigfoot', action = 'store_true', help = 'Turn BIGFOOT on', default = False)
     parser_tools.add_argument('-P', '--piler_on', dest = 'run_piler', action = 'store_true', help = 'Turn PILER on', default = False)
-    parser_tools.add_argument('-A', '--all_tools', dest = 'all_tools', action = 'store_true', help = 'Turn all tolls on (overide all other tool arguments)', default = False)
-    parser_tools.add_argument('--A2', '--all_tools2', dest = 'all_tools2', action = 'store_true', help = 'Turn all tolls on except araider (overide all other tool arguments)', default = False)
+    parser_tools.add_argument('-A', '--all_tools', dest = 'all_tools', action = 'store_true', help = 'Turn all tools on (overide all other tool arguments)', default = False)
+    parser_tools.add_argument('--A2', '--all_tools2', dest = 'all_tools2', action = 'store_true', help = 'Turn all tools on except araider (overide all other tool arguments)', default = False)
     parser_tools.add_argument('--tl', '--time_limit', dest = 'time_limit', help = 'Redhawk time limit (max: 400:00:00 default: 4:00:00)', default = default_time_limit)
     # Will later add: RepeatModeler, RECON, PILER (other?)
 
@@ -176,13 +176,17 @@ def parse_params(args):
     raider2_argument = parser.add_argument_group("RAIDER2 parameters")
     raider2_argument.add_argument('--age', type = int, help="Use older version of raider2", default=1) 
[…]

### 2015-07-28 RAIDER_src_v2_options_new/src/Family.h
modifications to allow for proactive splitting

@@ -4,49 +4,134 @@
 #include "LmerVector.h"
 #include <iostream>
 #include <assert.h>
+#include <algorithm>
 
 class Family {
 public:
-  void adopt(LmerVector *v);
-  void adopt(LmerVector *v, uint L);
-
-  uint size() const;
-
-  uint getOffset() const { return offset; }
-  void setOffset(uint off){ offset = off; }
-  void addToOffset(uint off);
-
-  uint getRepeatLength() const { return repeatlength; }
-  void setRepeatLength(uint L){ repeatlength = L; }
-  uint repeatLength(uint L) const;
-
-  std::vector<LmerVector*>* getLmers() { return &vectors; }
-
+  void adopt(LmerVector *v) {
+    v->setFamily(this);
+    vectors.push_back(v);
+    setLast(v);
+    setExpectedEnd(v->back() + 1);
+  }
+  
+  void adopt(LmerVector *v, uint L){
+    v->setFamily(this);
+    if (vectors.size() > 0){
+      uint off = v->front() - vectors.back()->front();
+      v->setOff(off - 1);
+    }
+    else{
+      v->setOff(0);
+    }
+    vectors.push_back(v);
+    setLast(v);
+    setExpectedEnd(v->back() + L);
+  }
+  
+  uint size() const {
+    if (vectors.size() == 0) {
+      return 0;
+    }
+    return vectors.front()->size();
+  }
+  
+  uint repeatLength(uint L) const {
+    return vectors.back()->front() - vectors.front()->front() + L;
+  }
+  
+  vector<LmerVector*>* getLmers() { return &vectors; }
+  
   LmerVector* at(uint index) { return vectors.at(index); }
   LmerVector* see(uint index) const { return vectors.at(index); }
   void push_back(LmerVector* v) { vectors.push_back(v); }
-  LmerVector* findByPos(uint pos);
-
-
+  
+  LmerVector* findByPos(uint pos){
+    uint sumOff = 0;
+    if (pos == 0){
+      return vectors.front();
+    }
+    for (LmerVector* v : vectors){
+      sumOff += v->getOff();
+      if (pos == sumOff){
+        return v;
+      }
+      sumOff++;
+    }
+    return vectors.front();
+  }
+    
   LmerVector* getPrefix() const { return vectors.front(); }
-  LmerVector* getSuffix() const { return vectors.back(); }
-
+  LmerVector* getSuffix() const { return vectors.back();  }
+  
[…]

### 2015-07-28 RAIDER_src_v2_options_new/src/LmerVector.h

@@ -4,6 +4,8 @@
 #include <stdexcept>
 #include <iostream>
 
+using namespace std;
+
 #ifndef uint
 typedef unsigned int uint;
 #endif
@@ -13,36 +15,58 @@ class Family;
 class LmerVector {
 public:
   
-  LmerVector(uint position);
-  ~LmerVector() { }
+  LmerVector(uint position) {
+    family = nullptr;
+    push_back(position);
+  }
   
-  void push_back(uint val);
+  ~LmerVector() {
+  }
+  
+  void push_back(uint val) {
+    if(lmers.size() > 0){
+      previous = back();
+    }
+    lmers.push_back(val);
+  }
   
   uint operator [](uint i) const { return lmers[i]; }
-  uint & operator [](uint i) { return lmers[i]; }
+  uint & operator [](uint i)     { return lmers[i]; }
   
-  uint prev() const { return previous; }
-  uint back() const { return lmers.back(); }
+  uint prev()  const { return previous; }
+  uint back()  const { return lmers.back(); }
   uint front() const { return lmers.front(); }
   
-  uint getPosition() const { return position; }
-  void setPosition(uint i) { position = i; }
-  
   uint getOff() const { return off; }
-  void setOff(uint i) { off = i; }
+  void setOff(uint i) { off = i;    }
   
   uint size() const { return lmers.size(); }
   
-  Family* getFamily() const { return family; }
-  void setFamily(Family *fam) { family = fam; }
+  Family* getFamily() const   { return family; }
+  void setFamily(Family *fam) {
+    if (family) {
+      prevFamily = family;
+    }
+    family = fam;
+  }
   
-  friend std::ostream &operator<<( std::ostream &output, const LmerVector &L);
+  Family* getPrevFamily() const   { return prevFamily; }
+  void setPrevFamily(Family *fam) { prevFamily = fam;  }
   
-  std::vector<uint> lmers;
-  Family* family;
+  friend ostream &operator<<( ostream &output, const LmerVector &L) {
+    output  << "(\tFront:\t" << L.front()
+    << "\tBack:\t" << L.back()
+    << "\tSize:\t" << L.size()
+    << "\tOffset In Family\t" << L.getOff()
+    << "\tFamily:\t" << L.getFamily() << "\t)";
+    return output;
+  }
+  
+  vector<uint> lmers;
   uint previous;
-  uint position;
   uint off;
+  Family* family;
+  Family* prevFamily;
 };
 
 #endif //LMER_VECTOR_H

### 2015-07-28 RAIDER_src_v2_options_new/src/LmerVector2.h

@@ -1,0 +1,48 @@
+#ifndef LMER_VECTOR_H
+#define LMER_VECTOR_H
+#include <vector>
+#include <stdexcept>
+#include <iostream>
+
+#ifndef uint
+typedef unsigned int uint;
+#endif
+
+class Family;
+
+class LmerVector {
+public:
+  
+  LmerVector(uint position);
+  ~LmerVector() { }
+  
+  void push_back(uint val);
+  
+  uint operator [](uint i) const { return lmers[i]; }
+  uint & operator [](uint i) { return lmers[i]; }
+  
+  uint prev() const { return previous; }
+  uint back() const { return lmers.back(); }
+  uint front() const { return lmers.front(); }
+  
+  uint getPosition() const { return position; }
+  void setPosition(uint i) { position = i; }
+  
+  uint getOff() const { return off; }
+  void setOff(uint i) { off = i; }
+  
+  uint size() const { return lmers.size(); }
+  
+  Family* getFamily() const { return family; }
+  void setFamily(Family *fam) { family = fam; }
+  
+  friend std::ostream &operator<<( std::ostream &output, const LmerVector &L);
+  
+  std::vector<uint> lmers;
+  Family* family;
+  uint previous;
+  uint position;
+  uint off;
+};
+
+#endif //LMER_VECTOR_H

### 2015-07-28 RAIDER_src_v2_options_new/src/main.cpp

@@ -52,13 +52,17 @@ struct AppOptions {
   bool overlaps;
   // Whether or not to allow for modified tie up
   bool tieup;
+  // Do cleanup while going
+  bool proactive_split;
+  // Keep track of previous family
+  bool prev_fam;
   
   seqan::CharString seed;
   seqan::CharString sequence_file;
   seqan::CharString output_directory;
   
   AppOptions() :
-  verbosity(1), age(1), family_array(true), excising(false), overlaps(true), tieup(false){}
+  verbosity(1), age(1), family_array(true), excising(false), overlaps(true), tieup(false), proactive(false), prevFam(false){}
 };
 
 // ==========================================================================
@@ -111,7 +115,8 @@ seqan::ArgumentParser::ParseResult parseCommandLine(AppOptions & options, int ar
   addOption(parser, seqan::ArgParseOption("e", "excise", "Enable excising (disabled by default)."));
   addOption(parser, seqan::ArgParseOption("no", "nooverlaps", "Do not require overlaps (required by default)."));
   addOption(parser, seqan::ArgParseOption("t", "tieup", "Enable alternate tie up (disabled by default)."));
-
+  addOption(parser, seqan::ArgParseOption("ps", "prosplit", "Enable proactive splitting(disabled by default)."));
+  addOption(parser, seqan::ArgParseOption("pf", "prevfam", "Enable pointers to prev family (disabled by default)."));
   
   // Add Examples Section.
   addTextSection(parser, "Examples");
@@ -176,6 +181,14 @@ seqan::ArgumentParser::ParseResult parseCommandLine(AppOptions & options, int ar
   
   if (isSet(parser, "tieup")){
     options.tieup = true;
+  }
+  
+  if (isSet(parser, "prosplit")){
+    options.proactive_split = true;
+  }
+  
+  if (isSet(parser, "prevfam")){
+    options.prev_fam = true;
   }
   
   // Ensure a trailing /
@@ -403,7 +416,7 @@ int main(int argc, char const ** argv) {
   seeds.push_back(options.seed);
   
 
-  getElementaryFamilies(sequence, seeds, families, options.verbosity, options.family_array, options.excising, options.overlaps, options.tieup);
+  getElementaryFamilies(sequence, seeds, families, options);
   
   if (options.verbosity > 0) {
     cout << "Writing results elements..." << endl;

### 2015-07-28 RAIDER_eval.py
faster family/lmervector?

@@ -180,7 +180,7 @@ def parse_params(args):
     raider2_argument.add_argument('--na', '--no_family_array', dest="family_array", action="store_false", help="Disable family array in Raider2", default=True)
     raider2_argument.add_argument('--ex', '--excise', dest="excising", action="store_true", help="Enable excising in RAIDER2", default=False)
     raider2_argument.add_argument('--no', '--no_overlaps', dest="overlaps", action="store_false", help="Do not require overlaps in RAIDER2", default=True)
-    raider2_argument.add_argument('--tu', '--tie_up', dest="tieup", action="store_true", "Enable alternative tie ups", default=False)
+    raider2_argument.add_argument('--tu', '--tie_up', dest="tieup", action="store_true", help="Enable alternative tie ups", default=False)
 
     # REPSCOUT ARGUMENTS
     repscout_argument = parser.add_argument_group("REPSCOUT parameters")
@@ -456,12 +456,16 @@ def run_composites_finder(elements_file, seq_file, compositesFinderDir):
 
     return p
 
-def run_raider2(seed, seed_num, f, m, input_file, raider2_dir, age):
+def run_raider2(seed, seed_num, f, m, input_file, raider2_dir, family_array, excise, overlaps, tieup, age, age_only):
     """Given raider parameters and an input file, run RAIDER and put the output into
     the directory specified in output_dir (creating a random name is none is
     specified."""
 
     input_base = file_base(input_file).rstrip(".fa")
+    raider2_dir += "NO_FA." if not family_array else "FA."
+    raider2_dir += "EXC." if excise else "NO_EXC."
+    raider2_dir += "NO_OV." if not overlaps else "OV."
+    raider2_dir += "TU" if tieup else "NO_TU"
     output_dir = raider2_dir + "/" + input_base.upper() + ".s" + str(seed_num)
     if not os.path.exists(output_dir):
         os.makedirs(output_dir)
@@ -471,9 +475,17 @@ def run_raider2(seed, seed_num, f, m, input_file, raider2_dir, age):
         seed_string = "-s " + " -s ".join(seed)
     else:
         seed_string = "-s {seed}".format(seed = seed)
-    
-    
-    cmd1 = "{raider2} -q -c {f} --age {version} {min_arg} {seed} {input_file} {output_dir}".format(raider2 = Locations['raider2'], f = f, version = age, min_arg = min_arg, seed = seed_string, input_file = input_file, output_dir = output_dir)
+
+    opt_str = ""
+    if not age_only:
[…]

### 2015-07-28 RAIDER_src_v2_options_new/src/Family.h

@@ -1,52 +1,158 @@
-#ifndef FAMILY_H
-#define FAMILY_H
+#ifndef SCANER_FAMILY
+#define SCANER_FAMILY
 #include <vector>
 #include "LmerVector.h"
 #include <iostream>
 #include <assert.h>
 
+using namespace std;
+
 class Family {
 public:
-  void adopt(LmerVector *v);
-  void adopt(LmerVector *v, uint L);
 
-  uint size() const;
+	void adopt(LmerVector *v) {
+		v->setFamily(this);
+		vectors.push_back(v);
+		setLast(v);
+		setExpectedEnd(v->back() + 1);
+	}
 
-  uint getOffset() const { return offset; }
-  void setOffset(uint off){ offset = off; }
-  void addToOffset(uint off);
+    void adopt(LmerVector *v, uint L){
+        v->setFamily(this);
+        //cout << L << endl;
+        if (vectors.size() > 0){
+            uint off = v->front() - vectors.back()->front();
+            addToOffset(off - 1);
+            setRepeatLength(getRepeatLength() + off);
+            v->setOff(off - 1);
+        }
+        else{
+            setRepeatLength(L);
+            setOffset(0);
+            v->setOff(0);
+        }
+        vectors.push_back(v);
+        setLast(v);
+        setExpectedEnd(v->back() + L);
+    }
 
-  uint getRepeatLength() const { return repeatlength; }
-  void setRepeatLength(uint L){ repeatlength = L; }
-  uint repeatLength(uint L) const;
+	uint size() const {
+		if (vectors.size() == 0) {
+			return 0;
+		}
+		return vectors.front()->size();
+	}
 
-  std::vector<LmerVector*>* getLmers() { return &vectors; }
+    void addToOffset(uint off){
+        offset += off;
+    }
 
-  LmerVector* at(uint index) { return vectors.at(index); }
-  LmerVector* see(uint index) const { return vectors.at(index); }
-  void push_back(LmerVector* v) { vectors.push_back(v); }
-  LmerVector* findByPos(uint pos);
+    void setOffset(uint off){
+        offset = off;
+    }
+
+    uint getOffset() const {
+        return offset;
+    }
+
+    uint getRepeatLength() const {
+        return repeatlength;
+    }
+
+    void setRepeatLength(uint L){
+        repeatlength = L;
+    }
+
+	uint repeatLength(uint L) const {
+		return repeatlength + L - L;
[…]

### 2015-07-28 RAIDER_src_v2_options_new/src/LmerVector.h

@@ -1,8 +1,10 @@
-#ifndef LMER_VECTOR_H
-#define LMER_VECTOR_H
+#ifndef LMER_VECTOR
+#define LMER_VECTOR
 #include <vector>
 #include <stdexcept>
 #include <iostream>
+
+using namespace std;
 
 #ifndef uint
 typedef unsigned int uint;
@@ -12,37 +14,71 @@ class Family;
 
 class LmerVector {
 public:
-  
-  LmerVector(uint position);
-  ~LmerVector() { }
-  
-  void push_back(uint val);
-  
-  uint operator [](uint i) const { return lmers[i]; }
-  uint & operator [](uint i) { return lmers[i]; }
-  
-  uint prev() const { return previous; }
-  uint back() const { return lmers.back(); }
-  uint front() const { return lmers.front(); }
-  
-  uint getPosition() const { return position; }
-  void setPosition(uint i) { position = i; }
-  
-  uint getOff() const { return off; }
-  void setOff(uint i) { off = i; }
-  
-  uint size() const { return lmers.size(); }
-  
-  Family* getFamily() const { return family; }
-  void setFamily(Family *fam) { family = fam; }
-  
-  friend std::ostream &operator<<( std::ostream &output, const LmerVector &L);
-  
-  std::vector<uint> lmers;
-  Family* family;
-  uint previous;
-  uint position;
-  uint off;
+
+	LmerVector(uint position) {
+		family = nullptr;
+		push_back(position);
+	}
+    
+    ~LmerVector() {
+    }
+
+	void push_back(uint val) {
+        if(lmers.size() > 0){
+            previous = back();
+        }
+		lmers.push_back(val);
+	}
+
+	uint operator [](uint i) const {
+		return lmers[i];
+	}
+
+	uint & operator [](uint i) {
+		return lmers[i];
+	}
+
+    uint prev() const {
+        return previous;
+    }
+
+	uint back() const {
+		return lmers.back();
+	}
+
+	uint front() const {
+		return lmers.front();
+	}
+
+    uint getOff() const {
+        return off;
+    }
+    
+    void setOff(uint i) {
+        off = i;
+    }
+
+	uint size() const {
+		return lmers.size();
+	}
+
+	Family* getFamily() const {
+		return family;
+	}
+
+	void setFamily(Family *fam) {
+		family = fam;
+	}
+
[…]

### 2015-07-28 RAIDER_src_v2_options_new/src/Family.h
family file that compiles

@@ -110,7 +110,7 @@ public:
   void removeOne(LmerVector* v){
     std::vector<LmerVector*>::iterator start = std::find(vectors.begin(), vectors.end(), v);
     (*(start+1))->setOff((*(start+1))->front() - (*(start-1))->front());
-    vectors.erase(start, start+1, vectors.end()); // no longer part of backbone of this family
+    vectors.erase(start, start+1); // no longer part of backbone of this family
     vectors.shrink_to_fit();
   }
 
@@ -135,4 +135,3 @@ public:
 };
 
 #endif //FAMILY_H
->>>>>>> b2cb1aabbca0804265664c37b15c9782a2bf08ba

### 2015-07-28 RAIDER_eval.py
raider eval update

@@ -181,6 +181,8 @@ def parse_params(args):
     raider2_argument.add_argument('--ex', '--excise', dest="excising", action="store_true", help="Enable excising in RAIDER2", default=False)
     raider2_argument.add_argument('--no', '--no_overlaps', dest="overlaps", action="store_false", help="Do not require overlaps in RAIDER2", default=True)
     raider2_argument.add_argument('--tu', '--tie_up', dest="tieup", action="store_true", help="Enable alternative tie ups", default=False)
+    raider2_argument.add_argument('--ps', '--prosplit', dest="prosplit", action="store_true", help="Enable proactive splitting(disabled by default).", default=False)
+    raider2_argument.add_argument("--pf", "--prevfam", dest="prevfam", action="store_true", help="Enable pointers to prev family (disabled by default).", default=False)
 
     # REPSCOUT ARGUMENTS
     repscout_argument = parser.add_argument_group("REPSCOUT parameters")
@@ -456,16 +458,18 @@ def run_composites_finder(elements_file, seq_file, compositesFinderDir):
 
     return p
 
-def run_raider2(seed, seed_num, f, m, input_file, raider2_dir, family_array, excise, overlaps, tieup, age, age_only):
+def run_raider2(seed, seed_num, f, m, input_file, raider2_dir, family_array, excise, overlaps, tieup, prosplit, prevfam, age, age_only):
     """Given raider parameters and an input file, run RAIDER and put the output into
     the directory specified in output_dir (creating a random name is none is
     specified."""
 
     input_base = file_base(input_file).rstrip(".fa")
-    raider2_dir += "NO_FA." if not family_array else "FA."
-    raider2_dir += "EXC." if excise else "NO_EXC."
-    raider2_dir += "NO_OV." if not overlaps else "OV."
-    raider2_dir += "TU" if tieup else "NO_TU"
+    #raider2_dir += "NO_FA." if not family_array else "FA."
+    #raider2_dir += "EXC." if excise else "NO_EXC."
+    #raider2_dir += "NO_OV." if not overlaps else "OV."
+    #raider2_dir += "TU." if tieup else "NO_TU."
+    #raider2_dir += "PS" if prosplit else "NO_PS."
+    #raider2_dir += "PF" if prevfam else "NO_PF"
     output_dir = raider2_dir + "/" + input_base.upper() + ".s" + str(seed_num)
     if not os.path.exists(output_dir):
         os.makedirs(output_dir)
[…]

### 2015-07-28 RAIDER_src_v2_options_new/src/main.cpp
Updating main.cpp: Now generating (but not yet using) the skip back list

@@ -5,6 +5,7 @@
 // Based off of original RAIDER by Nathan Figueroa
 // ==========================================================================
 
+
 #include <seqan/basic.h>
 #include <seqan/sequence.h>
 #include <seqan/arg_parse.h>
@@ -16,6 +17,8 @@
 #include <stdlib.h>
 #include <stdio.h>
 #include <unordered_map>
+#include <set>
+#include <numeric>
 
 #include "SeedChain.h"
 #include "Family.h"
@@ -83,6 +86,7 @@ seqan::ArgumentParser::ParseResult parseCommandLine(AppOptions & options, int ar
   addOption(parser, seqan::ArgParseOption("t", "tieup", "Enable alternate tie up (disabled by default)."));
   addOption(parser, seqan::ArgParseOption("ps", "prosplit", "Enable proactive splitting(disabled by default)."));
   addOption(parser, seqan::ArgParseOption("pf", "prevfam", "Enable pointers to prev family (disabled by default)."));
+  addOption(parser, seqan::ArgParseOption("sbl", "skipbacklist", "Enable skip back list (disabled by default)."));
   
   // Add Examples Section.
   addTextSection(parser, "Examples");
@@ -155,6 +159,10 @@ seqan::ArgumentParser::ParseResult parseCommandLine(AppOptions & options, int ar
   
   if (isSet(parser, "prevfam")){
     options.prev_fam = true;
+  }
+
+  if (isSet(parser, "sbl")){
+    options.sbl = true;
   }
   
   // Ensure a trailing /
@@ -360,6 +368,46 @@ void writeResults(vector<Family*> &families, vector<idThreshold> &thresholds, Ap
 }
 
 
+// Create the skip-back list for the seed.  
+vector<int> createSBL(vector<seqan::CharString> seeds, bool sbl) {
+  int l = accumulate(seeds.cbegin(), seeds.cend(), 0, [](int x, seqan::CharString s) {return max(x, (int)length(s));});
+  set<int> S({1, l+1});
+
+  if (not sbl) {
+    vector<int> SBL(l+1);
+    iota(SBL.begin(), SBL.end(), 1);
+    return SBL;
+  }
+
+  set<int> V;
+
+
+
+  for (auto i=seeds.begin(); i!=seeds.end(); i++)
+    for (int j=0; j < (int)length(*i); j++)
+      if ((*i)[j]=='0')
+	V.insert(j);
+
+
[…]

### 2015-07-28 RAIDER_src_v2_options_new/src/main.cpp
Fixing bug in main.cpp

@@ -428,11 +428,8 @@ int main(int argc, char const ** argv) {
   vector<Family*> families;
   vector<seqan::CharString> seeds;
   seeds.push_back(options.seed);
-  seeds.push_back("1011111");
-
 
   vector<int> SBL = createSBL(seeds, options.sbl);
-
 
   getElementaryFamilies(sequence, seeds, families, options);
   

### 2015-07-28 manuscript/BMC/bmc_article.tex

@@
 %\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
+
+% Added by karro for commenting -- should be removed before submission.
+\usepackage{color}
+\newcommand{\red}[1]{{\color{red}#1}}
+\newcommand{\blue}[1]{{\color{blue}#1}}
+\newcommand{\cyan}[1]{{\color{cyan}#1}}
+\newcommand{\magenta}[1]{{\color{magenta}#1}}
 
@@
 sequence variation.  Here we present phRAIDER (Pattern Hunter based
-Rapid Ab Initio Detection of Elementary Repeats) for quickly masking a
-newly sequence, or poorly annotated, genome of transposable elements
-that is robust to the sequence variation present in real data.
+Rapid Ab Initio Detection of Elementary Repeats), a tool that can
+quickly identify and mask tranposable elements in a a newly sequenced
+genome and is robust to the sequence variation present in real data.
 
-\parttitle{Results}
+\parttitle{Results} \red{To be written later.}
 
-\parttitle{Conclusions}
+\parttitle{Conclusions} \red{To be written later.}
 
@@
 order genome, covering as much as 45\% of the human genome and 90\% of
-the maize genome [CHECK THIS] \cite{Venter:2001p92} [CITE MAIZE].  TEs
-have proved an importent sorce of data in a number of studies
+the maize genome \cite{Venter:2001p92,SanMiguel:1996wa}.  TEs have
+proved an important source of data in a number of studies
 \cite{Arndt:2005p279,Karro:2008p123,Mugal:2009p581,Hardison:2003p97},
 but given their prevalence, it is just important for those studying
-the genome in them to have them masked out -- their bases replaced by
-N to allow for easy identification and filtering.  For example. if not
-filtered, TEs can trigger huge numbers of false positives automated
-gene finding tool \cite{Jiang:2013jt}, as well as inflate tool
-runtime.
+the genome to have TEs masked out -- their bases replaced by N to
+allow for easy identification and filtering.  Failure to filter can
+reek havoc with genomic analysis tools.  For example. if not filtered,
+TEs can trigger huge numbers of false positives automated gene finding
+tool \cite{Jiang:2013jt}, as well as inflate tool runtime.
 
-The most well known tools for repeat identification are RepeatMasker
-and nHMMER \cite{RepeatMaskerOpen:XkNxXSd7,Wheeler:2013gj}, both of
-which work on the principle of a library-based search.  That is, to
-identify transposable elements to a known family, the tools require
-some description of sequences in the family (e.g. a ancestral sequence
-for BLASTing, or a profile HMM description) that is used as the basis
-for identification.  But, much like we ask how the snow plow driver
-gets to work, or how dictionary authors look up the spelling of a word
-\cite{Pratchett:uw}, we must ask how the libraries on which the tools are based are
-initially compiled.  Such tools are useful for the discovery of new
-families, and cannot be applied to a newly sequenced genome for which
-such a library has not been compiled.
+The most well known tools for repeat identification is RepeatMasker,
+with the new nHMMER showing promise as well)
+\cite{RepeatMaskerOpen:XkNxXSd7,Wheeler:2013gj}.  Both work on the
+principle of a library-based search.  That is, to identify
+transposable elements to a known family, the tools require some
+description of sequences in the family (e.g. a ancestral sequence for
+BLASTing, or a profile HMM description) that is used as the basis for
+identification.  But, much like we ask how the snow plow driver gets
+to work, or how the author of a dictionary authors look up the spelling of a word
+\cite{Pratchett:uw}, we must ask how the libraries on which the tools
+are based are initially compiled.  Library-based tools are useless for the
+discovery of new families, and cannot be applied to a newly sequenced
+genome for which such a library has not been compiled.
 
-Within mammalian species we can, to some extent, rely on homology
+Within mammalian species we can largely rely on homology
 relationships to port libraries across species.  A library description
@@
 computationally intensive and infeasible for use on whole genomes
-(requiring 60 hours for 18Mb rice genome int 2013 \cite{Jiang:2013jt}), while
+(requiring 60 hours for 18Mb rice genome int 2013), while
 PILER achieves a good runtime with very low sensitivity
-\cite{Jiang:2013jt}.  ReAS \cite{Li:2005p3138} and RepeatScout
+\cite{Jiang:2013jt}.  ReAS \cite{Li:2005he} and RepeatScout
 \cite{Price:2005p1247} are based on $k$-mer searches, with the
 earlier showing less sensitivity than RECON \cite{Jiang:2013jt}.
-Finally RepeatGluer \cite{Pevzner:2004p3157,Zhi:2006p3199} is
+RepeatGluer \cite{Pevzner:2004p3157,Zhi:2006p3199} is
 based on a variation of DeBrujin graphs, which allows for a
@@
 
-[TO DO: Make sure we mention somewhere that RAIDER should work for
-  unassembled genomes.]
-
 \subsection*{Elementary Repeats}
@@
 these building blocks are sufficient for the purpose
-of masking, and can be assembled into Transposible Elements for those
-interested in studying those elements.
+of masking, and can be assembled into Transposable Elements for those
+interested in TEs themselves.
 
-While it is noteriously difficult to state a mathematical definition
-of elementary repeats \cite{Bao:2002}, elemenary repeats are a bit
-easier to model.  For a given genome, a nucleotide sequence $r$ is an
-elementary repeat if:
+While it is notoriously difficult to mathematically model transposable
+elements \cite{Bao:2002}, elementary repeats are more conducive to a
+formal description.  For a given genome, a nucleotide sequence $r$ is
+an elementary repeat if:
 \begin{enumerate}
@@
 \end{enumerate}
-The first two criteria are straight-foward, allowing us to designate a
-threshold length and frequency for our model.  The minimality
+The first two criteria are straight-forward, allowing us to designate
+length and frequency thresholds for our model.  The minimality
 requirement ensures that an elementary repeat does not contain
@@
 \cite{He:2006gpa,Huo:2009hoa} based on variations of suffix tree
-approaches, but because of that use of suffix trees they are limited
+approaches, but are limited
 in their ability to handle sequence variation.  As transposable
@@
   either an elementary repeat or belongs to one.
-\item Any two $l$-ers belonging to an elementary repeat must appear
+\item Any two $l$-mers belonging to an elementary repeat must appear
   the same number of times in the genome. 
-\item Given two sequences in the genome that are {\it maximally
+\item If two sequences in the genome that are {\it maximally
   identical} (that is, cannot be extended in either direction and still
@@
 \end{enumerate}
-\end{observation}
 By ``belong'' we mean ``is a substring of''; the reason for the
@@
 Results of the preliminary implementation were promising.  Based on
-the Saha {\it et al} analysis putting RepeatMasker as the top {\it de
-  novo} search tool, comparisons were against that.  RAIDER result
+the Saha {\it et al.} analysis putting RepeatScout as the top {\it de
+  novo} search tool, comparisons were made against that tool.  RAIDER result
 quality was close, but slightly behind -- with runtime
@@
 $12\times$ speedup from RepeatScout to RAIDER (2344 seconds to 192 seconds), while coverage
-of the RepBase ancesteral sequecne \cite{Jurka:2005bl} actually improved (77\% to 84\%), while on mouse chromosome 19 we saw 
+of the RepBase  \cite{Jurka:2005bl} ancestral sequence actually improved (77\% to 84\%), while on mouse chromosome 19 we saw 
 the same speedup with a significant drop in coverage (53\% to 30\%).  On
@@
 might or might not match).  In the toy seed $s=110101$, we would say the
-strings $AACACA$ and $AAGACA$ because all positions corresponding to a
-1 match.  The string $AACACA$ does not math $TACACA$ (with respect to
+strings $AACACA$ and $AAGACA$ match because all miss-matched positions correspond to a 
+0 match.  The string $AACACA$ does not math $TACACA$ (with respect to
 $s$), as the single miss-match does not correspond to the 1.  Nor does
 $AACACA$ match any substring of $TTAACTCA$, though it does match a
-substring of $T{\mathbf AAGATA}C$.
+substring of $T\mathbf{AAGATA}C$.
 
@@
 improvements at virtually no cost in runtime \cite{Li:2004wl}.
-SpecificallyIn our context, we will use them to allow limited
+In our context, we will use them to allow limited
 miss-matches between instances of ``the same'' elementary repeat.  The
 preliminary version of RAIDER from Figueroa {\it et al.} made some
-attempt to incorporate this -- was successful enough to provide a
+attempt to incorporate this.  It was successful enough to provide a
 proof-of-concept, but a significant amount of work was left to be
-done.
+done to allow for true space seed usage.
 
-In this paper we present phRAIDER (PatternHunter-based RAIDER),
-implemented as open source code that may be obtained [NEED WEB
-  ADDRESS].  In our Methods section we present the new model that
-allows us to extend RAIDER to correctly use spaced seeds, and briefly
-outline the algorithm (with more details provided in the supplmentary
-materials), with a analsis of phRAIDER performance in our Results.
+In this paper we present phRAIDER (PatternHunter-based RAIDER), a fast
+tool for the identification and making of transposible elements in
+both assembled and unassembled genomes. Code  may be obtained [NEED WEB ADDRESS].  In our Methods section
+we present the new model that allows us to extend RAIDER to correctly
+use spaced seeds, and briefly outline the algorithm (with more details
+provided in the supplementary materials), with a analysis of phRAIDER
+performance in our Results.
 
@@
     symbols (indicating that a position where base content does not
-    matter).  A string matches a sequence decriptor of equal length if
+    matter).  A string matches a sequence descriptor of equal length if
     it has the same base in all non-* positions.
@@
     hits both $AA*CC$ and $TTTTT$, but not $A*TTT$.)
-  \item The {\it decomposition} of $r$ (with respect to $s$) is the set of all substrings of
-    $r$ that are hit by $s$.  For the {\it generalized decomposition},
-    we take each member of the decomposition and replace by $*$ any
-    base that aligned to a 0 in the seed.  (Example: if $s=11011$ and
-    $r=AAATTT$, then $AAATT$ is in the decomposition, so $AA*TT$ is in
-    the generalized decomposition.)
+  \item The {\it decomposition} of $r$ (with respect to seed $s$) is
+    the set of all substrings of $r$ that are hit by $s$.  For the
+    {\it generalized decomposition}, we take each member of the
+    decomposition and replace by $*$ any base that aligned to a 0 in
+    the seed.  (Example: if $s=11011$ and $r=AAATTT$, then $AAATT$ is
+    in the decomposition, so $AA*TT$ is in the generalized
+    decomposition.)
   \item We say a string $s$ {\it covers} a sequence descriptor $r$ if
@@
 \end{enumerate}
-
-The Zheng and Lonardi elementary repeats are all identical, hence can
-be described by a single string.  Because we allow variation, we
-required the flexibility of the sequences descriptor.  Leading us to
-our definition:
 
@@
 sequence is such that (4) holds as well, this is an elementary repeat
-descriptor for this family of sequences.  (Note $AA*AA*TT*TT$ -- the
-elementary repeat descriptor is not unique.)  But if we add a a second
+descriptor for this family of sequences.  (Note $AA*AA*TT*TT$ also qualifies -- the
+elementary repeat descriptor is not necessarily unique.)  But if we add a second
 instance of $g_3 = AAAAAGTTTTT$, something happens that was not
@@
 one base to the left, as the $AAAAA$ substring would cause a violation
-of condition (3)).  So $r_2$ corresponds to strings in $G$ that are
+of condition (3)).  So $r_2$ matches strings in $G$ that are
 substrings of those that $r_1$ matches -- a condition not possible
@@
   matches a member of the generalized decomposition of $r$ (w.r.t. to
-  $s$.)
+  $s$).
 \end{definition}
@@
 repeats.  But the string $AAAAT$ does not match any string in the
-generalized decomposition of $r$, and hence was free to belong to
-the descriptor of a ``distinct'' family.
+generalized decomposition of $r$, and hence was free to belong to the
+descriptor of a ``distinct'' family.  See the Supplementary Materials
+for a proof of all the core observations.
 
 \subsection*{phRAIDER Algorithm}
+
+\red{
+TO DO: Give a pseudocode description of the algoirthm and some minimal
+discussion.
+}
+
+\subsection*{Analysis}
+
+\red{
+TO DO: Describe pra analysis and quality metrics.
+}
+
+\section*{Results}
+
+\red{
+Points to be made:
+\begin{enumerate}
+\item Show: RAIDER2 has better quality masking results then RS.
+\item Show: RAIDER2 is better than RAIDER.
+\item Show: Seeds matter.
+\item Show: Runtime / memory usage
+\end{enumerate}
+
+Possible tables / plots:
+\begin{itemize}
+\item Table showing results on different chromosomes (pra, runtime, memory).
+\item Plot of runtime against genome size (simulated data to control size)?
+\item Plot of memory usage (simlated data to control size)?
+\item We might want to compare the ability to detect transposable
+  elements againsts other types of repeats.  (Should be easy to do
+  using simulator.)
+\end{itemize}
+}
+
+\section*{Conclusions}
+
+\red {
+Hopefully will beconcluding that RAIDER is a better tool than
+RepeatScout.  Posible discussion of tool features.
+}
 
@@
 %%  LocalWords:  TTAACTCA TTTTT TTT substrings AAATTT AAATT TT AAAAA
-%%  LocalWords:  FIGUREOA AAAAAGTTTTT AAAAGTTTT AAAA TTTT AAAAT
-%%  LocalWords:  Acknowledgements
+%%  LocalWords:  FIGUREOA AAAAAGTTTTT AAAAGTTTT AAAA TTTT AAAAT novo
+%%  LocalWords:  Acknowledgements transposable tranposable runtime
+%%  LocalWords:  unassembled RepBase AAGATA

### 2015-07-29 RAIDER_src_v2_options_new/src/Family.h
New approach to skipped

@@ -74,45 +74,66 @@ public:
     last_index = v->back();
   }
   
-  LmerVector* getOneAfterLast(){
-    std::vector<LmerVector*>::iterator start = std::find(vectors.begin(), vectors.end(), last);
-    return *(start+1);
-  }
+  //LmerVector* getOneAfterLast(){
+  //  std::vector<LmerVector*>::iterator start = std::find(vectors.begin(), vectors.end(), last);
+  //  return *(start+1);
+  //}
   
   uint getExpectedEnd() const { return expected_end; }
   void setExpectedEnd(uint expected) { expected_end = expected; }
   
-  std::vector<LmerVector*>* getSkipped() { return &skipped; }
-
+  void resetOffs(){
+    for (uint i = 0; i < vectors.size(); i++){
+      if (v != vectors.front()){
+        v->setOff(v->front() - vectors.at(i-1)->front());
+      }
+    }
+  }
+  
   std::vector<LmerVector*> popSkipped() {
-    std::vector<LmerVector*> ret(skipped);
-    skipped.clear();
-    return ret;
+    //std::vector<LmerVector*> ret(skipped);
+    //skipped.clear();
+    std::vector<LmerVector*> skipped(vectors.size());
+    std::remove_copy_if (vectors.begin(),vectors.end(),skipped.begin(), [](LmerVector* v){return v->gotSkipped()});
+    vectors.shrink_to_fit();
+    skipped.shrink_to_fit();  // shrink container to new size
+    resetOffs();
+    return skipped;
+  }
+  
+
+  
+  void setSkippedRange(LmerVector* lastSeen){
+    std::vector<LmerVector*>::iterator start = std::find(vectors.begin(), vectors.end(), last);
+    std::vector<LmerVector*>::iterator end = std::find(vectors.begin(), vectors.end(), lastSeen);
+    std::for_each(start+1, end, setSkipped);
   }
   
   void setRemainingSkipped(){
-    std::vector<LmerVector*>::iterator start = std::find(vectors.begin(), vectors.end(), last);
-    std::move(start+1, vectors.end(), std::back_inserter(skipped));
-    vectors.erase(start+1, vectors.end()); // no longer part of backbone of this family
-    vectors.shrink_to_fit();
+    setSkippedRange(vectors.back());
   }
+  //  std::vector<LmerVector*>::iterator start = std::find(vectors.begin(), vectors.end(), last);
[…]

### 2015-07-29 RAIDER_src_v2_options_new/src/SeedChain.h

@@ -351,10 +351,10 @@ bool fragmentSplit(LmerVector* v, uint L, vector<Family*> &families, AppOptions 
   // If v occurs after it is supposed to (over L past the last index of the family),
   // split the family into prefix...last, (last+1)...suffix
   else if (!closeEnough(fam->getLast(), v, L, options.overlaps)) {
-    if (options.prev_fam && v->getPrevFamily() && closeEnough(v->getPrevFamily()->getLast(), v , L, options.overlaps)){
-      fam->removeOne(v);
-      v->getPrevFamily()->adopt(v,L);
-    }
+    // if (options.prev_fam && v->getPrevFamily() && closeEnough(v->getPrevFamily()->getLast(), v , L, options.overlaps)){
+    //   fam->removeOne(v);
+    //   v->getPrevFamily()->adopt(v,L);
+    // }
     else {
       Family* newFam = splitRepeatsByLmer(fam, fam->getLast(), true, L, options);
       families.push_back(newFam);
@@ -490,7 +490,7 @@ void getElementaryFamilies(seqan::Dna5String &sequence, vector<seqan::CharString
           if (!canMergeSkipped(fam, L, options.overlaps)){
             families.push_back(mergeIntoFamily(fam->popSkipped(), L));
           }
-          fam->moveSkippedRange(v);
+          fam->setSkippedRange(v);
         }
         fam->setLast(v);
         if (options.proactive_split && fam->lastRepeatComplete() && fam->getSkipped()->size() > 0){

### 2015-07-30 RAIDER_src_v2_options_new/src/SeedChain.h
Updating SeedChain.h

@@ -40,14 +40,15 @@ struct AppOptions {
   bool prev_fam;
   // Do we use a skip-back list?
   bool sbl;
-  // 
+  // Trun on masking
+  bool mask;
   
   seqan::CharString seed;
   seqan::CharString sequence_file;
   seqan::CharString output_directory;
   
   AppOptions() :
-  age(1), verbosity(1), family_array(true), excising(false), overlaps(true), tieup(false), proactive_split(false), prev_fam(false), sbl(false){}
+  age(1), verbosity(1), family_array(true), excising(false), overlaps(true), tieup(false), proactive_split(false), prev_fam(false), sbl(false), mask(false) {}
 };
 
 typedef unordered_map<size_t, LmerVector*> LmerMap;

### 2015-07-30 RAIDER_src_v2_options_new/src/SeedChain.h
New approach to skipping

@@ -355,7 +355,7 @@ bool fragmentSplit(LmerVector* v, uint L, vector<Family*> &families, AppOptions 
     //   fam->removeOne(v);
     //   v->getPrevFamily()->adopt(v,L);
     // }
-    else {
+    //else {
       Family* newFam = splitRepeatsByLmer(fam, fam->getLast(), true, L, options);
       families.push_back(newFam);
       // If splitting proactively, we know (last+1)...(v-1) did not
@@ -364,7 +364,7 @@ bool fragmentSplit(LmerVector* v, uint L, vector<Family*> &families, AppOptions 
         families.push_back(newFam2);
       }
       return true;
-    }
+    //}
   }
   return false;
 }
@@ -385,15 +385,15 @@ bool isPrefix(LmerVector *v) {
 }
 
 bool canMergeSkipped(Family* fam, uint L, bool overlaps) {
-  if (fam->getSkipped()->size() > 0){
-    return closeEnough(fam->getSkipped()->back(), fam->getOneAfterLast(), L, overlaps);
+  if (fam->getLastSkipped() && fam->getOneAfterLast()){
+    return closeEnough(fam->getLastSkipped(), fam->getOneAfterLast(), L, overlaps);
   }
   return true;
 }
 
-Family* mergeIntoFamily(vector<LmerVector*> lmers, uint L){
+Family* mergeIntoFamily(Family* fam, uint L){
   Family* newFam = new Family();
-  for (LmerVector* v : lmers){
+  for (LmerVector* v : fam->popSkipped()){
     newFam->adopt(v,L);
   }
   return newFam;
@@ -406,14 +406,15 @@ void tieLooseEnds(vector<Family*> &families, uint L, AppOptions options) {
     
     if (options.tieup){
       if (!canMergeSkipped(fam, L, options.overlaps)){
-        families.push_back(mergeIntoFamily(fam->popSkipped(), L));
+         
+        families.push_back(mergeIntoFamily(fam, L));
       }
     }
     if (!fam->lastRepeatComplete()) {
       Family* newFam;
       if (options.tieup){
         fam->setRemainingSkipped();
-        newFam = mergeIntoFamily(fam->popSkipped(), L);
+        newFam = mergeIntoFamily(fam, L);
       }
       else{
         newFam = splitRepeatsByLmer(fam, fam->getLast(), true, L, options);
@@ -488,13 +489,13 @@ void getElementaryFamilies(seqan::Dna5String &sequence, vector<seqan::CharString
[…]

### 2015-07-31 RAIDER_src_v2_options_new/src/SeedChain.h
main.cpp / SeedChain.h: Adding mask_file option

@@ -40,15 +40,15 @@ struct AppOptions {
   bool prev_fam;
   // Do we use a skip-back list?
   bool sbl;
-  // Trun on masking
-  bool mask;
+  // File name for masked file (if specified)
+  seqan::CharString mask_file;
   
   seqan::CharString seed;
   seqan::CharString sequence_file;
   seqan::CharString output_directory;
   
   AppOptions() :
-  age(1), verbosity(1), family_array(true), excising(false), overlaps(true), tieup(false), proactive_split(false), prev_fam(false), sbl(false), mask(false) {}
+  age(1), verbosity(1), family_array(true), excising(false), overlaps(true), tieup(false), proactive_split(false), prev_fam(false), sbl(false), mask_file("") {}
 };
 
 typedef unordered_map<size_t, LmerVector*> LmerMap;
@@ -359,7 +359,7 @@ bool fragmentSplit(LmerVector* v, uint L, vector<Family*> &families, AppOptions 
     //   fam->removeOne(v);
     //   v->getPrevFamily()->adopt(v,L);
     // }
-    else {
+    //else {
       Family* newFam = splitRepeatsByLmer(fam, fam->getLast(), true, L, options);
       families.push_back(newFam);
       // If splitting proactively, we know (last+1)...(v-1) did not
@@ -368,7 +368,7 @@ bool fragmentSplit(LmerVector* v, uint L, vector<Family*> &families, AppOptions 
         families.push_back(newFam2);
       }
       return true;
-    }
+    //}
   }
   return false;
 }
@@ -389,15 +389,15 @@ bool isPrefix(LmerVector *v) {
 }
 
 bool canMergeSkipped(Family* fam, uint L, bool overlaps) {
-  if (fam->getSkipped()->size() > 0){
-    return closeEnough(fam->getSkipped()->back(), fam->getOneAfterLast(), L, overlaps);
+  if (fam->getLastSkipped() && fam->getOneAfterLast()){
+    return closeEnough(fam->getLastSkipped(), fam->getOneAfterLast(), L, overlaps);
   }
   return true;
 }
 
-Family* mergeIntoFamily(vector<LmerVector*> lmers, uint L){
+Family* mergeIntoFamily(Family* fam, uint L){
   Family* newFam = new Family();
-  for (LmerVector* v : lmers){
+  for (LmerVector* v : fam->popSkipped()){
     newFam->adopt(v,L);
   }
   return newFam;
[…]

### 2015-07-31 RAIDER_src_v2_options_new/src/main.cpp

@@ -64,6 +64,11 @@ seqan::ArgumentParser::ParseResult parseCommandLine(AppOptions & options, int ar
             parser,
             seqan::ArgParseOption("s", "seed", "Spaced seed/mask to use. Defaults to 111110011111110001111111000000000000011111.",
                                   seqan::ArgParseOption::STRING));
+
+  addOption(
+	    parser,
+	    seqan::ArgParseOption("mf", "mask_file", "Masked sequence file.Default: off",
+				  seqan::ArgParseOption::STRING));
   
   addOption(
             parser,
@@ -122,6 +127,11 @@ seqan::ArgumentParser::ParseResult parseCommandLine(AppOptions & options, int ar
     seqan::getOptionValue(options.seed, parser, "seed");
   else
     options.seed = "111110011111110001111111000000000000011111";
+
+  if (isSet(parser, "mask_file"))
+    seqan::getOptionValue(options.mask_file, parser, "mask_file");
+  else
+    options.seed = "";
   
   if (isSet(parser, "min"))
     seqan::getOptionValue(options.min, parser, "min");
@@ -164,6 +174,7 @@ seqan::ArgumentParser::ParseResult parseCommandLine(AppOptions & options, int ar
   if (isSet(parser, "sbl")){
     options.sbl = true;
   }
+
   
   // Ensure a trailing /
   if (options.output_directory[seqan::length(options.output_directory) - 1] != '/') {
@@ -420,6 +431,10 @@ int main(int argc, char const ** argv) {
     // TODO throw exceptions
     return 1;
   }
+
+
+  cout << sequence[20] << endl;
+  exit(1);
   
   if (options.verbosity > 0) {
     cout << "BASE PAIRS\t" << seqan::length(sequence) << endl;

### 2015-08-03 RAIDER_src_v2_options_new/src/main.cpp
Adding file masking

@@ -131,7 +131,7 @@ seqan::ArgumentParser::ParseResult parseCommandLine(AppOptions & options, int ar
   if (isSet(parser, "mask_file"))
     seqan::getOptionValue(options.mask_file, parser, "mask_file");
   else
-    options.seed = "";
+    options.mask_file = "";
   
   if (isSet(parser, "min"))
     seqan::getOptionValue(options.min, parser, "min");
@@ -180,7 +180,7 @@ seqan::ArgumentParser::ParseResult parseCommandLine(AppOptions & options, int ar
   if (options.output_directory[seqan::length(options.output_directory) - 1] != '/') {
     seqan::append(options.output_directory, "/");
   }
-  
+
   return seqan::ArgumentParser::PARSE_OK;
 }
 
@@ -188,7 +188,9 @@ seqan::ArgumentParser::ParseResult parseCommandLine(AppOptions & options, int ar
  * Given a sequence stream, read in all sequences and concatenate into a master sequence
  */
 bool concatenateSequences(seqan::SequenceStream &seqStream, vector<idThreshold> &thresholds,
-                          seqan::Dna5String &outSequence, int verbosity) {
+                          vector<pair<uint,uint>>& seq_coords, seqan::Dna5String &outSequence, int verbosity, int L) {
+  thresholds.clear();
+  seq_coords.clear();
   seqan::CharString id;
   if (seqan::readRecord(id, outSequence, seqStream) != 0) {
     return false;
@@ -199,14 +201,26 @@ bool concatenateSequences(seqan::SequenceStream &seqStream, vector<idThreshold> 
   uint currentThreshold = seqan::length(outSequence);
   thresholds.push_back(make_pair(currentThreshold, id));
   seqan::Dna5String other;
+  seq_coords.push_back(make_pair<uint,uint>(0, seqan::length(outSequence)));
+
+  seqan::Dna5String buffer;
+  for (int i=0; i < L + 1; i++) // Better way to do this?
+    seqan::append(buffer, "N");
+
   while (seqan::readRecord(id, other, seqStream) == 0) {
     if (verbosity > 0) {
       cout << "Preparing " << id << endl;
     }
+    seqan::append(outSequence, buffer);
     currentThreshold += seqan::length(other);
+
+    uint start = seqan::length(outSequence);
+    uint finish = start + length(other);
+    seq_coords.push_back(make_pair(start,finish));
[…]

### 2015-08-04 RAIDER_src_v2_options_new/src/Family.h
Adding filters.  Tested; 90% its working.

@@ -5,6 +5,7 @@
 #include <iostream>
 #include <assert.h>
 #include <algorithm>
+#include <unordered_set>
 
 class Family {
 public:
@@ -163,6 +164,11 @@ public:
   uint expected_end;
   std::vector<LmerVector*> vectors;
   // std::vector<LmerVector*> skipped;
+
+  // Instances which are "removed" in post-processing, by LmetVector index.
+  // Removal only occurs in post-processing -- not relevant to RAIDER algorithm.
+  unordered_set<uint> excluded;  
+  
 };
 
 #endif //FAMILY_H

### 2015-08-04 RAIDER_src_v2_options_new/src/SeedChain.h

@@ -42,13 +42,15 @@ struct AppOptions {
   bool sbl;
   // File name for masked file (if specified)
   seqan::CharString mask_file;
+  // File name for filter_file (if specified)
+  seqan::CharString filter_file;
   
   seqan::CharString seed;
   seqan::CharString sequence_file;
   seqan::CharString output_directory;
   
   AppOptions() :
-  age(1), verbosity(1), family_array(true), excising(false), overlaps(true), tieup(false), proactive_split(false), prev_fam(false), sbl(false), mask_file("") {}
+  age(1), verbosity(1), family_array(true), excising(false), overlaps(true), tieup(false), proactive_split(false), prev_fam(false), sbl(false), mask_file(""), filter_file("") {}
 };
 
 typedef unordered_map<size_t, LmerVector*> LmerMap;

### 2015-08-04 RAIDER_src_v2_options_new/src/main.cpp

@@ -10,6 +10,7 @@
 #include <seqan/sequence.h>
 #include <seqan/arg_parse.h>
 #include <seqan/seq_io.h>
+#include <seqan/gff_io.h>
 
 #include <iostream>
 #include <sstream>
@@ -18,6 +19,7 @@
 #include <stdio.h>
 #include <unordered_map>
 #include <set>
+#include <unordered_set>
 #include <numeric>
 
 #include "SeedChain.h"
@@ -26,13 +28,15 @@
 
 using namespace std;
 
-typedef pair<uint, seqan::CharString> idThreshold;
+typedef pair<uint, string> idThreshold;
 
 const string OUTPUT_ELEMENTARY_REPEATS_FILENAME = "elements";
 const string OUTPUT_REPEAT_FAMILIES_FILENAME = "families";
 const string OUTPUT_COMPOSITES_FILENAME = "composites";
 const string OUTPUT_SUMMARY_FILENAME = "summary_info";
 
+
+typedef pair<int,int> Interval;
 
 // ==========================================================================
 // Functions
@@ -67,8 +71,13 @@ seqan::ArgumentParser::ParseResult parseCommandLine(AppOptions & options, int ar
 
   addOption(
 	    parser,
-	    seqan::ArgParseOption("mf", "mask_file", "Masked sequence file.Default: off",
+	    seqan::ArgParseOption("mf", "mask_file", "Masked sequence file. Default: none",
 				  seqan::ArgParseOption::STRING));
+
+  addOption(
+	     parser,
+	     seqan::ArgParseOption("ff", "filter_file", "Filter file. Default: none",
+				   seqan::ArgParseOption::STRING));
   
   addOption(
             parser,
@@ -132,6 +141,11 @@ seqan::ArgumentParser::ParseResult parseCommandLine(AppOptions & options, int ar
     seqan::getOptionValue(options.mask_file, parser, "mask_file");
   else
     options.mask_file = "";
+
+  if (isSet(parser, "filter_file"))
+    seqan::getOptionValue(options.filter_file, parser, "filter_file");
+  else
+    options.filter_file = "";
   
   if (isSet(parser, "min"))
     seqan::getOptionValue(options.min, parser, "min");
@@ -188,9 +202,15 @@ seqan::ArgumentParser::ParseResult parseCommandLine(AppOptions & options, int ar
  * Given a sequence stream, read in all sequences and concatenate into a master sequence
  */
[…]

### 2015-08-05 RAIDER_src_v2_options_new/src/Family.h
Modificiations for splitting

@@ -100,7 +100,7 @@ public:
     //std::vector<LmerVector*> ret(skipped);
     //skipped.clear();
     std::vector<LmerVector*> skipped(vectors.size());
-    std::remove_copy_if (vectors.begin(),vectors.end(),skipped.begin(), [](LmerVector* v){return v->gotSkipped();});
+    std::remove_copy_if (vectors.begin(),vectors.end(),skipped.begin(), [](LmerVector* v){return !v->gotSkipped();});
     vectors.shrink_to_fit();
     skipped.shrink_to_fit();  // shrink container to new size
     resetOffs();

### 2015-08-05 RAIDER_src_v2_options_new/src/LmerVector.h

@@ -58,8 +58,12 @@ public:
   
   friend ostream &operator<<( ostream &output, const LmerVector &L) {
     output  << "(\tFront:\t" << L.front()
-    << "\tBack:\t" << L.back()
-    << "\tSize:\t" << L.size()
+    << "\tBack:\t" << L.back() << "\tLocations:\t";
+    for (uint loc : L.lmers){
+      output << loc << " ";
+    }
+    output << "\tSize:\t" << L.size()
+    << "\tSkipped?\t" << L.gotSkipped()
     << "\tOffset In Family\t" << L.getOff()
     << "\tFamily:\t" << L.getFamily() << "\t)";
     return output;

### 2015-08-05 RAIDER_src_v2_options_new/src/SeedChain.h

@@ -76,6 +76,11 @@ void prettyPrintFamily(uint t, Family* fam, bool isNew, bool isMod){
   }
 }
 
+void prettyPrintModifiedFamily(uint t, Family* fam1, Family* fam2){
+  prettyPrintFamily(t, fam1, false, true);
+  prettyPrintFamily(t, fam2, true, false);
+}
+
 void prettyPrintMethodState(uint t, string methodname, LmerVector* v, Family* fam, bool isNew, bool isMod){
   prettyPrintTabbing(t);
   cout << "--- " << methodname << " ---" << endl;
@@ -253,11 +258,10 @@ Family* exciseRepeatsByLmer(Family* fam, LmerVector *v, uint L, AppOptions optio
   
   newFam->setLast(v);
   newFam->setExpectedEnd(v->back() + newFam->repeatLength(L));
-  
-  if (options.verbosity > 2){
-    prettyPrintFamily(2, fam, false, true);
-    prettyPrintFamily(2, newFam, true, false);
-  }
+ 
+
+  if (options.verbosity > 2)
+    prettyPrintModifiedFamily(2, fam, newFam);
   return newFam;
   
 }
@@ -300,10 +304,8 @@ Family* splitRepeatsByLmer(Family* fam, LmerVector *v, bool keepV, uint L, AppOp
     newFam->setExpectedEnd(v->back() + newFam->repeatLength(L));
   }
   
-  if (options.verbosity > 2){
-    prettyPrintFamily(2, fam, false, true);
-    prettyPrintFamily(2, newFam, true, false);
-  }
+  if (options.verbosity > 2)
+    prettyPrintModifiedFamily(2, fam, newFam);
   
   return newFam;
 }
@@ -391,11 +393,19 @@ bool canMergeSkipped(Family* fam, uint L, bool overlaps) {
   return true;
 }
 
-Family* mergeIntoFamily(Family* fam, uint L){
+Family* mergeIntoFamily(Family* fam, uint L, uint verbosity){
+  if (verbosity > 2){
+    cout << "-- Merge Into Family --" << endl;
+    prettyPrintFamily(1, fam, false, false);   
+  }
   Family* newFam = new Family();
   for (LmerVector* v : fam->popSkipped()){
-    newFam->adopt(v,L);
+    if (v)
+      newFam->adopt(v,L);
   }
+  if(newFam->size() == 0)
+    return 0;
[…]

### 2015-08-05 RAIDER_src_v2_options_new/src/Family.h
licensing

@@ -1,3 +1,20 @@
+// Family.h is part of phRAIDER.
+//
+// phRAIDER is free software: you can redistribute it and/or modify
+// it under the terms of the GNU General Public License as published by
+// the Free Software Foundation, either version 3 of the License, or
+// (at your option) any later version.
+//
+// phRAIDER is distributed in the hope that it will be useful,
+// but WITHOUT ANY WARRANTY; without even the implied warranty of
+// MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+// GNU General Public License for more details.
+//
+// You should have received a copy of the GNU General Public License
+// along with phRAIDER.  If not, see <http://www.gnu.org/licenses/>.
+
+// Created by Carly Schaeffer, Nathan Figueroa, and John Karro
+
 #ifndef FAMILY_H
 #define FAMILY_H
 #include <vector>

### 2015-08-05 RAIDER_src_v2_options_new/src/SeedChain.h

@@ -1,3 +1,20 @@
+// SeedChain.h is part of phRAIDER.
+//
+// phRAIDER is free software: you can redistribute it and/or modify
+// it under the terms of the GNU General Public License as published by
+// the Free Software Foundation, either version 3 of the License, or
+// (at your option) any later version.
+//
+// phRAIDER is distributed in the hope that it will be useful,
+// but WITHOUT ANY WARRANTY; without even the implied warranty of
+// MERCHANTABILITY or FITNESS FOR A PARTICULAR PURPOSE.  See the
+// GNU General Public License for more details.
+//
+// You should have received a copy of the GNU General Public License
+// along with phRAIDER.  If not, see <http://www.gnu.org/licenses/>.
+
+// Created by Carly Schaeffer, Nathan Figueroa, and John Karro
+
 #ifndef RAIDER2_SEEDCHAIN_H
 #define RAIDER2_SEEDCHAIN_H
 

### 2015-08-05 RAIDER_src_v2_options_new/src/main.cpp
Removing debugging code from main.cpp

@@ -420,7 +420,6 @@ void writeSummary(vector<Family*> &families, AppOptions &options) {
 }
 
 void writeResults(vector<Family*> &families, vector<idThreshold> &thresholds, AppOptions &options) {
-  cout << "Y: " << families[0]->size() << endl;
   writeFamilies(families, options);
   writeRepeats(families, thresholds, options);
   writeSummary(families, options);
@@ -574,7 +573,6 @@ int main(int argc, char const ** argv) {
     mask_sequence(families, sequence, thresholds, seq_coords, options.mask_file, L, options.count);
   }
 
-  cout << "X: " << families[0]->size() << endl;
   writeResults(families, thresholds, options);
   
   return 0;

### 2015-08-05 RAIDER_eval.py
removed print statements remaining from debugging

@@ -718,7 +718,7 @@ def run_scout_second_filter_RM(p, num_processors):
         p2.input_name = p.input_name
         p2.stage = "RM"
         p2.description = p.description#"rep_scout"
-        print("F1 resources : " + str(p2.tool_resources))
+        #print("F1 resources : " + str(p2.tool_resources))
         return p2
         
     else:
@@ -747,10 +747,10 @@ def run_scout_second_filter(p):
             p2.submit_timed_job(preserve=True)
         p2.description = p.description#"rep_scout"
         p2.stage = "2"
-        print("RM resources : " + str(p.getResources(cleanup=False)))
+        #print("RM resources : " + str(p.getResources(cleanup=False)))
 
         p2.tool_resources = [x + y for x, y in zip(p.tool_resources, p.getResources(cleanup=False))]
-        print("F1 + RM resources : " + str(p2.tool_resources))
+        #print("F1 + RM resources : " + str(p2.tool_resources))
         p2.seq_file = p.seq_file
         p2.lib_file = filter_stage2_output
         p2.should_filter_stage2 = p.should_filter_stage2
@@ -1158,10 +1158,10 @@ def run_timed_tool_jobs(jobs, run_rm, pa, run_pra, blast_db, RM_jobs=None, PRA_j
                         added_jobs.add(run_scout_second_filter(j))
                 else:
                     if "rep_scout" in j.description and j.stage == "2":
-                        print("F2 resources : " + str(list(j.getResources(cleanup=False))))
+                        #print("F2 resources : " + str(list(j.getResources(cleanup=False))))
                         #j.tool_resources = j.tool_resources + j.getResources(cleanup=False) #should we include the RM inside RS in timing? #p.getResources(cleanup=False)
                         j.tool_resources = [x + y for x, y in zip(j.tool_resources, j.getResources(cleanup=False))] #j.tool_resources + j.getResources(cleanup = False)
-                        print("F1 + RM + F2 resources : " + str(j.tool_resources))
+                        #print("F1 + RM + F2 resources : " + str(j.tool_resources))
                     rm_job = None
                     pra_job = None
                     if run_rm:
@@ -1227,7 +1227,7 @@ def run_timed_analysis_jobs(run_rm, run_pra, RM_jobs, PRA_jobs, results_dir , st
                 exit_now()
             pra_job_set = pra_job_set - finished_pra_jobs
[…]

### 2015-08-05 rev_sf.sh
Added script for running RAIDER_eval on phRAIDER

@@ -1,0 +1,78 @@
+#!/bin/bash
+
+cd ~/karro/RAIDER_eval/
+##comp=comparisons_overlap3
+
+
+
+if [ -n "$sfile" ]; then
+    id=$(basename $sfile .txt)
+    seed_str="--sf ${sfile}"
+else
+    L=${#seed}
+    weight=$(grep -o "1" <<< "$seed" | wc -l)
+    id=L${L}_W${weight}_S${seed}
+    seed_str="-s ${seed}"
+fi
+
+echo $id
+basedata=${data##*/}
+fname=${dir}/F${freq}_${id}
+
+
+module load python-3.3.3
+##data=hg18.chr22.fa
+
+opts_dir_str=""
+opts=""
+#if [ -v $PS ]; then
+#    opts="${opts} --ps"
+#    opts_dir_str=${opts_dir_str}_PROSPLIT
+#fi
+
+
+echo "Output directory: ${fname}"
+echo "Data: ${basedata}"
+echo "Seed file: ${sfile}"
+
+if [ ${type} == "chrom_sim" ]; then
+    fname=${dir}/sim/F${freq}_${id}
+    if [ -z $T ]; then
+        t_str=""
+        T=1
+    else
+        t_str="-t ${T}"
+    fi
+    if [ -z $ST ]; then
+        st_str=""
+        ST=0
+    else
+        st_str="--st ${ST}"
+    fi
+    if [ -z $RNG ]; then
+        rng_str=""
+        RNG=None
+    else
+        rng_str="--rng_seed ${RNG}"
+    fi
+    fname=${fname}_T${T}_ST${ST}_RNG${RNG}
+    
+    if [ -z $FAM ]; then
+        fam_str=""
+    else
+        fam_str="--family_file ${FAM}"
+        fname=${fname}_FAM
+    fi
+
+
+    ./RAIDER_eval.py --R2 -R --nuke -r $fname -f ${freq} ${seed_str} --age 1  --timing_jobs chrom_sim ${t_str} ${st_str} ${rng_str} ${fam_str} ${data} 
+    mv ${fname}/SOURCE_DATA/ sim_data/SIM_T${T}_ST${ST}_RNG${RNG}_FAM${FAM}
+else
+    fname=${dir}/seq/F${freq}_${id}_${opts_dir_str}
[…]

### 2015-08-05 aseed_files/programs/create_all_pals.py
Commented the helper scripts I have used in testing/analysis. Also created Carly.README with descriptions

@@ -1,10 +1,16 @@
 #!/software/python/3.3.3/bin/python3.3
 
+# create_all_pals.py: generate all palindromic spaced seeds possible of specified length and weight
+# unrealistically long seed file when |length - weight| gets large.
+# Arguments:    optionally specify length (-l/--length) and weight (-w/--weight)
+#               must specify seed_file (output file path)
+# by Carly Schaeffer
+
+
 import itertools
 import argparse
 
-
-def generate_seed_file(seed_file, length, weight): 
+def generate_seed_file(seed_file, length, weight):
     f = open(seed_file, 'w')
     for bits in itertools.combinations(range((length - 2)//2), (weight - 2)//2):
         s = ['0'] * length

### 2015-08-05 doCombinedAnalysis.R

@@ -1,53 +1,62 @@
 #!/software/R/3.0.2-gcc/bin/Rscript
+
+# doCombinedAnalysis.R : script for combining stats data from multiple RAIDER_eval
+#   output directories. Recursively searches for all files in specified directory
+#   that are named "stats.txt". Combines all data together, and allows options for
+#   adding/removing/formatting data.
+#   Arguments:
+#     -v/--verbose  specifies whether output will print to stdout. default = True
+#     -d/--dir      specifies directory in which to recursively search for stats files
+#     -o/--output   specifies file to print output to
+#     --formula     specifies sorting formula for output. default = ~-tpr (Descending by sensitivity)
+#     -a/--auto_out specifies to print output to default file in same directory
+#     -t/--add_type specifies to add info about whether analyzing seq/sim data
+#     -c/--add_chrom specifies to add info about chromosome being analyzed
+#     -s/--suppress_seeds specifies not to add seed specific information
+#     -i            specifies to not include time/space complexity information
+#     -m            specifies to include directory information
+# by Carly Schaeffer
+
 suppressPackageStartupMessages(require(optparse))
 source("sorted_analyze.R")
 
-# specify our desired options 
-# by default ArgumentParser will add an help option 
+# allowed arguments
 option_list_2 = list(
-    make_option(c("-v", "--verbose"), action="store_true", default=FALSE,
-        help="Print extra output [default]"),
-    make_option(c("-d", "--dir"), action="store", type="character", default=NA,
-        help="Directory to recursively search for stats files"),
-    make_option(c("-o", "--output"), type="character", default=NA, 
-        help="File to print output to"),
-    make_option(c("--formula"), type="character", default="~-tpr", 
-        help = "Sorting formula [default \"%(default)s\"]"),
-    make_option(c("-a", "--auto_out"), action="store_true", default=FALSE,
-        help="Automatically create output file in same directory."),
-    make_option(c("-t", "--add_type"), action="store_true", default=FALSE,
-        help="Automatically add type info (sim vs seq_files)."),
-    make_option(c("-c", "--add_chrom"), action="store_true", default=FALSE,
-        help="Automatically add chrom info."),
-    make_option(c("-s", "--suppress_seeds"), action="store_false", default=TRUE,
[…]

### 2015-08-05 sorted_analyze.R

@@ -1,244 +1,214 @@
 #!/software/R/3.0.2-gcc/bin/Rscript
 
+# sorted_analyze.R : helper R script for formatting stats data from RAIDER_eval.
+#   provides functions for sorting data, as well as adding extra/removing excess
+#   information to/from the data. also includes function for outputting data to
+#   file/stdout.
+# by Carly Schaeffer
+
 suppressPackageStartupMessages(require(optparse))
 suppressPackageStartupMessages(require(gdata))
-getDf <- function(fname) {
-    #dat <- readLines(fname,n=-1)
-    #names <- unlist(strsplit(dat[1]," "))
-    #names <- names[names != ""]
-    #dat <- dat[-1]
-    #dat <- dat[grep("INCOMPLETE",dat,invert=TRUE)]
-    #dat <- unlist(lapply(dat,strsplit," "))
-    #dat <- dat[dat != ""]
-    #dat <- data.frame(matrix(dat,ncol=length(names),byrow=TRUE))
-    #colnames(dat)<-names
-    #as.data.frame(sapply(dat, function(x) gsub("\"", "", x)))
-    #dat[2:length(names)] <- lapply(dat[2:length(names)], function(x) as.numeric(levels(x)[x]))
-    #dat
-    dat <- read.table(fname,header=TRUE,comment.char="",fill=TRUE)
-    dat <- dat[rowSums(is.na(dat)) < 4, ]
 
-    #cat(dim(dat))
-    #cat(length(names(dat)))
-    #write.table(dat)
+# Reads in data frame from stats file
+getDf <- function(fname) {
+  dat <- read.table(fname,header=TRUE,comment.char="",fill=TRUE)
+  dat <- dat[rowSums(is.na(dat)) < 4, ]
 }
 
-
 sumOverMatches <- function(matches){
-    sum(as.numeric(unlist(matches)))
+  sum(as.numeric(unlist(matches)))
 }
 
 getMatches <- function(expr, seed){
-    regmatches(seed, gregexpr(expr, seed, perl=TRUE))
+  regmatches(seed, gregexpr(expr, seed, perl=TRUE))
 }
 
+# Determines number of times val is to be seen in expanded seed (accounting for formatting)
 numberOf <- function(seed, val){
-    match.string <- paste("(?<=", val, "\\^\\{)[[:digit:]]+(?=\\})", sep="")
-    numOnes <- sumOverMatches(getMatches(match.string, as.character(seed)))
[…]

### 2015-08-06 RAIDER_eval.py
Updating for Oakley (semi-tested)

@@ -1,17 +1,28 @@
-#!/software/python/3.3.3/bin/python3.3
 import sys
 import subprocess
 import os
 import os.path
 import shutil
 import argparse
-from redhawk import *
 import tempfile
 import re
 import perform_stats
 import time
 import pickle
 from checkpoint_jobs import *
+
+try:
+    fp = open("locations.txt")
+except: 
+    sys.stderr.println("locations.txt file not present.")
+    exit(1);
+
+location = fp.readline().strip()
+if not (location.lower() in {'redhawk', 'oakley', 'osx'}):
+    sys.stderr.println("location.txt: Bad content")
+    exit(1)
+
+
 
 #################################################################
 # The following global variables are related to debugging issues.
@@ -66,7 +77,9 @@ MacLocations = {'build_lmer_table':'/usr/local/RepeatScout/build_lmer_table',
                 'bigfoot':'./bigfoot',
                 'python':'python3.4',
                 'araider':'./araider',
-                'raider2': './raiderv2_options'}#,'raider2_old':'./raiderv2_old','raider2_oldest':'./raiderv2_oldest'}
+                'raider2': './raiderv2_options',
+                'rm_modules': [],
+                'RepeatMasker' : 'RepeatMasker'}
 RedhawkLocations = {'build_lmer_table':'./build_lmer_table',
                     'RptScout':'./RepeatScout',
                     'filter_stage-1':'./filter-stage-1.prl',
@@ -76,7 +89,21 @@ RedhawkLocations = {'build_lmer_table':'./build_lmer_table',
                     'bigfoot':'./bigfoot',
                     'python':'python3.3',
                     'araider':'./araider',
-                    'raider2': './raiderv2_options'}#,'raider2_old':'./raiderv2_old','raider2_oldest':'./raiderv2_oldest'}
+                    'raider2': './raiderv2_options',
+                    'rm_modules' : ['RepeatMasker', 'python-3.3.3'],
+                    'RepeatMasker' : 'RepeatMasker'}
+OakleyLocations = {'build_lmer_table':'./build_lmer_table',
+                   'RptScout':'./RepeatScout',
+                   'filter_stage-1':'./filter-stage-1.prl',
+                   'filter_stage-2':'./filter-stage-2.prl',
+                   'raider':'./raider',
+                   'raider_pre':'./raider_pre',
+                   'bigfoot':'./bigfoot',
[…]

### 2015-08-06 redhawk.py

@@ -23,7 +23,6 @@ import pickle
 import tempfile
 import logging
 
-
 epilogue_str = """#!/bin/sh
 echo "Redhawk Epilogue Args:" >&2
 echo "Job ID: $1" >&2
@@ -44,17 +43,18 @@ Redhawk Epilogue Args:
 Resources Used: cput=00:00:00,mem=0kb,vmem=0kb,walltime=00:00:00
 """
 
+
 ############################################################
 # Library-global variables
-log_file = "/usr/local/torque/current/var/spool/torque/server_priv/accounting"
 uid = os.getuid()
 current_user = pwd.getpwuid(uid)[0]  # When run on redhawk with a nohup, os.getlogin() does not work
 try:
-    subprocess.check_call(["qstat"], stdout = subprocess.PIPE, stderr = subprocess.PIPE)
+    subprocess.check_call(["qstat > /dev/null"], shell = True)
 except:
     pbs_present = False
 else:
     pbs_present = True
+
 
 job_list = set()    # Global list of all jobs that have been submitted and not yet identified as having quit
 
@@ -762,6 +762,9 @@ def relaunch(args = sys.argv, force = False, walltime = "40:00:00", python = "py
         sys.stderr.write("STDERR: " + e)
         return True
     return False
+
+
+
         
 ############## Allow for a direct launch of a program
 if __name__ == "__main__":

### 2015-08-07 RAIDER_eval.py
Updating RAIDER_eval with oakley modifications

@@ -99,7 +99,7 @@ OakleyLocations = {'build_lmer_table':'./build_lmer_table',
                    'raider':'./raider',
                    'raider_pre':'./raider_pre',
                    'bigfoot':'./bigfoot',
-                   'python':'python3.3',
+                   'python':'python',
                    'araider':'./araider',
                    'raider2': './raiderv2_options',
                    'rm_modules' : [],
@@ -518,7 +518,7 @@ def run_raider2(seed, seed_num, f, m, input_file, raider2_dir, family_array, exc
     else:
         opt_str += "--age " + str(age) 
 
-    cmd1 = "{raider2} --vv -q -c {f} {version} {min_arg} {seed} {input_file} {output_dir}".format(raider2 = Locations['raider2'], f = f, version = opt_str, min_arg = min_arg, seed = seed_string, input_file = input_file, output_dir = output_dir)
+    cmd1 = "{raider2} -q -c {f} {version} {min_arg} {seed} {input_file} {output_dir}".format(raider2 = Locations['raider2'], f = f, version = opt_str, min_arg = min_arg, seed = seed_string, input_file = input_file, output_dir = output_dir)
 
     out_file = raider2_dir + "/" + input_base + ".s" + str(seed_num) + ".raider2_consensus.txt"
     lib_file = raider2_dir + "/" + input_base + ".s" + str(seed_num) + ".raider2_consensus.fa"
@@ -787,7 +787,7 @@ def run_scout_second_filter(p):
 
 def scout_second_filter(p, min_freq):
     """NOT CURRENTLY WORKING!!! Does not run correctly, and does not properly adjust time"""
-    p.wait()
+    p.wait(100)
 
     filter2_stage_output = p.seq_file.rstrip(".fa") + ".repscout.filtered2.fa"
     cmd = "cat {output} | perl {filter} --cat={cat} --thresh={thresh} > {final}".format(output = p.lib_file, filter = Locations['filter_stage-2'], cat = p.rm_output, thresh = min_freq, final = filter2_stage_output)
@@ -821,7 +821,7 @@ def run_repeat_masker(p, num_processors):
     """Given the pbs object used to start a consensus sequence job as well as
     repeatmasker arguments, wait until the job is done and then call repeatmasker 
     on the output and put results in masker_dir (current dir if unspecified)"""
-    p.wait()
+    p.wait(100)
     p.loadResources()
 
     input_base = file_base(p.seq_file)  # Base name of the file used for input
[…]

### 2015-08-08 RAIDER_eval.py
Updating RAIDER_eval.py.  (Untested.)

@@ -79,7 +79,8 @@ MacLocations = {'build_lmer_table':'/usr/local/RepeatScout/build_lmer_table',
                 'araider':'./araider',
                 'raider2': './raiderv2_options',
                 'rm_modules': [],
-                'RepeatMasker' : 'RepeatMasker'}
+                'RepeatMasker' : 'RepeatMasker',
+                'proc_per_node' : 1}
 RedhawkLocations = {'build_lmer_table':'./build_lmer_table',
                     'RptScout':'./RepeatScout',
                     'filter_stage-1':'./filter-stage-1.prl',
@@ -91,7 +92,8 @@ RedhawkLocations = {'build_lmer_table':'./build_lmer_table',
                     'araider':'./araider',
                     'raider2': './raiderv2_options',
                     'rm_modules' : ['RepeatMasker', 'python-3.3.3'],
-                    'RepeatMasker' : 'RepeatMasker'}
+                    'RepeatMasker' : 'RepeatMasker',
+                    'proc_per_node' : 4}
 OakleyLocations = {'build_lmer_table':'./build_lmer_table',
                    'RptScout':'./RepeatScout',
                    'filter_stage-1':'./filter-stage-1.prl',
@@ -103,7 +105,8 @@ OakleyLocations = {'build_lmer_table':'./build_lmer_table',
                    'araider':'./araider',
                    'raider2': './raiderv2_options',
                    'rm_modules' : [],
-                   'RepeatMasker' : 'RepeatMasker'}
+                   'RepeatMasker' : 'RepeatMasker',
+                   'proc_per_node' : 12}
 Locations = None;    # This will be set to one of the above two, and references to find exectuable locations.
 
 
@@ -209,7 +212,8 @@ def parse_params(args):
     raider2_argument.add_argument('--no', '--no_overlaps', dest="overlaps", action="store_false", help="Do not require overlaps in RAIDER2", default=True)
     raider2_argument.add_argument('--tu', '--tie_up', dest="tieup", action="store_true", help="Enable alternative tie ups", default=False)
     raider2_argument.add_argument('--ps', '--prosplit', dest="prosplit", action="store_true", help="Enable proactive splitting(disabled by default).", default=False)
-    raider2_argument.add_argument("--pf", "--prevfam", dest="prevfam", action="store_true", help="Enable pointers to prev family (disabled by default).", default=False)
[…]

### 2015-08-08 RAIDER_eval.py
Updatind RAIDER_eval.py (not finished)

@@ -213,7 +213,7 @@ def parse_params(args):
     raider2_argument.add_argument('--tu', '--tie_up', dest="tieup", action="store_true", help="Enable alternative tie ups", default=False)
     raider2_argument.add_argument('--ps', '--prosplit', dest="prosplit", action="store_true", help="Enable proactive splitting(disabled by default).", default=False)
     raider2_argument.add_argument("--pf", '--prevfam', dest="prevfam", action="store_true", help="Enable pointers to prev family (disabled by default).", default=False)
-    raider2_arugment.add_arggument("--mn", '--max_nodes', dest = "max_nodes", action="store_true", help="Reserve all nodes of a processor (disabled by default).", default=False)
+    raider2_argument.add_argument("--mn", '--max_nodes', dest = "max_nodes", action="store_true", help="Reserve all nodes of a processor (disabled by default).", default=False)
 
     # REPSCOUT ARGUMENTS
     repscout_argument = parser.add_argument_group("REPSCOUT parameters")

### 2015-08-08 RAIDER_eval.py
Correction to RAIDER_eval.py

@@ -1412,7 +1412,7 @@ if __name__ == "__main__":
             jobs += [run_raider2(seed = convert_seed(seed), seed_num = i, f = args.f, m = args.raider_min, input_file = file,
                                  raider2_dir = args.results_dir + "/" + args.raider2_dir, family_array = args.family_array, excise = args.excising, 
                                  overlaps = args.overlaps, tieup = args.tieup, prosplit=args.prosplit, prevfam=args.prevfam,
-                                 age=args.age, age_only=False) for i,seed in enumerate(seed_list) for file in file_list]
+                                 age=args.age, age_only=False, max_nodes=args.max_nodes) for i,seed in enumerate(seed_list) for file in file_list]
             #if args.all_ages:
             #    if not args.multi_seed:
             #        jobs += [run_raider2(seed = convert_seed(seed), seed_num = i, f = args.f, m = args.raider_min, input_file = file, 

### 2015-09-12 testing_pipeline.py
Updating pipeline software

@@ -23,7 +23,7 @@ seed_list = None      # Sorted list of seeds
 
 #######################
 # Defaults 
-walltime_default = "4:00:00"
+walltime_default = "10:00:00"
 delay_default = 300           # Number of seconds to sleep when cycling on a redhawk wait
 
 tool_prefix = {'phRAIDER' : 'phRA', 'RepeatScout' : 'RS', 'RAIDER': 'RA', 'pre-phRAIDER' : 'prephRA'}
@@ -129,7 +129,7 @@ def parse_params():
     repeatmasker_arguments.add_argument('--RM', '--suppress_rm', dest = 'run_rm', action="store_false", help = "suppress RepeatMasker run", default = True)
     repeatmasker_arguments.add_argument('--masker_dir', help = "Repeat masker output directory", default = None)
     repeatmasker_arguments.add_argument('-p', '--pa', type = int, help = "Number of processors will be using", default = 1)    
-    repeatmasker_arguments.add_argument('--rwt', '--rm_walltime', dest = "rm_walltime", help = "Wall time limit for repeat masker", default = "10:00:00")
+    repeatmasker_arguments.add_argument('--rwt', '--rm_walltime', dest = "rm_walltime", help = "Wall time limit for repeat masker", default = "30:00:00")
 
     # BLAST ARGUMENTS
     blast_arguments = parser.add_argument_group("BLAST parameters")
@@ -137,6 +137,8 @@ def parse_params():
     blast_arguments.add_argument('--evalue', dest = 'evalue', help = "BLASE evalue", default = "0.000001");
     blast_arguments.add_argument('--short', dest = 'short', action = 'store_false', help = "Turn off blast-short on blast run", default = True)
     blast_arguments.add_argument('--max_target', dest = 'max_target', action = 'store', help = "BLAST --max_target option", default = '999999999') # HACK!!!  Need to fix this
+    blast_arguments.add_argument('--bl_walltime', help = "Blast walltime.", default = "30:00:00")
+    blast_arguments.add_argument('--pb', dest = "pb", type = int, help = "Number of cpus for blast", default = 1)
 
     # DEBUGGING ARGUMENTS
     debug_group = parser.add_argument_group(title = "debugging")
@@ -195,8 +197,8 @@ def launch_job(cmd, title, base_dir, walltime = walltime_default, ppn = 1, bigme
     p = pbsJobHandler(batch_file = batch_file, executable = cmd, job_name = job_name,
                       stdout_file = stdout_file, stderr_file = stderr_file, res_file = res_file,
[…]

### 2015-09-12 testing_pipeline.py
Adding testing_pipeline.py

@@ -195,7 +195,7 @@ def launch_job(cmd, title, base_dir, walltime = walltime_default, ppn = 1, bigme
     p = pbsJobHandler(batch_file = batch_file, executable = cmd, job_name = job_name,
                       stdout_file = stdout_file, stderr_file = stderr_file, res_file = res_file,
                       walltime = walltime, depends = depend,
-                      mem = Locations['high_mem_arch'] if args.mem else False,
+                      mem = Locations['high_mem_arch'] if bigmem else False,
                       RHmodules = modules)
 
 
@@ -237,10 +237,21 @@ def setup():
     progress_fp = open(debug_file, "w")
 
     global seed_map
-    if args.seed_file:
-        seed_map = {convert_seed(seed):(i,seed) for i,line in enumerate(open(args.seed_file)) for seed in [line.strip()]}
+    if os.path.exists(args.results_dir + "/seed_file.txt"):
+        seed_map = {convert_seed(seed):i for line in open(args.results_dir + "/seed_file.txt") for i,seed in [re.split("\t+", l.strip())]}
     else:
-        seed_map = {convert_seed(args.seed):(0,args.seed)}
+        seed_map = {}
+
+    offset = len(seed_map)    
+    if args.seed_file:
+        for line in open(args.seed_file):
+            seed = line.rstrip()
+            if seed not in seed_map:
+                seed_map[seed] = offset
+            offset += 1
+    else:
+        if args.seed not in seed_map:
+            seed_map[args.seed] = offset
 
     global seed_list
     seed_list = sorted(seed_map.values(), key = lambda x: x[0])

### 2015-09-12 testing_pipeline.py
Adding testing_pipeline

@@ -1,0 +1,488 @@
+#!/software/python/3.3.3/bin/python3.3
+import sys
+import os
+import argparse
+import locations
+import subprocess
+import pickle
+from redhawk import *
+
+# Second version of the script for testing RAIDER.
+# Generates runs only; will use a seperate script for testing.
+
+
+
+#######################
+# Globals: So we don't have to pass everything through parameters.
+args = None     # Command line arguments object
+Locations = locations.Locations
+progress_fp = None
+job_log_dir = None
+seed_map = {}         # Maps seed to (index,short rep.) pairs
+seed_list = None      # Sorted list of seeds
+
+#######################
+# Defaults 
+walltime_default = "4:00:00"
+delay_default = 300           # Number of seconds to sleep when cycling on a redhawk wait
+
+tool_prefix = {'phRAIDER' : 'phRA', 'RepeatScout' : 'RS', 'RAIDER': 'RA', 'pre-phRAIDER' : 'prephRA'}
+
+#######################
+# Useful utilitiy functions
+def file_base(file):
+    """Extract the name of a file froma directory"""
+    return os.path.basename(file)
+
+def file_dir(file):
+    """Extract the directory a file is contained in"""
+    return file.rstrip(file_base(file)).rstrip("/")
+
+def make_dir(DIR):
+    if not os.path.exists(DIR):
+        os.makedirs(DIR)
+    return DIR
+
+def convert_seed(seed):
+    """Convert an abriviated seed to a full seed (e.g. "1{2}0{3}1{2}" => "1100011" """
+    i = 0
+    while (i < len(seed)-1):
+        if seed[i+1] == '^':
+            j = i+2
+            assert seed[j] == "{"
+            k = j+1
+            while seed[k] != '}':
+                k += 1
+            n = int(seed[j+1:k])
+            seed = seed[:i] + seed[i]*n + seed[k+1:]
+        i += 1
+    return seed
+
+def seed_weight(seed):
+    return sum([x == 1 for x in seed])
+
+def cmp_seed(s1, s2):
+    d = len(s1) - len(s2)
+    if (d != 0):
+        return d < 0
+    d = weight(s1) - weight(s2)
+    if f != 0:
+        return d < 0
+    return s1 < s2
+
+#######################
+# Command line parsing
+def parse_params():
+    """Parse command line arguments, set them equal to the global args, and set certain global variables"""
+    parser = argparse.ArgumentParser(description = "Evaluate RAIDER against RepeatScout")
+
+    # TOOL SELECTION
[…]

### 2015-09-12 testing_pipeline.py
Updating testing_pipeline.py

@@ -273,7 +273,7 @@ def setup():
 
 
         
-raider_cmd = "{raider} -q -c {f} -s {seed} {input_file} {output_dir}"
+raider_cmd = "/usr/bin/time {raider} -q -c {f} -s {seed} {input_file} {output_dir}"
 consensus_cmd = "{python} consensus_seq.py -s {data_file} -e {elements_file} {consensus_txt} {consensus_fa}"
 repeat_masker_cmd = "{RepeatMasker} -nolow -lib {library_file} -pa {pa} -dir {output_dir} {seq_file}"
 blast_format = "6 qseqid sseqid qstart qend qlen sstart send slen"
@@ -304,6 +304,9 @@ def raider_pipeline(raider_exe, input_file, seed, f):
     # Step 1: Run phRAIDER
     cmd1 = raider_cmd.format(raider=Locations[raider_exe], f=f, seed=seed,
                              input_file=input_file, output_dir=elements_dir)
+
+    if raider_exe == "RAIDER":  # Hack -- I should modify the raider to code
+        raider_exe = re.sub("-s ", " ", raider_exe)
     title1 = title;
     p1 = launch_job(cmd=cmd1, title=title1, base_dir=elements_dir, ppn = Locations['proc_per_node'] if args.max_nodes else 1, bigmem = args.mem, attrs = {'elements':'elements_dir/elements'})
 
@@ -337,8 +340,8 @@ def raider_pipeline(raider_exe, input_file, seed, f):
 
 
 #################################
-build_lrm_cmd = "{build_lmer_table_exe} -min {min} -sequence {seq_file} -freq {lmer_output}"
-rptscout_cmd = "{RptScout_exe} -sequence {seq_file} -freq {lmer_output} -output {output}"
+build_lrm_cmd = "/usr/bin/time {build_lmer_table_exe} -min {min} -sequence {seq_file} -freq {lmer_output}"
+rptscout_cmd = "/usr/bin/time {RptScout_exe} -sequence {seq_file} -freq {lmer_output} -output {output}"
 filter1_cmd = "{filter} {input} > {filter_output}"
 filter2_cmd = "cat {filtered} | {filter} --cat={rm_output} --thresh={thresh} > {filter_output}"
 def rptscout_pipeline(input_file, f):
@@ -476,7 +479,7 @@ if __name__ == "__main__":
         for f in args.f:
             if args.run_repscout:
                 rptscout_pipeline(file, f)
-            if args.run_raider:
+            if args.run<_raider:
[…]

### 2015-09-12 redhawk.py
Updating redhawk and testing_pipeline

@@ -184,7 +184,7 @@ class pbsJobHandler:
 
         if self.depends and len(self.depends) > 0:
             s = "#PBS -W depend=" + ",".join(["afterok:" + str(x.jobid) for x in self.depends]) + "\n"
-        f.write(s)
+            f.write(s)
         
         if stdout_file:
             self.ofile = self.output_location + "/" + stdout_file

### 2015-09-12 testing_pipeline.py

@@ -196,7 +196,7 @@ def launch_job(cmd, title, base_dir, walltime = walltime_default, ppn = 1, bigme
                       stdout_file = stdout_file, stderr_file = stderr_file, res_file = res_file,
                       walltime = walltime, depends = depend,
                       mem = Locations['high_mem_arch'] if bigmem else False,
-                      RHmodules = modules)
+                      RHmodules = modules, ppn = ppn)
 
 
     if attrs:
@@ -305,8 +305,8 @@ def raider_pipeline(raider_exe, input_file, seed, f):
     cmd1 = raider_cmd.format(raider=Locations[raider_exe], f=f, seed=seed,
                              input_file=input_file, output_dir=elements_dir)
 
-    if raider_exe == "RAIDER":  # Hack -- I should modify the raider to code
-        raider_exe = re.sub("-s ", " ", raider_exe)
+    if raider_exe == "RAIDER":  # Hack -- I should modify the raider code
+        cmd1 = re.sub("-s", " ", cmd1)
     title1 = title;
     p1 = launch_job(cmd=cmd1, title=title1, base_dir=elements_dir, ppn = Locations['proc_per_node'] if args.max_nodes else 1, bigmem = args.mem, attrs = {'elements':'elements_dir/elements'})
 
@@ -479,7 +479,7 @@ if __name__ == "__main__":
         for f in args.f:
             if args.run_repscout:
                 rptscout_pipeline(file, f)
-            if args.run<_raider:
+            if args.run_raider:
                 for seed in seed_map:
                     raider_pipeline('RAIDER', file, seed, f)
             if args.run_phraider:

### 2015-09-13 testing_pipeline.py
Update testing_pipeline.py

@@ -238,23 +238,30 @@ def setup():
 
     global seed_map
     if os.path.exists(args.results_dir + "/seed_file.txt"):
-        seed_map = {convert_seed(seed):(int(i),seed) for line in open(args.results_dir + "/seed_file.txt") for i,seed in [re.split("\t+", line.strip())]}
+        with open(args.results_dir + "/seed_file.txt") as fp:
+            seed_map = {convert_seed(seed):(int(i),seed) for line in fp for i,seed in [re.split("\t+", line.strip())]}
     else:
         seed_map = {}
 
+    seed_map2 = {}
+        
     offset = len(seed_map)    
     if args.seed_file:
         for line in open(args.seed_file):
             seed = convert_seed(line.rstrip())
             if seed not in seed_map:
                 seed_map[seed] = (offset,line.rstrip())
+                seed_map2[seed] = (offset,line.rstrip())
                 offset += 1
+            else:
+                seed_map2[seed] = (offset,line.rstrip())
     else:
         if args.seed not in seed_map:
             seed_map[convert_seed(args.seed)] = (offset,args.seed)
+            seed_map2[convert_seed2(args.seed)] = (offset,args.seed)
 
     global seed_list
-    seed_list = sorted(seed_map.values(), key = lambda x: x[0])
+    seed_list = sorted(seed_map2.values(), key = lambda x: x[0])
     with open(args.results_dir + "/seed_file.txt", "w") as fp:
         fp.write("\n".join([str(x[0]) + "\t" + x[1] for x in seed_list]))
 
@@ -480,12 +487,12 @@ if __name__ == "__main__":
             if args.run_repscout:
                 rptscout_pipeline(file, f)
             if args.run_raider:
-                for seed in seed_map:
+                for seed in seed_list:
                     raider_pipeline('RAIDER', file, seed, f)
             if args.run_phraider:
-                for seed in seed_map:
+                for seed in seed_list:
                     raider_pipeline('phRAIDER', file, seed, f)
             if args.run_prephraider:
-                for seed in seed_map:
+                for seed in seed_list:
                     raider_pipeline('pre-phRAIDER', file, seed, f)
             

### 2015-09-13 testing_pipeline.py
Update testing_pipeline.py

@@ -260,10 +260,12 @@ def setup():
             seed_map[convert_seed(args.seed)] = (offset,args.seed)
             seed_map2[convert_seed2(args.seed)] = (offset,args.seed)
 
+    with open(args.results_dir + "/seed_file.txt", "w") as fp:
+        fp.write("\n".join([str(x[0]) + "\t" + x[1] for x in sorted(seed_map.values(), key = lambda x: x[0]])))
+            
     global seed_list
     seed_list = sorted(seed_map2.values(), key = lambda x: x[0])
-    with open(args.results_dir + "/seed_file.txt", "w") as fp:
-        fp.write("\n".join([str(x[0]) + "\t" + x[1] for x in seed_list]))
+
 
 
     # Create data_files.txt

### 2015-09-13 testing_pipeline.py
Adding testing_pipeline

@@ -137,6 +137,7 @@ def parse_params():
     blast_arguments.add_argument('--evalue', dest = 'evalue', help = "BLASE evalue", default = "0.000001");
     blast_arguments.add_argument('--short', dest = 'short', action = 'store_false', help = "Turn off blast-short on blast run", default = True)
     blast_arguments.add_argument('--max_target', dest = 'max_target', action = 'store', help = "BLAST --max_target option", default = '999999999') # HACK!!!  Need to fix this
+    blast_arguments.add_argument('--nt', '--num_threads', dest = 'num_threads', action = 'store', type = int', "BLAST --num_threads option', default = 1
 
     # DEBUGGING ARGUMENTS
     debug_group = parser.add_argument_group(title = "debugging")
@@ -286,7 +287,7 @@ raider_cmd = "/usr/bin/time {raider} -q -c {f} -s {seed} {input_file} {output_di
 consensus_cmd = "{python} consensus_seq.py -s {data_file} -e {elements_file} {consensus_txt} {consensus_fa}"
 repeat_masker_cmd = "{RepeatMasker} -nolow -lib {library_file} -pa {pa} -dir {output_dir} {seq_file}"
 blast_format = "6 qseqid sseqid qstart qend qlen sstart send slen"
-blast_cmd = "{blast} -out {output} -outfmt \"{blast_format}\" -query {consensus_file} -db {db_file} -evalue {evalue} {short} -max_target_seqs {max_target}"
+blast_cmd = "{blast} -out {output} -outfmt \"{blast_format}\" -query {consensus_file} -db {db_file} -evalue {evalue} {short} -max_target_seqs {max_target} --num_threads {num_threads}"
 
 def raider_pipeline(raider_exe, input_file, seed, f):
     ##########################
@@ -342,9 +343,9 @@ def raider_pipeline(raider_exe, input_file, seed, f):
     # Step 4: Apply blast to consensus sequences:
     if args.run_blast:
         cmd4 = blast_cmd.format(blast = Locations['blast'], blast_format = blast_format, output = blast_output, consensus_file = consensus_fa, db_file = database_file, 
-                                evalue = args.evalue, short = "-task blastn-short" if args.short else "", max_target = args.max_target)
+                                evalue = args.evalue, short = "-task blastn-short" if args.short else "", max_target = args.max_target, num_threads = args.num_threads)
         title4 = "bl." + title
[…]

### 2015-09-13 testing_pipeline.py
Updating testing_pipeline.py

@@ -137,7 +137,6 @@ def parse_params():
     blast_arguments.add_argument('--evalue', dest = 'evalue', help = "BLASE evalue", default = "0.000001");
     blast_arguments.add_argument('--short', dest = 'short', action = 'store_false', help = "Turn off blast-short on blast run", default = True)
     blast_arguments.add_argument('--max_target', dest = 'max_target', action = 'store', help = "BLAST --max_target option", default = '999999999') # HACK!!!  Need to fix this
-    blast_arguments.add_argument('--nt', '--num_threads', dest = 'num_threads', action = 'store', type = int', "BLAST --num_threads option', default = 1
 
     # DEBUGGING ARGUMENTS
     debug_group = parser.add_argument_group(title = "debugging")
@@ -255,19 +254,19 @@ def setup():
                 seed_map2[seed] = (offset,line.rstrip())
                 offset += 1
             else:
-                seed_map2[seed] = (offset,line.rstrip())
+                seed_map2[seed] = (seed_map[seed][0],line.rstrip())
     else:
         if args.seed not in seed_map:
             seed_map[convert_seed(args.seed)] = (offset,args.seed)
             seed_map2[convert_seed2(args.seed)] = (offset,args.seed)
 
     with open(args.results_dir + "/seed_file.txt", "w") as fp:
-        fp.write("\n".join([str(x[0]) + "\t" + x[1] for x in sorted(seed_map.values(), key = lambda x: x[0]])))
+        fp.write("\n".join([str(x[0]) + "\t" + x[1] for x in sorted(seed_map.values(), key = lambda x: x[0])]))
             
     global seed_list
-    seed_list = sorted(seed_map2.values(), key = lambda x: x[0])
-
-
+    seed_list = sorted(seed_map2.keys(), key = lambda v: seed_map2[v][0])
+    print(seed_map2)
+    print(seed_list)
 
     # Create data_files.txt
     data_files = args.results_dir + "/data_files.txt"
@@ -287,7 +286,7 @@ raider_cmd = "/usr/bin/time {raider} -q -c {f} -s {seed} {input_file} {output_di
 consensus_cmd = "{python} consensus_seq.py -s {data_file} -e {elements_file} {consensus_txt} {consensus_fa}"
 repeat_masker_cmd = "{RepeatMasker} -nolow -lib {library_file} -pa {pa} -dir {output_dir} {seq_file}"
 blast_format = "6 qseqid sseqid qstart qend qlen sstart send slen"
[…]

### 2015-09-13 testing_pipeline.py
testing_pipeline.py

@@ -137,7 +137,8 @@ def parse_params():
     blast_arguments.add_argument('--evalue', dest = 'evalue', help = "BLASE evalue", default = "0.000001");
     blast_arguments.add_argument('--short', dest = 'short', action = 'store_false', help = "Turn off blast-short on blast run", default = True)
     blast_arguments.add_argument('--max_target', dest = 'max_target', action = 'store', help = "BLAST --max_target option", default = '999999999') # HACK!!!  Need to fix this
-
+    blast_arguments.add_argument('--nt', '--num_threads', dest = 'num_threads', action = 'store', type = int, help = "BLAST --num_threads argument", default = 1)
+    
     # DEBUGGING ARGUMENTS
     debug_group = parser.add_argument_group(title = "debugging")
     debug_group.add_argument('--sp', '--show_progress', dest = 'show_progress', action = 'store_true', help = "Print reports on program progress to stderr", default = False)
@@ -286,7 +287,7 @@ raider_cmd = "/usr/bin/time {raider} -q -c {f} -s {seed} {input_file} {output_di
 consensus_cmd = "{python} consensus_seq.py -s {data_file} -e {elements_file} {consensus_txt} {consensus_fa}"
 repeat_masker_cmd = "{RepeatMasker} -nolow -lib {library_file} -pa {pa} -dir {output_dir} {seq_file}"
 blast_format = "6 qseqid sseqid qstart qend qlen sstart send slen"
-blast_cmd = "{blast} -out {output} -outfmt \"{blast_format}\" -query {consensus_file} -db {db_file} -evalue {evalue} {short} -max_target_seqs {max_target}"
+blast_cmd = "{blast} -out {output} -outfmt \"{blast_format}\" -query {consensus_file} -db {db_file} -evalue {evalue} {short} -max_target_seqs {max_target} -num_threads {num_threads}"
 
 def raider_pipeline(raider_exe, input_file, seed, f):
     ##########################
@@ -342,9 +343,9 @@ def raider_pipeline(raider_exe, input_file, seed, f):
     # Step 4: Apply blast to consensus sequences:
     if args.run_blast:
         cmd4 = blast_cmd.format(blast = Locations['blast'], blast_format = blast_format, output = blast_output, consensus_file = consensus_fa, db_file = database_file, 
-                                evalue = args.evalue, short = "-task blastn-short" if args.short else "", max_target = args.max_target)
+                                evalue = args.evalue, short = "-task blastn-short" if args.short else "", max_target = args.max_target, num_threads = args.num_threads)
[…]

### 2015-09-13 testing_pipeline.py
Updating testing_pipeline

@@ -266,8 +266,6 @@ def setup():
             
     global seed_list
     seed_list = sorted(seed_map2.keys(), key = lambda v: seed_map2[v][0])
-    print(seed_map2)
-    print(seed_list)
 
     # Create data_files.txt
     data_files = args.results_dir + "/data_files.txt"
@@ -345,7 +343,7 @@ def raider_pipeline(raider_exe, input_file, seed, f):
         cmd4 = blast_cmd.format(blast = Locations['blast'], blast_format = blast_format, output = blast_output, consensus_file = consensus_fa, db_file = database_file, 
                                 evalue = args.evalue, short = "-task blastn-short" if args.short else "", max_target = args.max_target, num_threads = args.num_threads)
         title4 = "bl." + title
-        p4 = launch_job(cmd=cmd4, title=title4, base_dir=elements_dir, modules=Locations['blast_modules'], depend=[p2], attrs={'blast_output':blast_output}, ppn = args.num_thrreads)
+        p4 = launch_job(cmd=cmd4, title=title4, base_dir=elements_dir, modules=Locations['blast_modules'], depend=[p2], attrs={'blast_output':blast_output}, ppn = args.num_threads)
 
 
 

### 2015-09-14 manuscript/JCB/bmc_article.tex
Manuscript update

--- 2015-07-28 manuscript/BMC/bmc_article.tex
@@
 \section*{Background}
-Transposable elements, or TEs, are genomic sequences that have at some
+Transposable Elements (TEs) are genomic sequences that have at some
 point had the capacity to insert copies of themselves into other
 genomic locations, resulting in homologous families of sequences
-spread across the genome.  TEs are present in almost every higher
-order genome, covering as much as 45\% of the human genome and 90\% of
-the maize genome \cite{Venter:2001p92,SanMiguel:1996wa}.  TEs have
-proved an important source of data in a number of studies
-\cite{Arndt:2005p279,Karro:2008p123,Mugal:2009p581,Hardison:2003p97},
-but given their prevalence, it is just important for those studying
+spread across the genome.  Present in almost every higher
+order genome (covering as much as 45\% of the human genome and 90\% of
+the maize genome \cite{Venter:2001p92,SanMiguel:1996wa}), TEs have
+proved an important source of data in numerous studies of genomic structure (e.g.
+\cite{Arndt:2005p279,Karro:2008p123,Mugal:2009p581,Hardison:2003p97}).
+But given their prevalence, it is important for those studying other aspect
 the genome to have TEs masked out -- their bases replaced by N to
 allow for easy identification and filtering.  Failure to filter can
-reek havoc with genomic analysis tools.  For example. if not filtered,
+reek havoc with genomic analysis tools.  For example,
 TEs can trigger huge numbers of false positives automated gene finding
@@
 
-The most well known tools for repeat identification is RepeatMasker,
-with the new nHMMER showing promise as well)
-\cite{RepeatMaskerOpen:XkNxXSd7,Wheeler:2013gj}.  Both work on the
-principle of a library-based search.  That is, to identify
-transposable elements to a known family, the tools require some
+The best tools for repeat identification are RepeatMasker and nHMMER
+\cite{RepeatMaskerOpen:XkNxXSd7,Wheeler:2013gj}, but both employ a
+library-based search strategy.  They require some pre-compiled
 description of sequences in the family (e.g. a ancestral sequence for
-BLASTing, or a profile HMM description) that is used as the basis for
-identification.  But, much like we ask how the snow plow driver gets
-to work, or how the author of a dictionary authors look up the spelling of a word
-\cite{Pratchett:uw}, we must ask how the libraries on which the tools
-are based are initially compiled.  Library-based tools are useless for the
-discovery of new families, and cannot be applied to a newly sequenced
-genome for which such a library has not been compiled.
+BLASTing, or a profile HMM) that is used as the basis for
+identification of new member of the family.  But, much like we ask how
+the snow plow driver gets to work \cite{Pratchett:uw}, we must ask how
+these libraries are based are initially compiled.  Library-based tools
+are useless for the discovery of new families, and cannot be applied
+to a newly sequenced genome for which such a library has not been
+compiled.
 
-Within mammalian species we can largely rely on homology
-relationships to port libraries across species.  A library description
-of a human ALU sequence can likely be used for the initial annotation
-of a newly sequenced primate genome, and many human LINEs will extend to
-rodent, or even more distant, mammalian species.  This does not hold
-so well in plants: in many cases TE composition of a given plant organism
-is species-specific.  For example, a rice-based TE library will only identify 25\%
-of the TEs in the maize genome \cite{Jiang:2013jt}.
+Within mammalian species we can largely rely on homology relationships
+to port libraries across species.  This does not hold
+so well in plants: in many cases TE composition of a given plant
+organism is species-specific.  For example, a rice-based TE library
+will only identify 25\% of the TEs in the maize genome
+\cite{Jiang:2013jt}.
 
 To solve this problem we turn to {\it de novo} TE identification
-tools: tools identifying TEs though a genome using only the genome
+tools, identifying TEs using only the genome
 sequence information.  A number of such tools are discussed in the
-literature, operating on a number of different principles.  Tools such
-as RECON and PILER are based on
+literature.   RECON and PILER are based on
 self-alignment, using WU-BLAST and LASTZ for the alignment
@@
 computationally intensive and infeasible for use on whole genomes
-(requiring 60 hours for 18Mb rice genome int 2013), while
+(requiring 60 hours for 18Mb rice genome in a 2013 study), while
 PILER achieves a good runtime with very low sensitivity
@@
 formal description.  For a given genome, a nucleotide sequence $r$ is
-an elementary repeat if:
-\begin{enumerate}
-\item It is of at least length $l$.  (The length requirement.)
-\item At least $f$ exact copies of $r$ appear in the genome.  (The
-  frequency requirement.)
-\item There is no proper substring of $r$ that is of at least length
-  $l$ and appears in the genome independently of $r$. (The minimality
-  requirement.)
-\item There is no string that contains $r$, appears in the genome
-  the same number of times as $r$, and satisfies conditions 1-3.  (The
-  maximality requirement.)
-\end{enumerate}
-The first two criteria are straight-forward, allowing us to designate
-length and frequency thresholds for our model.  The minimality
-requirement ensures that an elementary repeat does not contain
-anything that could be viewed as an independent elementary repeat --
-that it is the ``smallest building block'' that we can find.  While
-the maximality requirement ensures that it is not tied to some other, larger,
-elementary repeat.
-
-Having proposed this definition, Zheng and Lonardi developed an
+an elementary repeat if: {\bf (1)} It is of at least length $l$ (the
+length requirement); {\bf (2)} there are at least $f$ copies of $r$
+appear (the frequency requirement) {\bf (3)} there is no proper
+substring of $r$ of length $\leq l$ that appears in the genome
+independently of $r$ (the minimality requirement); {\bf (4)} $r$ is a
+maximal string w.r.t (1-3) (the maximality requirement).  Having
+proposed this definition, Zheng and Lonardi developed an
 identification algorithm that had a runtime quadratic in the genome
-size -- so not of practical use on whole-genome analysis.  This was
-refined to linear time by He and also by Huo {\it et al.}
-\cite{He:2006gpa,Huo:2009hoa} based on variations of suffix tree
-approaches, but are limited
-in their ability to handle sequence variation.  As transposable
-elements naturally suffer from copy mistakes and accumulate
-instance-specific base substitutions over time, this is a significant
-limitation.
+size. This was refined to linear time by He and also by Huo {\it et
+  al.}  \cite{He:2006gpa,Huo:2009hoa} based on variations of suffix
+tree approaches, but are limited in their ability to handle sequence
+variation.  As we are looking a genome size inputs, and TEs
+transposable elements naturally suffer from copy mistakes and
+accumulate instance-specific base substitutions over time, this is a
+significant limitation.
 
@@
 Figueroa {\it et al.}, with more details in the Figueroa masters
-thesis \cite{Figueroa:2014uk,Figueroa:2013cz}.  Not itself fully able
-to handle the spaced seed component, RAIDER was built along an
+thesis \cite{Figueroa:2014uk,Figueroa:2013cz}.  RAIDER was built along an
 alternate, but equivalent definition of elementary repeats based on
@@
 that the minimality condition could be rewritten as: 
-There is no $l$-mer contained with $r$ that appears in the genome
-more times than $r$.  From there we make our four core observations that form
+There is no $l$-mer contained within elementary repeat $r$ that appears in the genome
+more times than $r$.  From there we make four core observations that form
 the basis for the RAIDER algorithm:
@@
 
-Results of the preliminary implementation were promising.  Based on
-the Saha {\it et al.} analysis putting RepeatScout as the top {\it de
-  novo} search tool, comparisons were made against that tool.  RAIDER result
-quality was close, but slightly behind -- with runtime
-order of magnitudes better.  On human chromosome 22 we saw a
-$12\times$ speedup from RepeatScout to RAIDER (2344 seconds to 192 seconds), while coverage
-of the RepBase  \cite{Jurka:2005bl} ancestral sequence actually improved (77\% to 84\%), while on mouse chromosome 19 we saw 
-the same speedup with a significant drop in coverage (53\% to 30\%).  On
-the full human genome RAIDER ran in $6.3$ hours, while RepeatScout was
-unable to complete it run.  For details, see Figueroa {\it et al.} \cite{Figueroa:2014uk}.
+Results of the preliminary implementation were promising. On human
+chromosome 22 we saw a $12\times$ speedup over RepeatScout to RAIDER
+(2344 seconds to 192 seconds), while coverage of the RepBase
+\cite{Jurka:2005bl} ancestral sequence improved (77\% to 84\%), while
+on mouse chromosome 19 we saw the same speedup with a significant drop
+in coverage (53\% to 30\%).  On the full human genome RAIDER ran in
+$6.3$ hours, while RepeatScout was unable to complete it run.  For
+details, see Figueroa {\it et al.} \cite{Figueroa:2014uk}.
 
@@
 
-PatternHunter, first proposed as a very successful augmentation to
+PatternHunter, a very successful augmentation to
 BLAST \cite{Li:2004wl,Altschul:1997p843}, is based on the notion of {\it
-  spaced seeds}.  In the context of BLAST, instead of looking for a
-contiguous substring match to trigger the alignment extension,
-PatternHunter uses a more refined match pattern.  For example, the
-seed 1111110111111 would specify that we were looking for an exact
-match of two length-six substrings separated by a single base (which
-might or might not match).  In the toy seed $s=110101$, we would say the
-strings $AACACA$ and $AAGACA$ match because all miss-matched positions correspond to a 
-0 match.  The string $AACACA$ does not math $TACACA$ (with respect to
-$s$), as the single miss-match does not correspond to the 1.  Nor does
-$AACACA$ match any substring of $TTAACTCA$, though it does match a
-substring of $T\mathbf{AAGATA}C$.
+  spaced seeds}: improving the sensitivity of string matching based
+algorthms by allowing wild-cards in the match.  That is, instead of
+requiring two strings matching in 12 consecutive characters, we might
+instead require two six-character exact matches seperated by one base
+(represented by the {\it seed pattern} 11111101111111), or perhaps
+three consecutive four-character exact matches sperated by two bases
+each (1111001111001111). It has been demonstrated that certain seed
+patterns can idunce significant imporvements in BLAST sensitivity with
+out time penalty, though what makes a good pattern is not well understood.
 
-The incorporation of spaced seeds into BLAST lead to significant
-improvements at virtually no cost in runtime \cite{Li:2004wl}.
-In our context, we will use them to allow limited
-miss-matches between instances of ``the same'' elementary repeat.  The
-preliminary version of RAIDER from Figueroa {\it et al.} made some
-attempt to incorporate this.  It was successful enough to provide a
-proof-of-concept, but a significant amount of work was left to be
-done to allow for true space seed usage.
-
-In this paper we present phRAIDER (PatternHunter-based RAIDER), a fast
+RAIDER was designed with the intent of employing the spaced seed
+strategy, but this was only implemented heuristically for the Figuera
+{\it et al.} paper \cite{Figueroa:2013cz} -- serving primarily as a
+proof-of-concept.  Since its release we have develoed a formal model
+of elementary repeats that incorporated spaced seeds, and from that
+developed phRAIDER (PatternHunter-based RAIDER).  phRAIDER is a fast
 tool for the identification and making of transposible elements in
-both assembled and unassembled genomes. Code  may be obtained [NEED WEB ADDRESS].  In our Methods section
+both assembled and unassembled genomes aht outpreforms RptScout and
+other establihsed tools. Code is free available under the Gnu GPL
+lisence (v. 3) and  may be obtained [NEED WEB ADDRESS].  In our Methods section
 we present the new model that allows us to extend RAIDER to correctly
@@
 
-\section*{Methods}
+\section*{Model}
 
+Our goal is to redefine the concept of transposbile elements to
+accomodate a spaced seed strategy.  In the following section we will describe our
+identification algorithm, and then describe the success in using it to
+mase transposable elements. But we will start here with a brief
+outline of our theoretical model (with a more detailed description in
+the appendix).
+
+\subsection*{phRAIDER}
+
+In order to adapt the RAIDER algorithm to the new definition, we
+require observations parallel to those on which RAIDER was built.  As
+it turns out, its easy to adopt those.  Consider, for example, the
+first observation: ``An $l$-mer cannot belong to two different
+elementary repeats.''  We have already seen a violation, but can
+eliminate this by changing the definition of {\it belongs}:
+\begin{definition}
+  Given a elementary repeat descriptor $r$ for seed $s$ with length
+  $l$, we say a length $w$ substring \underline{belongs} to $r$ if it
+  matches a member of the generalized decomposition of $r$ (w.r.t. to
+  $s$).
+\end{definition}
+We notice in the above example, the string $AAAAA$ belongs to the
+generalized decomposition, hence could not appear in both elementary
+repeats.  But the string $AAAAT$ does not match any string in the
+generalized decomposition of $r$, and hence was free to belong to the
+descriptor of a ``distinct'' family.  See the Supplementary Materials
+for a proof of all the core observations.
+
+\subsection*{phRAIDER Algorithm}
+
+\red{
+TO DO: Give a pseudocode description of the algoirthm and some minimal
+discussion.
+}
+
+\subsection*{Analysis}
+
+\red{
+TO DO: Describe pra analysis and quality metrics.
+}
+
+\section*{Results}
+
+\red{
+Points to be made:
+\begin{enumerate}
+\item Show: RAIDER2 has better quality masking results then RS.
+\item Show: RAIDER2 is better than RAIDER.
+\item Show: Seeds matter.
+\item Show: Runtime / memory usage
+\end{enumerate}
+
+Possible tables / plots:
+\begin{itemize}
+\item Table showing results on different chromosomes (pra, runtime, memory).
+\item Plot of runtime against genome size (simulated data to control size)?
+\item Plot of memory usage (simlated data to control size)?
+\item We might want to compare the ability to detect transposable
+  elements againsts other types of repeats.  (Should be easy to do
+  using simulator.)
+\end{itemize}
+}
+
+\section*{Conclusions}
+
+\red {
+Hopefully will beconcluding that RAIDER is a better tool than
+RepeatScout.  Posible discussion of tool features.
+}
+
+\section*{Appendix}
 In order to allow the use of a spaced seed strategy, we will need to
@@
 too breaks the algorithm.
-
-\subsection*{phRAIDER}
-
-In order to adapt the RAIDER algorithm to the new definition, we
-require observations parallel to those on which RAIDER was built.  As
-it turns out, its easy to adopt those.  Consider, for example, the
-first observation: ``An $l$-mer cannot belong to two different
-elementary repeats.''  We have already seen a violation, but can
-eliminate this by changing the definition of {\it belongs}:
-\begin{definition}
-  Given a elementary repeat descriptor $r$ for seed $s$ with length
-  $l$, we say a length $w$ substring \underline{belongs} to $r$ if it
-  matches a member of the generalized decomposition of $r$ (w.r.t. to
-  $s$).
-\end{definition}
-We notice in the above example, the string $AAAAA$ belongs to the
-generalized decomposition, hence could not appear in both elementary
-repeats.  But the string $AAAAT$ does not match any string in the
-generalized decomposition of $r$, and hence was free to belong to the
-descriptor of a ``distinct'' family.  See the Supplementary Materials
-for a proof of all the core observations.
-
-\subsection*{phRAIDER Algorithm}
-
-\red{
-TO DO: Give a pseudocode description of the algoirthm and some minimal
-discussion.
-}
-
-\subsection*{Analysis}
-
-\red{
-TO DO: Describe pra analysis and quality metrics.
-}
-
-\section*{Results}
-
-\red{
-Points to be made:
-\begin{enumerate}
-\item Show: RAIDER2 has better quality masking results then RS.
-\item Show: RAIDER2 is better than RAIDER.
-\item Show: Seeds matter.
-\item Show: Runtime / memory usage
-\end{enumerate}
-
-Possible tables / plots:
-\begin{itemize}
-\item Table showing results on different chromosomes (pra, runtime, memory).
-\item Plot of runtime against genome size (simulated data to control size)?
-\item Plot of memory usage (simlated data to control size)?
-\item We might want to compare the ability to detect transposable
-  elements againsts other types of repeats.  (Should be easy to do
-  using simulator.)
-\end{itemize}
-}
-
-\section*{Conclusions}
-
-\red {
-Hopefully will beconcluding that RAIDER is a better tool than
-RepeatScout.  Posible discussion of tool features.
-}
 

### 2015-09-14 manuscript/JCB/manuscript.tex
Manuscript update

--- 2015-09-14 manuscript/JCB/bmc_article.tex
@@
 \section*{Background}
-Transposable Elements (TEs) are genomic sequences that have at some
-point had the capacity to insert copies of themselves into other
-genomic locations, resulting in homologous families of sequences
-spread across the genome.  Present in almost every higher
-order genome (covering as much as 45\% of the human genome and 90\% of
-the maize genome \cite{Venter:2001p92,SanMiguel:1996wa}), TEs have
-proved an important source of data in numerous studies of genomic structure (e.g.
+Transposable Elements (TEs) are genomic sequences thathad the capacity
+to insert copies of themselves into other genomic locations, resulting
+in homologous families of sequences spread across the genome.  Present
+in almost every higher order genome (covering as much as 45\% of the
+human genome and 90\% of the maize genome
+\cite{Venter:2001p92,SanMiguel:1996wa}), TEs have proved an important
+source of data in numerous studies of genomic structure (e.g.
 \cite{Arndt:2005p279,Karro:2008p123,Mugal:2009p581,Hardison:2003p97}).
-But given their prevalence, it is important for those studying other aspect
-the genome to have TEs masked out -- their bases replaced by N to
-allow for easy identification and filtering.  Failure to filter can
-reek havoc with genomic analysis tools.  For example,
-TEs can trigger huge numbers of false positives automated gene finding
-tool \cite{Jiang:2013jt}, as well as inflate tool runtime.
+But given their prevalence, it is important for those studying other
+aspect the genome to have TEs masked out -- their bases replaced by N
+to allow for easy identification and filtering.  Failure to filter can
+reek havoc with genomic analysis tools.  For example, unfiltered TEs can trigger
+huge numbers of false positives in automated gene finding tool
+\cite{Jiang:2013jt}, as well as inflate tool runtime.
 
 The best tools for repeat identification are RepeatMasker and nHMMER
-\cite{RepeatMaskerOpen:XkNxXSd7,Wheeler:2013gj}, but both employ a
-library-based search strategy.  They require some pre-compiled
+\cite{RepeatMaskerOpen:XkNxXSd7,Wheeler:2013gj}, but both employ 
+library-based search strategies using a pre-compiled
 description of sequences in the family (e.g. a ancestral sequence for
-BLASTing, or a profile HMM) that is used as the basis for
-identification of new member of the family.  But, much like we ask how
+BLASTing, or a profile HMM).  But, much like we ask how
 the snow plow driver gets to work \cite{Pratchett:uw}, we must ask how
-these libraries are based are initially compiled.  Library-based tools
-are useless for the discovery of new families, and cannot be applied
-to a newly sequenced genome for which such a library has not been
-compiled.
+these libraries are compiled.  Library-based tools
+are useless for the discovery of new families, and this are
+challanging to use on newly sequenced genomes.
 
@@
 To solve this problem we turn to {\it de novo} TE identification
-tools, identifying TEs using only the genome
-sequence information.  A number of such tools are discussed in the
-literature.   RECON and PILER are based on
-self-alignment, using WU-BLAST and LASTZ for the alignment
-\cite{Bao:2002,Edgar:2005p2365,Lopez:2003td,Harris:2007uf}.  RECON show good sensitivity but is
-computationally intensive and infeasible for use on whole genomes
-(requiring 60 hours for 18Mb rice genome in a 2013 study), while
-PILER achieves a good runtime with very low sensitivity
-\cite{Jiang:2013jt}.  ReAS \cite{Li:2005he} and RepeatScout
-\cite{Price:2005p1247} are based on $k$-mer searches, with the
-earlier showing less sensitivity than RECON \cite{Jiang:2013jt}.
-RepeatGluer \cite{Pevzner:2004p3157,Zhi:2006p3199} is
-based on a variation of DeBrujin graphs, which allows for a
-decomposition of TE families into domains, but is very, very
-computationally expensive.  In Saha {\it et al.} the authors perform
-an extensive comparison of the tools, and conclude that RepeatScout is
-the best tool overall for assembled genomes, while ReAS the best when
-dealing with unassembled sequence fragments \cite{Saha:2008dm}.
+tools, identifying TEs using only the genome sequence information.  A
+number of such tools are discussed in the literature.  RECON uses
+WU-BLAST and PILER using LASTZ to compute self-alignments
+\cite{Bao:2002,Edgar:2005p2365,Lopez:2003td,Harris:2007uf}.  RECON
+show good sensitivity but is computationally intensive and infeasible
+for use on whole genomes (requiring 60 hours for 18Mb rice genome in a
+2013 study), while PILER achieves a good runtime with very low
+sensitivity .  ReAS and RepeatScout \cite{Price:2005p1247} are based
+on $k$-mer searches, with the earlier showing less sensitivity than
+RECON \cite{Li:2005he, Price:2005p1247,Jiang:2013jt}.  RepeatGluer
+\cite{Pevzner:2004p3157,Zhi:2006p3199} is based on a variation of
+DeBrujin graphs, which allows for a decomposition of TE families into
+domains, but is very, very computationally expensive.  In Saha {\it et
+  al.} the authors perform an extensive comparison of the tools, and
+conclude that RepeatScout is the best tool overall for assembled
+genomes, while ReAS the best when dealing with unassembled sequence
+fragments \cite{Saha:2008dm}.
 
 \subsection*{Elementary Repeats}
-Another line of development, first proposed by Zheng and Lonardi
-\cite{Zheng:2005bl}, involves the use of {\it elementary repeats}.
+Zheng and Lonardi approached the {\it de novo} identification problem using {\it elementary repeats}
+\cite{Zheng:2005bl}.
 Similar to the RepeatGluer domains, elementary repeats are
@@
 formal description.  For a given genome, a nucleotide sequence $r$ is
-an elementary repeat if: {\bf (1)} It is of at least length $l$ (the
-length requirement); {\bf (2)} there are at least $f$ copies of $r$
-appear (the frequency requirement) {\bf (3)} there is no proper
-substring of $r$ of length $\leq l$ that appears in the genome
-independently of $r$ (the minimality requirement); {\bf (4)} $r$ is a
-maximal string w.r.t (1-3) (the maximality requirement).  Having
-proposed this definition, Zheng and Lonardi developed an
-identification algorithm that had a runtime quadratic in the genome
-size. This was refined to linear time by He and also by Huo {\it et
-  al.}  \cite{He:2006gpa,Huo:2009hoa} based on variations of suffix
-tree approaches, but are limited in their ability to handle sequence
+an elementary repeat if: (1) It is of at least length $l$ (the length
+requirement); (2) there are at least $f$ copies of $r$ appear (the
+frequency requirement) (3) there is no proper substring of $r$ of
+length $\leq l$ that appears in the genome independently of $r$ (the
+minimality requirement); (4) $r$ is a maximal string w.r.t (1-3) (the
+maximality requirement).  Having proposed this definition, Zheng and
+Lonardi developed an identification algorithm that had a runtime
+quadratic in the query sequence size \cite{Zheng:2005bl}. This was
+refined to linear time by He and also by Huo {\it et al.}
+\cite{He:2006gpa,Huo:2009hoa} based on variations of suffix tree
+approaches, but these appraoches are limited in their ability to handle sequence
 variation.  As we are looking a genome size inputs, and TEs
@@
 Figueroa {\it et al.}, with more details in the Figueroa masters
-thesis \cite{Figueroa:2014uk,Figueroa:2013cz}.  RAIDER was built along an
-alternate, but equivalent definition of elementary repeats based on
+thesis \cite{Figueroa:2014uk,Figueroa:2013cz}.  RAIDER was built along
+an alternate, but equivalent definition of elementary repeats based on
 $l$-mers (sequences of length $l$).  Specifically, it was observed
-that the minimality condition could be rewritten as: 
-There is no $l$-mer contained within elementary repeat $r$ that appears in the genome
-more times than $r$.  From there we make four core observations that form
-the basis for the RAIDER algorithm:
-\begin{enumerate}
-\item An $l$-mer cannot belong to two different elementary repeats.
-\item Any $l$-mer in the genome that occurs $f$ or more times is
-  either an elementary repeat or belongs to one.
-\item Any two $l$-mers belonging to an elementary repeat must appear
-  the same number of times in the genome. 
-\item If two sequences in the genome that are {\it maximally
-  identical} (that is, cannot be extended in either direction and still
-  be identical), these sequences cannot belong to a larger
-  elementary repeat.
-\end{enumerate}
-By ``belong'' we mean ``is a substring of''; the reason for the
-terminology will be explained shortly.  For discussion and proof, see
-the Figueroa Thesis \cite{Figueroa:2013cz}.
+that the minimality condition could be rewritten as: There is no
+$l$-mer contained within elementary repeat $r$ that appears in the
+genome more times than $r$.  From there we make four core observations
+that form the basis for the RAIDER algorithm: (1) An $l$-mer cannot
+belong to two different elementary repeats; (2) Any $l$-mer in the
+genome that occurs $f$ or more times is either an elementary repeat or
+belongs to one; (3) any two $l$-mers belonging to an elementary repeat
+must appear the same number of times in the genome; (4) If two
+sequences in the genome that are {\it maximally identical} (that is,
+cannot be extended in either direction and still be identical), these
+sequences cannot belong to a larger elementary repeat.  (By ``belong''
+we mean ``is a substring of'' -- a definition we will be generalizing
+shortly.)  For discussion and proof, see the Figueroa Thesis
+\cite{Figueroa:2013cz}.
 
@@
 
-PatternHunter, a very successful augmentation to
-BLAST \cite{Li:2004wl,Altschul:1997p843}, is based on the notion of {\it
+PatternHunter, a very successful augmentation to BLAST
+\cite{Li:2004wl,Altschul:1997p843}, is based on the notion of {\it
   spaced seeds}: improving the sensitivity of string matching based
@@
 instead require two six-character exact matches seperated by one base
-(represented by the {\it seed pattern} 11111101111111), or perhaps
-three consecutive four-character exact matches sperated by two bases
-each (1111001111001111). It has been demonstrated that certain seed
-patterns can idunce significant imporvements in BLAST sensitivity with
-out time penalty, though what makes a good pattern is not well understood.
+which may or may not match (represented by the {\it seed pattern}
+11111101111111), or perhaps three consecutive four-character exact
+matches sperated by two bases each (1111001111001111). It has been
+demonstrated that certain seed patterns can indunce significant
+imporvements in BLAST sensitivity with out time penalty, though what
+makes a good pattern is not well understood.
 
@@
 other establihsed tools. Code is free available under the Gnu GPL
-lisence (v. 3) and  may be obtained [NEED WEB ADDRESS].  In our Methods section
+lisence (v. 3) and  may be obtained [NEED WEB ADDRESS].
+
+In the following we will present a new theoretical model for {\it
+  seeded elementary repeats}, followed by an algorithm for identifying
+them.  In the results section we will show the result of using that
+algorithm for masking transp
+
+In our Methods section
 we present the new model that allows us to extend RAIDER to correctly
@@
 Our goal is to redefine the concept of transposbile elements to
-accomodate a spaced seed strategy.  In the following section we will describe our
-identification algorithm, and then describe the success in using it to
-mase transposable elements. But we will start here with a brief
+accomodate a spaced seed strategy.  In the next section we will describe our
+identification algorithm, and then quantified its success in 
+masking transposable elements. But we will start here with a brief
 outline of our theoretical model (with a more detailed description in
@@
 
-\subsection*{phRAIDER}
+We first need to redfine out terminimology regarding elementary
+repeats.  Under the Z\&L definition, all instances of one elementary
+repeat have the exact same sequence, and hence can be descrbied by a
+single string.  As we are allowing for variation, we will describe out
+instance set with a {\it sequence descriptor} consisting of bases
+letters and the wild-card character * (e.g. AAC*GG would describe a
+set of sequences with starting with an AAC, ending with a GG, and
+having any character in between).  Given a binary string $s$
+representing a spaced seed, we say a sequence descriptor $r$ is
+{\it consistent} with $s$ if we can align $s$ to a substring of $r$ such
+that every $*$ in $r$ aligns with a $0$ character in $s$. (Hence
+$ACG**T*A$ is consistent with the seed $11001$, but not $11011$.)
 
-In order to adapt the RAIDER algorithm to the new definition, we
-require observations parallel to those on which RAIDER was built.  As
-it turns out, its easy to adopt those.  Consider, for example, the
-first observation: ``An $l$-mer cannot belong to two different
-elementary repeats.''  We have already seen a violation, but can
-eliminate this by changing the definition of {\it belongs}:
-\begin{definition}
-  Given a elementary repeat descriptor $r$ for seed $s$ with length
-  $l$, we say a length $w$ substring \underline{belongs} to $r$ if it
-  matches a member of the generalized decomposition of $r$ (w.r.t. to
-  $s$).
+Given a sequence descriptor $r$, we can {\it decompose} $r$ with
+respect to $s$ by taking every length $|s|$ substring of $r$ that is consistant
+with $s$, replacing all letters of $r$ that match to a 0 in $s$ with a
+$*$, and creating a set from the results.  (Hence for $s = 11011$ and
+$r=AA*CCGTT$, the decomposition would be $\{AA*CC, CC*TT\}$.)  We say
+$s$ {\it covers} $r$ is every base in $r$ is contained in at least on
+string in the decompositio.  (Hence the previous $s$ does not cover $AAA*CC$, as
+the first base in not in any of the strings of the decomposition.)
+
+We can now modif the previous definition of elementary repeats as
+follows:
+\begin{definition} Given a genomic sequence $G$, an integer $f$, and a
+  spaced seed $s$, a sequence descriptor $r$ describes an elementary
+  repeat if it meets the four (moified) requirements of an elementary repeat:
+\begin{itemize}
+\item \underline{Structure requriement}: $s$ covers $r$.
+\item \underline{Frequency requirement}: There are at least $f$
+  substrings of $G$ that match $r$.
+\item \underline{Minimality requirement:} For every string $t$ in the
+  decomposition of $r$ w.r.t $s$, the number of occurences of $t$ in
+  $G$ is equal to the number of occurneces of $r$ in the genome.
+\item \underline{Maximality requirement:} There is no sequence
+  descriptor $r'$ of $r$ that contains $r$ as a proper substring and
+  satisfies conditions 1-1.
+\end{itemize}
 \end{definition}
-We notice in the above example, the string $AAAAA$ belongs to the
-generalized decomposition, hence could not appear in both elementary
-repeats.  But the string $AAAAT$ does not match any string in the
-generalized decomposition of $r$, and hence was free to belong to the
-descriptor of a ``distinct'' family.  See the Supplementary Materials
-for a proof of all the core observations.
+
+\begin{theorem} When the seed $s$ has no 0 characters, this
+    definition of elementary repeats is equivilent to the Z\&L
+    definition.
+\end{theorem}
 

### 2015-09-15 analysis2.R
Updated testing pipeline: Added /var/folders/4y/cc6kbbq16g10q22_hskf1h8rlsjht9/T/ usage (semi-tested)

@@ -18,10 +18,13 @@ read.stats <- function() {
   stats10 <<- read.file("seeds10.stats.txt", FALSE, 10)
   stats11 <<- read.file("seeds11.stats.txt", TRUE,  11)  
   stats12 <<- read.file("seeds12.stats.txt", FALSE, 12)
-  stats13 <<- read.file("seeds12.stats.txt", FALSE, 13)
+  stats13 <<- read.file("seeds13.stats.txt", FALSE, 13)
+  stats15 <<- read.file("seeds15.stats.txt", FALSE, 15)
   stats.all <<-rbind(stats1, stats2, stats3, stats4, stats5, 
                      stats6, stats7, stats8, stats9, stats10, 
                      stats11, stats12, stats13)
+  statsM1 <<- read.file("seeds1m.stats.txt", FALSE, 1)
+  best <<- read.file("best.stats.txt", FALSE, -1)
 }
 
 add.legend <- function(position) {

### 2015-09-15 seed_fun.py

@@ -38,10 +38,13 @@ def weight(key):
     return sum([1 for s in key if s == "1"])
 
 def check_weights(file):
+    i = 0
     for line in open(file):
         seed = expand_seed(line.rstrip())
         w = weight(seed)
         print(len(seed), w, float(w)/len(seed))
+        i += 1
+    return i
 
 def gen_seed(L):
     s = ""
@@ -102,7 +105,7 @@ def generate_list(w, p_lower, p_upper, p_step, n):
     return sorted(M.keys(), key=len)
         
 def main():
-    fp = open("seeds14.txt", "w")
-    L = generate_list(40, 0.80, 0.99, 0.02, 5)
+    fp = open("seeds15.txt", "w")
+    L = generate_list(40, 0.70, 0.91, 0.02, 20)
     fp.write("\n".join([compact_seed(s) for s in L]))
     fp.close()

### 2015-09-15 testing_pipeline.py

@@ -259,7 +259,7 @@ def setup():
     else:
         if args.seed not in seed_map:
             seed_map[convert_seed(args.seed)] = (offset,args.seed)
-            seed_map2[convert_seed2(args.seed)] = (offset,args.seed)
+            seed_map2[convert_seed(args.seed)] = (offset,args.seed)
 
     with open(args.results_dir + "/seed_file.txt", "w") as fp:
         fp.write("\n".join([str(x[0]) + "\t" + x[1] for x in sorted(seed_map.values(), key = lambda x: x[0])]))
@@ -283,9 +283,10 @@ def setup():
         
 raider_cmd = "/usr/bin/time {raider} -q -c {f} -s {seed} {input_file} {output_dir}"
 consensus_cmd = "{python} consensus_seq.py -s {data_file} -e {elements_file} {consensus_txt} {consensus_fa}"
-repeat_masker_cmd = "{RepeatMasker} -nolow -lib {library_file} -pa {pa} -dir {output_dir} {seq_file}"
+repeat_masker_cmd = "cd $TMPDIR; {RepeatMasker} -nolow -lib {library_file} -pa {pa} -dir $TMPDIR {seq_file}; mv $TMPDIR/{seq_file_base}.out {output_dir}/"
 blast_format = "6 qseqid sseqid qstart qend qlen sstart send slen"
-blast_cmd = "{blast} -out {output} -outfmt \"{blast_format}\" -query {consensus_file} -db {db_file} -evalue {evalue} {short} -max_target_seqs {max_target} -num_threads {num_threads}"
+blast_cmd = "{blast} -out $TMPDSK/{blast_file} -outfmt \"{blast_format}\" -query {consensus_file} -db {db_file} -evalue {evalue} {short} -max_target_seqs {max_target} -num_threads {num_threads}; " + \
+            "bzip2 $TMPDSK/{blast_file}; " + "cp $TMPDSK/{blast_file}.bz2 {blast_dir}/"
 
 def raider_pipeline(raider_exe, input_file, seed, f):
     ##########################
@@ -302,13 +303,13 @@ def raider_pipeline(raider_exe, input_file, seed, f):
 
     consensus_txt = elements_dir + "/" + consensus_name + ".consensus.txt"
     consensus_fa = elements_dir + "/" + consensus_name + ".consensus.fa"
-
     database_file = input_file.rstrip(".fa") + ".rptseq.fa"
[…]

### 2015-09-15 old.testing_pipeline.py
Adding old.testing

@@ -1,0 +1,502 @@
+#!/software/python/3.3.3/bin/python3.3
+import sys
+import os
+import argparse
+import locations
+import subprocess
+import pickle
+from redhawk import *
+
+# Second version of the script for testing RAIDER.
+# Generates runs only; will use a seperate script for testing.
+
+
+
+#######################
+# Globals: So we don't have to pass everything through parameters.
+args = None     # Command line arguments object
+Locations = locations.Locations
+progress_fp = None
+job_log_dir = None
+seed_map = {}         # Maps seed to (index,short rep.) pairs
+seed_list = None      # Sorted list of seeds
+
+#######################
+# Defaults 
+walltime_default = "4:00:00"
+blast_walltime_default = "10:00:00"
+rs_walltime_default = "10:00:00"
+rm_walltime_default = "10:00:00"
+delay_default = 300           # Number of seconds to sleep when cycling on a redhawk wait
+
+tool_prefix = {'phRAIDER' : 'phRA', 'RepeatScout' : 'RS', 'RAIDER': 'RA', 'pre-phRAIDER' : 'prephRA'}
+
+#######################
+# Useful utilitiy functions
+def file_base(file):
+    """Extract the name of a file froma directory"""
+    return os.path.basename(file)
+
+def file_dir(file):
+    """Extract the directory a file is contained in"""
+    return file.rstrip(file_base(file)).rstrip("/")
+
+def make_dir(DIR):
+    if not os.path.exists(DIR):
+        os.makedirs(DIR)
+    return DIR
+
+def convert_seed(seed):
+    """Convert an abriviated seed to a full seed (e.g. "1{2}0{3}1{2}" => "1100011" """
+    i = 0
+    while (i < len(seed)-1):
+        if seed[i+1] == '^':
+            j = i+2
+            assert seed[j] == "{"
+            k = j+1
+            while seed[k] != '}':
+                k += 1
+            n = int(seed[j+1:k])
+            seed = seed[:i] + seed[i]*n + seed[k+1:]
+        i += 1
+    return seed
+
+def seed_weight(seed):
+    return sum([x == 1 for x in seed])
+
+def cmp_seed(s1, s2):
+    d = len(s1) - len(s2)
+    if (d != 0):
+        return d < 0
+    d = weight(s1) - weight(s2)
+    if f != 0:
+        return d < 0
+    return s1 < s2
+
+#######################
+# Command line parsing
+def parse_params():
+    """Parse command line arguments, set them equal to the global args, and set certain global variables"""
[…]

### 2015-09-15 testing_pipeline.py
Updating testing_pipeline.py

@@ -24,6 +24,10 @@ seed_list = None      # Sorted list of seeds
 #######################
 # Defaults 
 walltime_default = "4:00:00"
+blast_walltime_default = "10:00:00"
+rs_walltime_default = "10:00:00"
+rm_walltime_default = "10:00:00"
+
 delay_default = 300           # Number of seconds to sleep when cycling on a redhawk wait
 
 tool_prefix = {'phRAIDER' : 'phRA', 'RepeatScout' : 'RS', 'RAIDER': 'RA', 'pre-phRAIDER' : 'prephRA'}
@@ -121,7 +125,7 @@ def parse_params():
     repscout_argument = parser.add_argument_group("REPSCOUT parameters")
     #repscout_argument.add_argument('--rs_min', type = int, help = "Minimum repeat length for repscout.", default = 10)
     repscout_argument.add_argument('--rs_min_freq', type = int, help = "Minimum repeat frequency for repscout.", default = 3)
-    repscout_argument.add_argument('--rs_walltime', help = "RepeatScout walltime.", default = walltime_default)
+    repscout_argument.add_argument('--rs_walltime', help = "RepeatScout walltime.", default = rs_walltime_default)
     repscout_argument.add_argument('--rs_filters', type = int, help = "0: no filters, 1: 1 filter, 2: both filters.", default = 0)
     
     # REPEAT MASKER ARGUMENTS
@@ -129,7 +133,7 @@ def parse_params():
     repeatmasker_arguments.add_argument('--RM', '--suppress_rm', dest = 'run_rm', action="store_false", help = "suppress RepeatMasker run", default = True)
     repeatmasker_arguments.add_argument('--masker_dir', help = "Repeat masker output directory", default = None)
     repeatmasker_arguments.add_argument('-p', '--pa', type = int, help = "Number of processors will be using", default = 1)    
-    repeatmasker_arguments.add_argument('--rwt', '--rm_walltime', dest = "rm_walltime", help = "Wall time limit for repeat masker", default = "10:00:00")
+    repeatmasker_arguments.add_argument('--rwt', '--rm_walltime', dest = "rm_walltime", help = "Wall time limit for repeat masker", default = rm_walltime_default)
 
     # BLAST ARGUMENTS
     blast_arguments = parser.add_argument_group("BLAST parameters")
@@ -283,7 +287,7 @@ def setup():
         
[…]

### 2015-09-15 testing_pipeline.py
Updating testing_pipeline.py

@@ -28,8 +28,16 @@ delay_default = 300           # Number of seconds to sleep when cycling on a red
 
 tool_prefix = {'phRAIDER' : 'phRA', 'RepeatScout' : 'RS', 'RAIDER': 'RA', 'pre-phRAIDER' : 'prephRA'}
 
+num_jobs = 0;
+
 #######################
 # Useful utilitiy functions
+def tmp_dir():
+    global num_jobs
+    name = "$TMPDIR/karro." + str(num_jobs)
+    num_jobs += 1
+    return name;
+
 def file_base(file):
     """Extract the name of a file froma directory"""
     return os.path.basename(file)
@@ -283,10 +291,10 @@ def setup():
         
 raider_cmd = "/usr/bin/time {raider} -q -c {f} -s {seed} {input_file} {output_dir}"
 consensus_cmd = "{python} consensus_seq.py -s {data_file} -e {elements_file} {consensus_txt} {consensus_fa}"
-repeat_masker_cmd = "cd $TMPDIR; {RepeatMasker} -nolow -lib {library_file} -pa {pa} -dir $TMPDIR {seq_file}; mv $TMPDIR/{seq_file_base}.out {output_dir}/"
+repeat_masker_cmd = "mkdir {TMPDIR}; cd {TMPDIR}; {RepeatMasker} -nolow -lib $PBS_O_WORKDIR/{library_file} -pa {pa} -dir {TMPDIR} $PBS_O_WORKDIR/{seq_file}; mv {TMPDIR}/{seq_file_base}.out $PBS_O_WORKDIR/{output_dir}/; rm -r -f {TMPDIR}"
 blast_format = "6 qseqid sseqid qstart qend qlen sstart send slen"
-blast_cmd = "{blast} -out $TMPDSK/{blast_file} -outfmt \"{blast_format}\" -query {consensus_file} -db {db_file} -evalue {evalue} {short} -max_target_seqs {max_target} -num_threads {num_threads}; " + \
-            "bzip2 $TMPDSK/{blast_file}; " + "cp $TMPDSK/{blast_file}.bz2 {blast_dir}/"
[…]

### 2015-09-16 testing_pipeline.py
Adding testing_pipeline.py

@@ -293,7 +293,7 @@ def setup():
 
 
         
-raider_cmd = "/usr/bin/time {raider} -q -c {f} -s {seed} {input_file} {TMPDIR}; cp {TMPDIR}/elements {output_dir}/; rm -r -f {TMPDIR}"
+raider_cmd = "mkdir {TMPDIR}; /usr/bin/time {raider} -q -c {f} -s {seed} {input_file} {TMPDIR}; cp {TMPDIR}/elements {output_dir}/; rm -r -f {TMPDIR}"
 consensus_cmd = "{python} consensus_seq.py -s {data_file} -e {elements_file} {consensus_txt} {consensus_fa}"
 repeat_masker_cmd = "mkdir {TMPDIR}; cd {TMPDIR}; {RepeatMasker} -nolow -lib $PBS_O_WORKDIR/{library_file} -pa {pa} -dir {TMPDIR} $PBS_O_WORKDIR/{seq_file}; mv {TMPDIR}/{seq_file_base}.out $PBS_O_WORKDIR/{output_dir}/; rm -r -f {TMPDIR}"
 blast_format = "6 qseqid sseqid qstart qend qlen sstart send slen"

### 2015-09-16 testing_pipeline.py
Add testing_pipeline.py

@@ -293,7 +293,7 @@ def setup():
 
 
         
-raider_cmd = "mkdir {TMPDIR}; /usr/bin/time {raider} -q -c {f} -s {seed} {input_file} {TMPDIR}; cp {TMPDIR}/elements {output_dir}/; rm -r -f {TMPDIR}"
+raider_cmd = "mkdir {TMPDIR}; /usr/bin/time {raider} -q -c {f} -s {seed} {input_file} {TMPDIR}; ls {TMPDIR}; cp {TMPDIR}/elements {output_dir}/; rm -r -f {TMPDIR}"
 consensus_cmd = "{python} consensus_seq.py -s {data_file} -e {elements_file} {consensus_txt} {consensus_fa}"
 repeat_masker_cmd = "mkdir {TMPDIR}; cd {TMPDIR}; {RepeatMasker} -nolow -lib $PBS_O_WORKDIR/{library_file} -pa {pa} -dir {TMPDIR} $PBS_O_WORKDIR/{seq_file}; mv {TMPDIR}/{seq_file_base}.out $PBS_O_WORKDIR/{output_dir}/; rm -r -f {TMPDIR}"
 blast_format = "6 qseqid sseqid qstart qend qlen sstart send slen"

### 2015-09-16 collect_stats.py
Updating collect_stats.py

@@ -74,12 +74,10 @@ def load_f_list(DIR):
 
         
 def collectRaider(DIR, tool):
-    print("HERE: " + DIR + " " + tool)
     # FIRST: Get any applicable RM output stats
     for org in data_map.keys():
         for seed_num in seed_map.keys():
             for f in f_list:
-                print(" ".join([org, str(seed_num), str(f)]))
                 RAIDER_job_file = DIR + "/../job_log/{prefix}.{org}.s{seed_num}.f{f}".format(prefix = tool_prefix[tool], org=org, seed_num=seed_num, f=f)        
                 RM_job_file = DIR + "/../job_log/rm.{prefix}.{org}.s{seed_num}.f{f}".format(prefix = tool_prefix[tool], org=org, seed_num=seed_num, f=f)
                 RM_dir = DIR + "/" + ("{org}.s{seed}.f{f}".format(org=org, seed=seed_num, f=f)).upper()
@@ -134,7 +132,7 @@ def collectRaider(DIR, tool):
                     redhawk.storePBS([p], open(RM_job_file, "wb"))
 
                 if os.path.exists(blast_file):
-                    cmd = "./pra_analysis2 {blast_output} {output}".format(blast_output=blast_file, output=pra_output)
+                    cmd = "bzcat {blast_output}.tgz | ./pra_analysis2 {blast_output} {output}".format(blast_output=blast_file, output=pra_output)
                     subprocess.call(re.split("\s+", cmd))
                     query_cover, target_cover, Used = parse_pra_output.parse_pra_output(pra_output, "exclude.txt")
                     H['ConCoverage'], H['QuCoverage'] = query_cover, target_cover

### 2015-09-16 pra_code2/pra_analysis.cpp
Update pra_analysis

@@ -41,10 +41,9 @@ int main(int argc, char** argv) {
     debug = false;
   }
 
-  string blast_file = argv[1];
-  string output_file = argv[2];
+  string output_file = argv[1];
 
-  ifstream fin(blast_file);
+ 
   ofstream fout(output_file);
 
   string qid;
@@ -60,7 +59,7 @@ int main(int argc, char** argv) {
   si_map tmap;
   unordered_map<string,int> lenmap;
   
-  while (fin >> qid >> sid >> qstart >> qend >> qlen >> sstart >> send >> slen) {
+  while (cin >> qid >> sid >> qstart >> qend >> qlen >> sstart >> send >> slen) {
     if (qend < qstart)
       std::swap(qstart, qend);
     if (send < sstart)

### 2015-09-17 collect_stats.py
Fixing bugs in testing_pipeline and collect_stats

@@ -82,7 +82,7 @@ def collectRaider(DIR, tool):
                 RM_job_file = DIR + "/../job_log/rm.{prefix}.{org}.s{seed_num}.f{f}".format(prefix = tool_prefix[tool], org=org, seed_num=seed_num, f=f)
                 RM_dir = DIR + "/" + ("{org}.s{seed}.f{f}".format(org=org, seed=seed_num, f=f)).upper()
                 RM_file = RM_dir + "/" + "{org}.fa.out".format(org=org, seed=seed_num, f=f)
-                blast_file = RM_dir + "/" + "{org}.s{seed}.f{f}.blast.6.txt".format(org=org, seed=seed_num, f=f)
+                blast_file = RM_dir + "/" + "{org}.s{seed}.f{f}.blast.6.txt.bz2".format(org=org, seed=seed_num, f=f)
                 pra_output = RM_dir + "/" + "{org}.s{seed}.f{f}.pra.txt".format(org=org, seed=seed_num, f=f)
 
                 tool_output = RM_file
@@ -131,9 +131,10 @@ def collectRaider(DIR, tool):
                         pass
                     redhawk.storePBS([p], open(RM_job_file, "wb"))
 
+                
                 if os.path.exists(blast_file):
-                    cmd = "bzcat {blast_output}.tgz | ./pra_analysis2 {blast_output} {output}".format(blast_output=blast_file, output=pra_output)
-                    subprocess.call(re.split("\s+", cmd))
+                    cmd = "bzcat {blast_output} | ./pra_analysis2 {output}".format(blast_output=blast_file, output=pra_output)
+                    subprocess.call(cmd, shell=True)
                     query_cover, target_cover, Used = parse_pra_output.parse_pra_output(pra_output, "exclude.txt")
                     H['ConCoverage'], H['QuCoverage'] = query_cover, target_cover
 
@@ -149,7 +150,7 @@ def collectRptScout(DIR, tool):
             RM_job_file = DIR + "/../job_log/rm.{prefix}.{org}.s0.f{f}".format(prefix = tool_prefix[tool], org=org, f=f)
             RS_dir = DIR + "/" + ("{org}.s0.f{f}".format(org=org, f=f)).upper()
             RM_file = RS_dir + "/" + "{org}.fa.out".format(org=org, f=f)
-            blast_file = RS_dir + "/" + "{org}.s0.f{f}.RS.blast.6.txt".format(org=org, f=f)
[…]

### 2015-09-17 testing_pipeline.py

@@ -431,7 +431,7 @@ def rptscout_pipeline(input_file, f):
         
     # Step 3: Run repeatmasker
     if (args.run_rm):
-        cmd3 = repeat_masker_cmd.format(RepeatMasker=Locations['RepeatMasker'], library_file=output, output_dir=rm_dir, seq_file=input_file, seq_file_base = file_base(input_file), pa = args.pa, TMPDIR = tmp_dir())
+        cmd3 = repeat_masker_cmd.format(RepeatMasker=Locations['RepeatMasker'], library_file=output +"/" + rpt_sct_out, output_dir=rm_dir, seq_file=input_file, seq_file_base = file_base(input_file), pa = args.pa, TMPDIR = tmp_dir())
         title3 = "rm." + title 
         p3 = launch_job(cmd=cmd3, title=title3, base_dir=output_dir, walltime = args.rm_walltime, ppn = args.pa, bigmem = False, modules = Locations['rm_modules'], depend=[p2])
     else:
@@ -439,7 +439,7 @@ def rptscout_pipeline(input_file, f):
 
     # Step 4: Apply blast
     if args.run_blast:
-        cmd4 = blast_cmd.format(blast = Locations['blast'], blast_format = blast_format, blast_dir = blast_dir, blast_file = blast_file, consensus_file = output,
+        cmd4 = blast_cmd.format(blast = Locations['blast'], blast_format = blast_format, blast_dir = blast_dir, blast_file = blast_file, consensus_file = output + "/" + rpt_sct_out,
                                 db_file = database_file, evalue = args.evalue, short = "-task blastn-short" if args.short else "", max_target = args.max_target,
                                 num_threads = args.num_threads, TMPDIR = tmp_dir())
         title4 =  "bl." + title 
@@ -450,6 +450,8 @@ def rptscout_pipeline(input_file, f):
 
     # Filter 1
     if args.rs_filters >= 1:
+        sys.stderr("Filter 1 not working")
+        exit(1)
         cmd5 = filter1_cmd.format(filter=Locations['filter_stage-1'], input=output, filter_output = filter1_output)
         title5 = "f1." + title 
         p5 = launch_job(cmd=cmd5, title=title5, base_dir=output_dir1, depend=[p2])

### 2015-09-17 manuscript/JCB/manuscript.tex
Updating manuscript

@@
-%% BioMed_Central_Tex_Template_v1.06
-%%                                      %
-%  bmc_article.tex            ver: 1.06 %
-%                                       %
-
-%%IMPORTANT: do not delete the first line of this template
-%%It must be present to enable the BMC Submission system to
-%%recognise this template!!
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                     %%
-%%  LaTeX template for BioMed Central  %%
-%%     journal article submissions     %%
-%%                                     %%
-%%          <8 June 2012>              %%
-%%                                     %%
-%%                                     %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                                                 %%
-%% For instructions on how to fill out this Tex template           %%
-%% document please refer to Readme.html and the instructions for   %%
-%% authors page on the biomed central website                      %%
-%% http://www.biomedcentral.com/info/authors/                      %%
-%%                                                                 %%
-%% Please do not use \input{...} to include other tex files.       %%
-%% Submit your LaTeX manuscript as one .tex document.              %%
-%%                                                                 %%
-%% All additional figures and files should be attached             %%
-%% separately and not embedded in the \TeX\ document itself.       %%
-%%                                                                 %%
-%% BioMed Central currently use the MikTex distribution of         %%
-%% TeX for Windows) of TeX and LaTeX.  This is available from      %%
-%% http://www.miktex.org                                           %%
-%%                                                                 %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%% additional documentclass options:
-%  [doublespacing]
-%  [linenumbers]   - put the line numbers on margins
-
-%%% loading packages, author definitions
-
-%\documentclass[twocolumn]{bmcart}% uncomment this for twocolumn layout and comment line below
-\documentclass{bmcart}
-
-%%% Load packages
-%\usepackage{amsthm,amsmath}
-%\RequirePackage{natbib}
-%\RequirePackage[authoryear]{natbib}% uncomment this for author-year bibliography
-%\RequirePackage{hyperref}
-\usepackage[utf8]{inputenc} %unicode support
-%\usepackage[applemac]{inputenc} %applemac support if unicode package fails
-%\usepackage[latin1]{inputenc} %UNIX support if unicode package fails
-
-% Added by karro for commenting -- should be removed before submission.
-\usepackage{color}
-\newcommand{\red}[1]{{\color{red}#1}}
-\newcommand{\blue}[1]{{\color{blue}#1}}
-\newcommand{\cyan}[1]{{\color{cyan}#1}}
-\newcommand{\magenta}[1]{{\color{magenta}#1}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                             %%
-%%  If you wish to display your graphics for   %%
-%%  your own use using includegraphic or       %%
-%%  includegraphics, then comment out the      %%
-%%  following two lines of code.               %%
-%%  NB: These line *must* be included when     %%
-%%  submitting to BMC.                         %%
-%%  All figure files must be submitted as      %%
-%%  separate graphics through the BMC          %%
-%%  submission process, not included in the    %%
-%%  submitted article.                         %%
-%%                                             %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
+\documentclass[10pt]{article}
+\usepackage{longtable}
+\usepackage{ulem}
+\usepackage{authblk}
+\usepackage{algorithmic}
 \def\includegraphic{}
 \def\includegraphics{}
+\pagestyle{myheadings}
 
-%%% Put your definitions there:
-\startlocaldefs
-\endlocaldefs
+\setlength{\parindent}{0pt}
+\setlength{\parskip}{1ex plus 0.5ex minus 0.2ex}
 
-%%% Begin ...
+\usepackage[paper=letterpaper,margin=1in]{geometry}
+
+\title{phRAIDER: Pattern-Hunter based Rapid Ad Initio \\ Identification of Elemenrary Repeats}
+\author[1]{Carly E. Schaeffer}
+\author[1]{Nathniel D. Figueroa}
+\author[2]{Xiaolin Liu}
+\author[1,2,3,4]{\\John E. Karro\thanks{Corresponding Author}}
+\affil[1]{Department of Computer Science and Software Engineering}
+\affil[2]{Cell, Molecular, and Structural Bology}
+\affil[3]{Department of Microbiology}
+\affil[4]{Department of Statisticcs, Miami University, Oxford, Ohio (USA)}
+\renewcommand\Affilfont{\itshape\small}
+
+\date{}
+
 \begin{document}
+\maketitle
+\abstract{This is the abstract.}
 
-%%% Start of article front matter
-\begin{frontmatter}
+\section*{Introduction}
 
-\begin{fmbox}
-\dochead{Research}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Enter the title of your article here     %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\title{Fast de novo transposable element annotation}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Enter the authors here                   %%
-%%                                          %%
-%% Specify information, if available,       %%
-%% in the form:                             %%
-%%   <key>={<id1>,<id2>}                    %%
-%%   <key>=                                 %%
-%% Comment or delete the keys which are     %%
-%% not used. Repeat \author command as much %%
-%% as required.                             %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\author[
-   addressref={aff1},
-   noteref={n1},
-   email={schaefce@miamiOH.edu}
-]{\inits{SE}\fnm{Carly E} \snm{Schaeffer}}
-\author[
-   addressref={aff1},                   % id's of addresses, e.g. {aff1,aff2}
-   noteref={n1},                        % id's of article notes, if any
-   email={figuernd@miamiOH.edu}   % email address
-]{\inits{ND}\fnm{Nathaniel D} \snm{Figueroa}}
-\author[
-   addressref={aff1,aff2},
-   email={liux17@miamiOH.edu}
-]{\inits{X}\fnm{Xiaolin} \snm{Liu}}
-\author[
-  addressref={aff1,aff2,aff3,aff4},
-  corref={aff1},
-  noteref={n2},
-  email={karroje@miamiOH.edu}
-]{\inits{JE}\fnm{John E} \snm{Karro}}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Enter the authors' addresses here        %%
-%%                                          %%
-%% Repeat \address commands as much as      %%
-%% required.                                %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\address[id=aff1]{%                           % unique id
-  \orgname{Department of Computer Science and Software Engineering}, % university, etc
-  %\street{Miami University},                     %
-  %\postcode{}                                % post or zip code
-  %\city{Oxford},                              % city
-  %\cny{USA}                                    % country
-}
-\address[id=aff2]{%                           % unique id
-  \orgname{Center for  Molecular and Structural Biology},
-  %\street{Miami University},                     %
-  %\postcode{}                                % post or zip code
-  %\city{Oxford},                              % city
-  %\cny{USA}                                    % country
-}
-\address[id=aff3]{%                           % unique id
-  \orgname{Department of Statistics}, % university, etc
-  %\street{Miami University},                     %
-  %\postcode{}                                % post or zip code
-  %\city{Oxford},                              % city
-  %\cny{USA}                                    % country
-}
-\address[id=aff4]{%                           % unique id
-  \orgname{Department of Microbiology}, % university, etc
-  \street{Miami University},                     %
-  %\postcode{}                                % post or zip code
-  \city{Oxford},                              % city
-  \cny{USA}                                    % country
-}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Enter short notes here                   %%
-%%                                          %%
-%% Short notes will be after addresses      %%
-%% on first page.                           %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{artnotes}
-%\note{Sample of title note}     % note to the article
-  \note[id=n1]{Equal contributor} % note, connected to author
-  \note[id=n2]{Corresponding Author}
-\end{artnotes}
-
-\end{fmbox}% comment this for two column layout
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% The Abstract begins here                 %%
-%%                                          %%
-%% Please refer to the Instructions for     %%
-%% authors on http://www.biomedcentral.com  %%
-%% and include the section headings         %%
-%% accordingly for your article type.       %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{abstractbox}
-
-\begin{abstract} % abstract
-\parttitle{Background} The problem of {\it de novo} identification of
-transposable elements (the discovery and annotation of transposable
-elements without the use of a pre-complied profile) has been addressed
-by a number of tools -- all of which are either based on
-computationally complex algorithms that cannot scale to whole-genome
-use, or whose sensitivity suffers significantly from the presence of
-sequence variation.  Here we present phRAIDER (Pattern Hunter based
-Rapid Ab Initio Detection of Elementary Repeats), a tool that can
-quickly identify and mask tranposable elements in a a newly sequenced
-genome and is robust to the sequence variation present in real data.
-
-\parttitle{Results} \red{To be written later.}
-
-\parttitle{Conclusions} \red{To be written later.}
-
-\end{abstract}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% The keywords begin here                  %%
-%%                                          %%
-%% Put each keyword in separate \kwd{}.     %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{keyword}
-\kwd{transposable elements}
-\kwd{elementary repeats}
-\kwd{pattern hunter}
-\kwd{genomic masking}
-\end{keyword}
-
-% MSC classifications codes, if any
-%\begin{keyword}[class=AMS]
-%\kwd[Primary ]{}
-%\kwd{}
-%\kwd[; secondary ]{}
-%\end{keyword}
-
-\end{abstractbox}
-%
-%\end{fmbox}% uncomment this for twcolumn layout
-
-\end{frontmatter}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% The Main Body begins here                %%
-%%                                          %%
-%% Please refer to the instructions for     %%
-%% authors on:                              %%
-%% http://www.biomedcentral.com/info/authors%%
-%% and include the section headings         %%
-%% accordingly for your article type.       %%
-%%                                          %%
-%% See the Results and Discussion section   %%
-%% for details on how to create sub-sections%%
-%%                                          %%
-%% use \cite{...} to cite references        %%
-%%  \cite{koon} and                         %%
-%%  \cite{oreg,khar,zvai,xjon,schn,pond}    %%
-%%  \nocite{smith,marg,hunn,advi,koha,mouse}%%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\newtheorem{definition}{Definition}
-\newtheorem{observation}{Observation}
-\newtheorem{lemma}{Lemma}
-\newtheorem{theorem}{Theorem}
-
-%%%%%%%%%%%%%%%%%%%%%%%%% start of article main body
-\section*{Background}
 Transposable Elements (TEs) are genomic sequences thathad the capacity
@@
 
-Within mammalian species we can largely rely on homology relationships
-to port libraries across species.  This does not hold
-so well in plants: in many cases TE composition of a given plant
-organism is species-specific.  For example, a rice-based TE library
-will only identify 25\% of the TEs in the maize genome
-\cite{Jiang:2013jt}.
+\begin{algorithmic}
+\If {$i\geq maxval$}
+    \State $i\gets 0$
+\Else
+    \If {$i+k\leq maxval$}
+        \State $i\gets i+k$
+    \EndIf
+\EndIf
+\end{algorithmic}
 
-To solve this problem we turn to {\it de novo} TE identification
-tools, identifying TEs using only the genome sequence information.  A
-number of such tools are discussed in the literature.  RECON uses
-WU-BLAST and PILER using LASTZ to compute self-alignments
-\cite{Bao:2002,Edgar:2005p2365,Lopez:2003td,Harris:2007uf}.  RECON
-show good sensitivity but is computationally intensive and infeasible
-for use on whole genomes (requiring 60 hours for 18Mb rice genome in a
-2013 study), while PILER achieves a good runtime with very low
-sensitivity .  ReAS and RepeatScout \cite{Price:2005p1247} are based
-on $k$-mer searches, with the earlier showing less sensitivity than
-RECON \cite{Li:2005he, Price:2005p1247,Jiang:2013jt}.  RepeatGluer
-\cite{Pevzner:2004p3157,Zhi:2006p3199} is based on a variation of
-DeBrujin graphs, which allows for a decomposition of TE families into
-domains, but is very, very computationally expensive.  In Saha {\it et
-  al.} the authors perform an extensive comparison of the tools, and
-conclude that RepeatScout is the best tool overall for assembled
-genomes, while ReAS the best when dealing with unassembled sequence
-fragments \cite{Saha:2008dm}.
-
-\subsection*{Elementary Repeats}
-Zheng and Lonardi approached the {\it de novo} identification problem using {\it elementary repeats}
-\cite{Zheng:2005bl}.
-Similar to the RepeatGluer domains, elementary repeats are
-decompositions of TEs into basic building blocks.  Identification of
-these building blocks are sufficient for the purpose
-of masking, and can be assembled into Transposable Elements for those
-interested in TEs themselves.
-
-While it is notoriously difficult to mathematically model transposable
-elements \cite{Bao:2002}, elementary repeats are more conducive to a
-formal description.  For a given genome, a nucleotide sequence $r$ is
-an elementary repeat if: (1) It is of at least length $l$ (the length
-requirement); (2) there are at least $f$ copies of $r$ appear (the
-frequency requirement) (3) there is no proper substring of $r$ of
-length $\leq l$ that appears in the genome independently of $r$ (the
-minimality requirement); (4) $r$ is a maximal string w.r.t (1-3) (the
-maximality requirement).  Having proposed this definition, Zheng and
-Lonardi developed an identification algorithm that had a runtime
-quadratic in the query sequence size \cite{Zheng:2005bl}. This was
-refined to linear time by He and also by Huo {\it et al.}
-\cite{He:2006gpa,Huo:2009hoa} based on variations of suffix tree
-approaches, but these appraoches are limited in their ability to handle sequence
-variation.  As we are looking a genome size inputs, and TEs
-transposable elements naturally suffer from copy mistakes and
-accumulate instance-specific base substitutions over time, this is a
-significant limitation.
-
-\subsection*{RAIDER}
-\label{RAIDERSec}
-It was with the objective of creating a linear time identification
-algorithm that could handle variation through use of
-PatternHunter-like spaced seeds that we developed the prototype RIADER
-\cite{Li:2004wl}.  A rough implementation was first presented in
-Figueroa {\it et al.}, with more details in the Figueroa masters
-thesis \cite{Figueroa:2014uk,Figueroa:2013cz}.  RAIDER was built along
-an alternate, but equivalent definition of elementary repeats based on
-$l$-mers (sequences of length $l$).  Specifically, it was observed
-that the minimality condition could be rewritten as: There is no
-$l$-mer contained within elementary repeat $r$ that appears in the
-genome more times than $r$.  From there we make four core observations
-that form the basis for the RAIDER algorithm: (1) An $l$-mer cannot
-belong to two different elementary repeats; (2) Any $l$-mer in the
-genome that occurs $f$ or more times is either an elementary repeat or
-belongs to one; (3) any two $l$-mers belonging to an elementary repeat
-must appear the same number of times in the genome; (4) If two
-sequences in the genome that are {\it maximally identical} (that is,
-cannot be extended in either direction and still be identical), these
-sequences cannot belong to a larger elementary repeat.  (By ``belong''
-we mean ``is a substring of'' -- a definition we will be generalizing
-shortly.)  For discussion and proof, see the Figueroa Thesis
-\cite{Figueroa:2013cz}.
-
-Based on these observations, we discover we can find all elementary
-repeats in a single scan of the genome.  Specifically, as we scan from
-left to right, we track $l$-mer occurrences and identify multiple
-copies of the same $l$-mer.  When we find the same sequence of $l$-mers
-occurring multiple times in a row, we can mark it as a tentative family,
-then break it down later if we discover violations of the minimality
-condition.  The algorithm is summarized in Figueroa {\it et al.}, and
-discussed in detail in the Figueroa Thesis \cite{Figueroa:2014uk,Figueroa:2013cz}.
-
-Results of the preliminary implementation were promising. On human
-chromosome 22 we saw a $12\times$ speedup over RepeatScout to RAIDER
-(2344 seconds to 192 seconds), while coverage of the RepBase
-\cite{Jurka:2005bl} ancestral sequence improved (77\% to 84\%), while
-on mouse chromosome 19 we saw the same speedup with a significant drop
-in coverage (53\% to 30\%).  On the full human genome RAIDER ran in
-$6.3$ hours, while RepeatScout was unable to complete it run.  For
-details, see Figueroa {\it et al.} \cite{Figueroa:2014uk}.
-
-\subsection*{Spaced Seeds}
-
-PatternHunter, a very successful augmentation to BLAST
-\cite{Li:2004wl,Altschul:1997p843}, is based on the notion of {\it
-  spaced seeds}: improving the sensitivity of string matching based
-algorthms by allowing wild-cards in the match.  That is, instead of
-requiring two strings matching in 12 consecutive characters, we might
-instead require two six-character exact matches seperated by one base
-which may or may not match (represented by the {\it seed pattern}
-11111101111111), or perhaps three consecutive four-character exact
-matches sperated by two bases each (1111001111001111). It has been
-demonstrated that certain seed patterns can indunce significant
-imporvements in BLAST sensitivity with out time penalty, though what
-makes a good pattern is not well understood.
-
-RAIDER was designed with the intent of employing the spaced seed
-strategy, but this was only implemented heuristically for the Figuera
-{\it et al.} paper \cite{Figueroa:2013cz} -- serving primarily as a
-proof-of-concept.  Since its release we have develoed a formal model
-of elementary repeats that incorporated spaced seeds, and from that
-developed phRAIDER (PatternHunter-based RAIDER).  phRAIDER is a fast
-tool for the identification and making of transposible elements in
-both assembled and unassembled genomes aht outpreforms RptScout and
-other establihsed tools. Code is free available under the Gnu GPL
-lisence (v. 3) and  may be obtained [NEED WEB ADDRESS].
-
-In the following we will present a new theoretical model for {\it
-  seeded elementary repeats}, followed by an algorithm for identifying
-them.  In the results section we will show the result of using that
-algorithm for masking transp
-
-In our Methods section
-we present the new model that allows us to extend RAIDER to correctly
-use spaced seeds, and briefly outline the algorithm (with more details
-provided in the supplementary materials), with a analysis of phRAIDER
-performance in our Results.
-
-\section*{Model}
-
-Our goal is to redefine the concept of transposbile elements to
-accomodate a spaced seed strategy.  In the next section we will describe our
-identification algorithm, and then quantified its success in 
-masking transposable elements. But we will start here with a brief
-outline of our theoretical model (with a more detailed description in
-the appendix).
-
-We first need to redfine out terminimology regarding elementary
-repeats.  Under the Z\&L definition, all instances of one elementary
-repeat have the exact same sequence, and hence can be descrbied by a
-single string.  As we are allowing for variation, we will describe out
-instance set with a {\it sequence descriptor} consisting of bases
-letters and the wild-card character * (e.g. AAC*GG would describe a
-set of sequences with starting with an AAC, ending with a GG, and
-having any character in between).  Given a binary string $s$
-representing a spaced seed, we say a sequence descriptor $r$ is
-{\it consistent} with $s$ if we can align $s$ to a substring of $r$ such
-that every $*$ in $r$ aligns with a $0$ character in $s$. (Hence
-$ACG**T*A$ is consistent with the seed $11001$, but not $11011$.)
-
-Given a sequence descriptor $r$, we can {\it decompose} $r$ with
-respect to $s$ by taking every length $|s|$ substring of $r$ that is consistant
-with $s$, replacing all letters of $r$ that match to a 0 in $s$ with a
-$*$, and creating a set from the results.  (Hence for $s = 11011$ and
-$r=AA*CCGTT$, the decomposition would be $\{AA*CC, CC*TT\}$.)  We say
-$s$ {\it covers} $r$ is every base in $r$ is contained in at least on
-string in the decompositio.  (Hence the previous $s$ does not cover $AAA*CC$, as
-the first base in not in any of the strings of the decomposition.)
-
-We can now modif the previous definition of elementary repeats as
-follows:
-\begin{definition} Given a genomic sequence $G$, an integer $f$, and a
-  spaced seed $s$, a sequence descriptor $r$ describes an elementary
-  repeat if it meets the four (moified) requirements of an elementary repeat:
-\begin{itemize}
-\item \underline{Structure requriement}: $s$ covers $r$.
-\item \underline{Frequency requirement}: There are at least $f$
-  substrings of $G$ that match $r$.
-\item \underline{Minimality requirement:} For every string $t$ in the
-  decomposition of $r$ w.r.t $s$, the number of occurences of $t$ in
-  $G$ is equal to the number of occurneces of $r$ in the genome.
-\item \underline{Maximality requirement:} There is no sequence
-  descriptor $r'$ of $r$ that contains $r$ as a proper substring and
-  satisfies conditions 1-1.
-\end{itemize}
-\end{definition}
-
-\begin{theorem} When the seed $s$ has no 0 characters, this
-    definition of elementary repeats is equivilent to the Z\&L
-    definition.
-\end{theorem}
-
-\subsection*{phRAIDER Algorithm}
-
-\red{
-TO DO: Give a pseudocode description of the algoirthm and some minimal
-discussion.
-}
-
-\subsection*{Analysis}
-
-\red{
-TO DO: Describe pra analysis and quality metrics.
-}
-
-\section*{Results}
-
-\red{
-Points to be made:
-\begin{enumerate}
-\item Show: RAIDER2 has better quality masking results then RS.
-\item Show: RAIDER2 is better than RAIDER.
-\item Show: Seeds matter.
-\item Show: Runtime / memory usage
-\end{enumerate}
-
-Possible tables / plots:
-\begin{itemize}
-\item Table showing results on different chromosomes (pra, runtime, memory).
-\item Plot of runtime against genome size (simulated data to control size)?
-\item Plot of memory usage (simlated data to control size)?
-\item We might want to compare the ability to detect transposable
-  elements againsts other types of repeats.  (Should be easy to do
-  using simulator.)
-\end{itemize}
-}
-
-\section*{Conclusions}
-
-\red {
-Hopefully will beconcluding that RAIDER is a better tool than
-RepeatScout.  Posible discussion of tool features.
-}
-
-\section*{Appendix}
-In order to allow the use of a spaced seed strategy, we will need to
-define a new model of elementary repeats that will accommodate it.  In
-the following we will first outline our theoretical framework
-(described in detail in the Supplementary Materials), and then review
-the phRAIDER algorithm.
-
-\subsection*{Seeded Elementary Repeats}
-In order to present our new model of elementary repeats, we first need
-to define a few terms:
-\begin{enumerate}
-  \item A spaced seed is a binary string that starts and ends with 1
-    symbols (as defined in Li {\it et al.} \cite{Li:2004wl}).
-  \item A {\it sequence descriptor} is a DNA string that allows *
-    symbols (indicating that a position where base content does not
-    matter).  A string matches a sequence descriptor of equal length if
-    it has the same base in all non-* positions.
-  \item We say the frequency of $r$ in a genomic sequence $G$ (denoted
-    $\nu_G(r)$) is the number of subsequences in $G$ that match $r$.
-  \item A spaced seed $s$ {\it hits} a substring of sequence
-    descriptor $r$ if that substring is the same length of $s$ and,
-    when aligned, every $*$ in the substring matches a 0 in the
-    string.  (Example: if $s=11011$ and $r=AAA*TTTTT$, then $s$
-    hits both $AA*CC$ and $TTTTT$, but not $A*TTT$.)
-  \item The {\it decomposition} of $r$ (with respect to seed $s$) is
-    the set of all substrings of $r$ that are hit by $s$.  For the
-    {\it generalized decomposition}, we take each member of the
-    decomposition and replace by $*$ any base that aligned to a 0 in
-    the seed.  (Example: if $s=11011$ and $r=AAATTT$, then $AAATT$ is
-    in the decomposition, so $AA*TT$ is in the generalized
-    decomposition.)
-  \item We say a string $s$ {\it covers} a sequence descriptor $r$ if
-  every base in $r$ is contained within at least one substring from
-  the decomposition of $r$.
-\end{enumerate}
-
-\begin{definition}
-  \label{REDDef}
-  Given a fixed genomic sequence $G$, a integer $f$, and spaced seed $s$, a sequence
-  descriptor $r$ is a {\it repeat element descriptor} if:
-  \begin{enumerate}
-  \item $s$ covers $r$.  (Minimum length requirement.)
-  \item $\nu_G(r) \geq f$.  (Frequency requirement.)
-  \item For every string $w$ in the generalized decomposition of $r$
-    (w.r.t. $s$), $\nu_G(w) = \nu_G(r)$.  (Minimality requirement.)
-  \item There is no sequence descriptor $r'$ such that (1) $r$ is
-    a substring of $r'$, (2) $\nu_G(r') = \nu_G(r)$, and (3) $r'$
-    satisfies conditions 1-3.  (Maximality requirement.)
-  \end{enumerate}
-\end{definition}
-
-We will refer to the set of sequences in the genome that match the
-repeat family descriptor to be the repeat element family of the
-descriptor.
-
-\begin{theorem}
-Definition~\ref{REDDef} is equivalent to the Z\&L
-definition of elementary repeats when $s$ contains no 0 symbols.
-\end{theorem}
-In the Supplementary materials we prove our definition equivalent to
-the Figueroa definition (Definition 5 from the thesis) -- which is
-proved to be equivalent to the Z\&L definition in that work \cite{Figueroa:2014uk}.
-
-It is instructive to look at a specific example.  Consider the toy
-seed $11011$, let $f=2$, and assume the genome contains a unique
-instance of $g_1 = AAAAA{\mathbf C}TTTTT$ and a unique instance of
-$g_2 = AAAAA{\mathbf G}TTTTT$.  Consider the sequence descriptor $r_1
-= AAAAA*TTTTT$.  A simple inspection shows us that conditions (1)-(3)
-of Definition~\ref{REDDef} hold, so if we assume the surrounding
-sequence is such that (4) holds as well, this is an elementary repeat
-descriptor for this family of sequences.  (Note $AA*AA*TT*TT$ also qualifies -- the
-elementary repeat descriptor is not necessarily unique.)  But if we add a second
-instance of $g_3 = AAAAAGTTTTT$, something happens that was not
-possible in the unseeded version.  While $r_1$ continues to meet the definition, the $g_2/g_3$
-subsequence $AAAAGTTTT$ also induces an elementary repeat family
-$r_2 = AAAA*TTTT$. (Note that we cannot extend this by, for example,
-one base to the left, as the $AAAAA$ substring would cause a violation
-of condition (3)).  So $r_2$ matches strings in $G$ that are
-substrings of those that $r_1$ matches -- a condition not possible
-with the seedless elementary repeats.  This also then violates one of
-our core observation on which RAIDER is broken: we now have an $l$-mer
-($AAAA*$) that belongs to multiple descriptors.
-
-A second complicating factor is that our repeat element descriptor is
-not a sequence of consecutive $l$-mers. In the seedless version, an
-elementary repeat of length $n$ can be viewed as a sequence of $n-l+1$
-$l$-mers, each overlapping the last by exactly $l-1$ bases. In this
-definition we use members of the generalize decomposition, and find
-they are not necessarily offset by just one base: given the $g_1$ and
-$g_2$ example above, we see that the $r_1 = AAAAA*TTTTT$ descriptor is
-composed of $AA*AA$, $AA*TT$, and $TT*TT$ -- each offset from the last
-by 3.  As RAIDER assumed that it need a look-back of only one, this
-too breaks the algorithm.
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                                          %%
-%% Backmatter begins here                   %%
-%%                                          %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\begin{backmatter}
-
-\section*{Competing interests}
-  The authors declare that they have no competing interests.
-
-\section*{Author's contributions}
-    Text for this section \ldots
-
-\section*{Acknowledgements}
-  Text for this section \ldots
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                  The Bibliography                       %%
-%%                                                         %%
-%%  Bmc_mathpys.bst  will be used to                       %%
-%%  create a .BBL file for submission.                     %%
-%%  After submission of the .TEX file,                     %%
-%%  you will be prompted to submit your .BBL file.         %%
-%%                                                         %%
-%%                                                         %%
-%%  Note that the displayed Bibliography will not          %%
-%%  necessarily be rendered by Latex exactly as specified  %%
-%%  in the online Instructions for Authors.                %%
-%%                                                         %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-% if your bibliography is in bibtex format, use those commands:
 \bibliographystyle{bmc-mathphys} % Style BST file (bmc-mathphys, vancouver, spbasic).
 \bibliography{bmc}      % Bibliography file (usually '*.bib' )
-% for author-year bibliography (bmc-mathphys or spbasic)
-% a) write to bib file (bmc-mathphys only)
-% @settings{label, options="nameyear"}
-% b) uncomment next line
-%\nocite{label}
 
-% or include bibliography directly:
-% \begin{thebibliography}
-% \bibitem{b1}
-% \end{thebibliography}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                               %%
-%% Figures                       %%
-%%                               %%
-%% NB: this is for captions and  %%
-%% Titles. All graphics must be  %%
-%% submitted separately and NOT  %%
-%% included in the Tex document  %%
-%%                               %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%%
-%% Do not use \listoffigures as most will included as separate files
-
-%% \section*{Figures}
-%%   \begin{figure}[h!]
-%%   \caption{\csentence{Sample figure title.}
-%%       A short description of the figure content
-%%       should go here.}
-%%       \end{figure}
-
-%% \begin{figure}[h!]
-%%   \caption{\csentence{Sample figure title.}
-%%       Figure legend text.}
-%%       \end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                               %%
-%% Tables                        %%
-%%                               %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% Use of \listoftables is discouraged.
-%%
-%% \section*{Tables}
-%% \begin{table}[h!]
-%% \caption{Sample table title. This is where the description of the table should go.}
-%%       \begin{tabular}{cccc}
-%%         \hline
-%%            & B1  &B2   & B3\\ \hline
-%%         A1 & 0.1 & 0.2 & 0.3\\
-%%         A2 & ... & ..  & .\\
-%%         A3 & ..  & .   & .\\ \hline
-%%       \end{tabular}
-%% \end{table}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-%%                               %%
-%% Additional Files              %%
-%%                               %%
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-%% \section*{Additional Files}
-%%   \subsection*{Additional file 1 --- Sample additional file title}
-%%     Additional file descriptions text (including details of how to
-%%     view the file, if it is in a non-standard format or the file extension).  This might
-%%     refer to a multi-page table or a figure.
-
-%%   \subsection*{Additional file 2 --- Sample additional file title}
-%%     Additional file descriptions text.
-
-\end{backmatter}
 \end{document}
-
-%%  LocalWords:  de addressref aff neteref Schaeffer noteref Xiaolin
-%%  LocalWords:  Liu corref JE Karro Initio phRAIDER genomic TEs Ns
-%%  LocalWords:  JIANG RepeatMasker nHMMER BLASTing homology ALU ReAS
-%%  LocalWords:  LINEs Jiang PILER RepeatScout mer RepeatGluer Saha
-%%  LocalWords:  infeasible DeBrujin et al Zheng Lonardi pre Huo mers
-%%  LocalWords:  PatternHunter decompositions subsequence minimality
-%%  LocalWords:  maximality RIADER ers substring AACACA AAGACA TACACA
-%%  LocalWords:  TTAACTCA TTTTT TTT substrings AAATTT AAATT TT AAAAA
-%%  LocalWords:  FIGUREOA AAAAAGTTTTT AAAAGTTTT AAAA TTTT AAAAT novo
-%%  LocalWords:  Acknowledgements transposable tranposable runtime
-%%  LocalWords:  unassembled RepBase AAGATA

### 2015-09-18 manuscript/JCB/manuscript.tex
Updating manuscript.txt

@@
 \usepackage{authblk}
-\usepackage{algorithmic}
+\usepackage{algorithm}
+\usepackage[noend]{algpseudocode}
+\usepackage{algorithmicx}
+\usepackage{color}
 \def\includegraphic{}
@@
 \pagestyle{myheadings}
+
+\newcommand{\red}[1]{{\color{red}#1}}
+\newcommand{\blue}[1]{{\color{blue}#1}}
+\newcommand{\cyan}[1]{{\color{cyan}#1}}
+\newcommand{\magenta}[1]{{\color{magenta}#1}}
 
@@
 \renewcommand\Affilfont{\itshape\small}
+
+\newcommand{\comment}[1]{\State \red{\# #1}}
 
@@
 
+\begin{algorithm}
 \begin{algorithmic}
-\If {$i\geq maxval$}
-    \State $i\gets 0$
+  \Function{phRAIDER}{Genome $G$, Seed $s$, MinFrequence $f$}
+    \State Queue$<$Family$>$ $Q$;
+
+    \State \Comment{Iterate over all $l$-mer positions in $G$}
+    \For {$i \leftarrow 1 \to |G|-|s|$}
+
+       \comment{Remove and families whose last $l$-mer did not occur in the last $i$}
+       \comment{bases (hence must be closd out).}
+       \While{$i - $location($Q$, finish, last, end)$ > |s|$}
+         \State $Q.$dequeue()
+       \EndWhile
+       \State
+       \comment{Get the $l$-mer occuring at $G_i$ and remove letters corresponding to $0$s in $s4$.}
+       \State $v \leftarrow \mbox{seeded}(G[i],x)$ 
+       \State
+       \comment{Add to list of $v$'s positions.}
+       \State $H[v].push(i)$
+       \State
+       \If {$|H|(v)| == 2$}
+          \comment{$v$ either is the beginning of a family, or should be combined}
+          \comment{with a previous $l$-mer to form a family.}
+          \State $F = \mbox{arg\_find}_{F \in Q}\{$loc($v$, start, first) - loc($f$, start, first, begin) $==i - $loc($f$, start last end)$\}$
+          \If {$F$}
+             $Q.$requeue($F$)
+          \EndIf
+          \State last($Q$).addLmer($v$)   
+       \ElsIf {$|H[v]| > 2$}
+          \comment{Is $v$ the proper continuation of a family, or does it}
+          \comment{need to be split off?}
+          \State $F \leftarrow v.$family()
+          \If {$f \not\in Q$}
+              \comment{$v$ belongs to a family whose last instance is too far away.}
+              \If {$v \neq F.first()$}
+                  \comment{$v$ is not the first $l$-mer of its family
+                  $F' \leftarrow F.$split($F.$v$)
               \Else
-    \If {$i+k\leq maxval$}
-        \State $i\gets i+k$
+                  $F' \leftarrow F.$split($F.$lastSeen())
+              \EndIf
+              \State $Q.$enqueue($F'$)
+          \Else
+              \If {$v == F.$first()}
+                 \comment{$v$ is the first $l$-mer of $F$, so $F$ is moved to the back of $Q$}
+                 $Q$.requeue($F$)
+              \ElsIf {$F.$lastSeen()$.$next$==v$}
+                 ???
+                 $F.$setLastSeen($v$)
               \EndIf
            \EndIf                             
+       \EndIf
+       \EndFor
+       \State tieLooseEnds()
+  \EndFunction
 \end{algorithmic}
+\end{algorithm}
 

### 2015-09-18 manuscript/JCB/manuscript.tex
Updating manuscript -- copied over text, worked on algorithm format.

--- 2015-09-17 manuscript/JCB/manuscript.tex
@@
-\documentclass[10pt]{article}
+\documentclass[11pt]{article}
 \usepackage{longtable}
@@
 \renewcommand\Affilfont{\itshape\small}
+
+\newcommand{\comment}[1]{\State \red{\# #1}}
+
+\newtheorem{definition}{Definition}
+\newtheorem{observation}{Observation}
+\newtheorem{lemma}{Lemma}
+\newtheorem{theorem}{Theorem}
 
@@
 
+Within mammalian species we can largely rely on homology relationships
+to port libraries across species.  This does not hold
+so well in plants: in many cases TE composition of a given plant
+organism is species-specific.  For example, a rice-based TE library
+will only identify 25\% of the TEs in the maize genome
+\cite{Jiang:2013jt}.
+
+To solve this problem we turn to {\it de novo} TE identification
+tools, identifying TEs using only the genome sequence information.  A
+number of such tools are discussed in the literature.  RECON uses
+WU-BLAST and PILER using LASTZ to compute self-alignments
+\cite{Bao:2002,Edgar:2005p2365,Lopez:2003td,Harris:2007uf}.  RECON
+show good sensitivity but is computationally intensive and infeasible
+for use on whole genomes (requiring 60 hours for 18Mb rice genome in a
+2013 study), while PILER achieves a good runtime with very low
+sensitivity .  ReAS and RepeatScout \cite{Price:2005p1247} are based
+on $k$-mer searches, with the earlier showing less sensitivity than
+RECON \cite{Li:2005he, Price:2005p1247,Jiang:2013jt}.  RepeatGluer
+\cite{Pevzner:2004p3157,Zhi:2006p3199} is based on a variation of
+DeBrujin graphs, which allows for a decomposition of TE families into
+domains, but is very, very computationally expensive.  In Saha {\it et
+  al.} the authors perform an extensive comparison of the tools, and
+conclude that RepeatScout is the best tool overall for assembled
+genomes, while ReAS the best when dealing with unassembled sequence
+fragments \cite{Saha:2008dm}.
+
+\subsection*{Elementary Repeats}
+Zheng and Lonardi approached the {\it de novo} identification problem using {\it elementary repeats}
+\cite{Zheng:2005bl}.
+Similar to the RepeatGluer domains, elementary repeats are
+decompositions of TEs into basic building blocks.  Identification of
+these building blocks are sufficient for the purpose
+of masking, and can be assembled into Transposable Elements for those
+interested in TEs themselves.
+
+While it is notoriously difficult to mathematically model transposable
+elements \cite{Bao:2002}, elementary repeats are more conducive to a
+formal description.  For a given genome, a nucleotide sequence $r$ is
+an elementary repeat if: (1) It is of at least length $l$ (the length
+requirement); (2) there are at least $f$ copies of $r$ appear (the
+frequency requirement) (3) there is no proper substring of $r$ of
+length $\leq l$ that appears in the genome independently of $r$ (the
+minimality requirement); (4) $r$ is a maximal string w.r.t (1-3) (the
+maximality requirement).  Having proposed this definition, Zheng and
+Lonardi developed an identification algorithm that had a runtime
+quadratic in the query sequence size \cite{Zheng:2005bl}. This was
+refined to linear time by He and also by Huo {\it et al.}
+\cite{He:2006gpa,Huo:2009hoa} based on variations of suffix tree
+approaches, but these appraoches are limited in their ability to handle sequence
+variation.  As we are looking a genome size inputs, and TEs
+transposable elements naturally suffer from copy mistakes and
+accumulate instance-specific base substitutions over time, this is a
+significant limitation.
+
+\subsection*{RAIDER}
+\label{RAIDERSec}
+It was with the objective of creating a linear time identification
+algorithm that could handle variation through use of
+PatternHunter-like spaced seeds that we developed the prototype RIADER
+\cite{Li:2004wl}.  A rough implementation was first presented in
+Figueroa {\it et al.}, with more details in the Figueroa masters
+thesis \cite{Figueroa:2014uk,Figueroa:2013cz}.  RAIDER was built along
+an alternate, but equivalent definition of elementary repeats based on
+$l$-mers (sequences of length $l$).  Specifically, it was observed
+that the minimality condition could be rewritten as: There is no
+$l$-mer contained within elementary repeat $r$ that appears in the
+genome more times than $r$.  From there we make four core observations
+that form the basis for the RAIDER algorithm: (1) An $l$-mer cannot
+belong to two different elementary repeats; (2) Any $l$-mer in the
+genome that occurs $f$ or more times is either an elementary repeat or
+belongs to one; (3) any two $l$-mers belonging to an elementary repeat
+must appear the same number of times in the genome; (4) If two
+sequences in the genome that are {\it maximally identical} (that is,
+cannot be extended in either direction and still be identical), these
+sequences cannot belong to a larger elementary repeat.  (By ``belong''
+we mean ``is a substring of'' -- a definition we will be generalizing
+shortly.)  For discussion and proof, see the Figueroa Thesis
+\cite{Figueroa:2013cz}.
+
+Based on these observations, we discover we can find all elementary
+repeats in a single scan of the genome.  Specifically, as we scan from
+left to right, we track $l$-mer occurrences and identify multiple
+copies of the same $l$-mer.  When we find the same sequence of $l$-mers
+occurring multiple times in a row, we can mark it as a tentative family,
+then break it down later if we discover violations of the minimality
+condition.  The algorithm is summarized in Figueroa {\it et al.}, and
+discussed in detail in the Figueroa Thesis \cite{Figueroa:2014uk,Figueroa:2013cz}.
+
+Results of the preliminary implementation were promising. On human
+chromosome 22 we saw a $12\times$ speedup over RepeatScout to RAIDER
+(2344 seconds to 192 seconds), while coverage of the RepBase
+\cite{Jurka:2005bl} ancestral sequence improved (77\% to 84\%), while
+on mouse chromosome 19 we saw the same speedup with a significant drop
+in coverage (53\% to 30\%).  On the full human genome RAIDER ran in
+$6.3$ hours, while RepeatScout was unable to complete it run.  For
+details, see Figueroa {\it et al.} \cite{Figueroa:2014uk}.
+
+\subsection*{Spaced Seeds}
+
+PatternHunter, a very successful augmentation to BLAST
+\cite{Li:2004wl,Altschul:1997p843}, is based on the notion of {\it
+  spaced seeds}: improving the sensitivity of string matching based
+algorthms by allowing wild-cards in the match.  That is, instead of
+requiring two strings matching in 12 consecutive characters, we might
+instead require two six-character exact matches seperated by one base
+which may or may not match (represented by the {\it seed pattern}
+11111101111111), or perhaps three consecutive four-character exact
+matches sperated by two bases each (1111001111001111). It has been
+demonstrated that certain seed patterns can indunce significant
+imporvements in BLAST sensitivity with out time penalty, though what
+makes a good pattern is not well understood.
+
+RAIDER was designed with the intent of employing the spaced seed
+strategy, but this was only implemented heuristically for the Figuera
+{\it et al.} paper \cite{Figueroa:2013cz} -- serving primarily as a
+proof-of-concept.  Since its release we have develoed a formal model
+of elementary repeats that incorporated spaced seeds, and from that
+developed phRAIDER (PatternHunter-based RAIDER).  phRAIDER is a fast
+tool for the identification and making of transposible elements in
+both assembled and unassembled genomes aht outpreforms RptScout and
+other establihsed tools. Code is free available under the Gnu GPL
+lisence (v. 3) and  may be obtained [NEED WEB ADDRESS].
+
+In the following we will present a new theoretical model for {\it
+  seeded elementary repeats}, followed by an algorithm for identifying
+them.  In the results section we will show the result of using that
+algorithm for masking transp
+
+In our Methods section
+we present the new model that allows us to extend RAIDER to correctly
+use spaced seeds, and briefly outline the algorithm (with more details
+provided in the supplementary materials), with a analysis of phRAIDER
+performance in our Results.
+
+\section*{Model}
+
+Our goal is to redefine the concept of transposbile elements to
+accomodate a spaced seed strategy.  In the next section we will describe our
+identification algorithm, and then quantified its success in 
+masking transposable elements. But we will start here with a brief
+outline of our theoretical model (with a more detailed description in
+the appendix).
+
+We first need to redfine out terminimology regarding elementary
+repeats.  Under the Z\&L definition, all instances of one elementary
+repeat have the exact same sequence, and hence can be descrbied by a
+single string.  As we are allowing for variation, we will describe out
+instance set with a {\it sequence descriptor} consisting of bases
+letters and the wild-card character * (e.g. AAC*GG would describe a
+set of sequences with starting with an AAC, ending with a GG, and
+having any character in between).  Given a binary string $s$
+representing a spaced seed, we say a sequence descriptor $r$ is
+{\it consistent} with $s$ if we can align $s$ to a substring of $r$ such
+that every $*$ in $r$ aligns with a $0$ character in $s$. (Hence
+$ACG**T*A$ is consistent with the seed $11001$, but not $11011$.)
+
+Given a sequence descriptor $r$, we can {\it decompose} $r$ with
+respect to $s$ by taking every length $|s|$ substring of $r$ that is consistant
+with $s$, replacing all letters of $r$ that match to a 0 in $s$ with a
+$*$, and creating a set from the results.  (Hence for $s = 11011$ and
+$r=AA*CCGTT$, the decomposition would be $\{AA*CC, CC*TT\}$.)  We say
+$s$ {\it covers} $r$ is every base in $r$ is contained in at least on
+string in the decompositio.  (Hence the previous $s$ does not cover $AAA*CC$, as
+the first base in not in any of the strings of the decomposition.)
+
+We can now modif the previous definition of elementary repeats as
+follows:
+\begin{definition} Given a genomic sequence $G$, an integer $f$, and a
+  spaced seed $s$, a sequence descriptor $r$ describes an elementary
+  repeat if it meets the four (moified) requirements of an elementary repeat:
+\begin{itemize}
+\item \underline{Structure requriement}: $s$ covers $r$.
+\item \underline{Frequency requirement}: There are at least $f$
+  substrings of $G$ that match $r$.
+\item \underline{Minimality requirement:} For every string $t$ in the
+  decomposition of $r$ w.r.t $s$, the number of occurences of $t$ in
+  $G$ is equal to the number of occurneces of $r$ in the genome.
+\item \underline{Maximality requirement:} There is no sequence
+  descriptor $r'$ of $r$ that contains $r$ as a proper substring and
+  satisfies conditions 1-1.
+\end{itemize}
+\end{definition}
+
+\begin{theorem} When the seed $s$ has no 0 characters, this
+    definition of elementary repeats is equivilent to the Z\&L
+    definition.
+\end{theorem}
+
+\subsection*{phRAIDER Algorithm}
+
+\begin{algorithm}
+\caption{This is the algorthm caption.}
 \begin{algorithmic}
-\If {$i\geq maxval$}
-    \State $i\gets 0$
+  \Function{sameOffset}{$l$-mer $v_1$, $l$-mer $v_2$}
+    \comment{Is the distance between the first two occurences of $v$ equal to the}
+    \comment{the distance betwen the most recent two occurences?}
+    \State \Return $H[v_2][0] - H[v_1][0]  == H[v_2][-1] - H[v_1][-1]$
+  \EndFunction
+  \State
+  \Function{phRAIDER}{Genome $G$, Seed $s$, MinFrequence $f$}
+    \comment{We are going to maintain a $Q$ of up to $|s|$ families, representing the families}
+    \comment{of the last $|L|$ $l$-mers seen, sorted by the position of those last $l$-mers.}
+    \State Queue$<$Family$>$ $Q$;
+    \State
+    \comment{Iterate over all $l$-mer positions in $G$}
+    \For {$i \leftarrow 1 \to |G|-|s|$}
+
+       \comment{Remove any families whose last $l$-mer did not occur in the last $i$ bases.}
+       \While{$i - $location($Q$, finish, last, end)$ > |s|$}
+         \State $Q.$dequeue()
+       \EndWhile
+       \State
+       \comment{Get the $l$-mer occuring at $G_i$ and remove letters corresponding to $0$s in $s$.}
+       \State $v \leftarrow \mbox{seeded}(G[i],x)$ 
+       \State
+       \comment{Add to list of $v$'s positions.}
+       \State $H[v].push(i)$
+       \State
+       \If {$|H|(v)| == 2$}
+          \comment{$v$ either is the beginning of a family, or should be combined with a}
+          \comment{previous $l$-mer to form a family.}
+          \State $F = \mbox{arg\_find}_{F \in Q}\{\mbox{sameOffset}(F.\mbox{start}(), v)\}$
+          \If {$F$}
+             \State $Q.$requeue($F$)
+          \EndIf
+          \State last($Q$).addLmer($v$)   
+       \ElsIf {$|H[v]| > 2$}
+          \comment{Is $v$ the proper continuation of a family, or does it need to be split off?}
+          \State $F \leftarrow v.$family()
+          \If {$H[F.\mbox{lastSeen}(i)][-1] < H[Q.\mbox{front}().\mbox{lastSeen}()]$}
+              \comment{$v$ belongs to a family whose last instance is too far back.}
+              \If {$v \neq F.$first$()$}
+                  \comment{$v$ is not the first $l$-mer of its family}
+                  \State $F' \leftarrow F.$split($F.v$)
               \Else
-    \If {$i+k\leq maxval$}
-        \State $i\gets i+k$
+                  \State $F' \leftarrow F.$split($F.$lastSeen())
+              \EndIf
+              \State $Q.$enqueue($F'$)
+          \Else
+              \If {$v == F.$first()}
+                 \comment{$v$ is the first $l$-mer of $F$, so $F$ is moved to the back of $Q$}
+                 \State $Q$.requeue($F$)
+              \ElsIf {sameOffSet($F.$first(), $v$)}
+                 \State $F.$setLastSeen($v$)
               \EndIf
            \EndIf                             
+       \EndIf
+       \EndFor
+       \State
+       \comment{Final break up of families as needed; filter families with frequency $\leq f$.}
+       \State tieLooseEnds()
+  \EndFunction
 \end{algorithmic}
+\end{algorithm}
 

### 2015-09-18 manuscript/JCB/old.manuscript.tex
Updating manuscript

--- 2015-09-14 manuscript/JCB/bmc_article.tex
@@
 
-\newtheorem{definition}{Definition}
-\newtheorem{observation}{Observation}
-\newtheorem{lemma}{Lemma}
-\newtheorem{theorem}{Theorem}
-
 %%%%%%%%%%%%%%%%%%%%%%%%% start of article main body
 \section*{Background}
-Transposable Elements (TEs) are genomic sequences that have at some
-point had the capacity to insert copies of themselves into other
-genomic locations, resulting in homologous families of sequences
-spread across the genome.  Present in almost every higher
-order genome (covering as much as 45\% of the human genome and 90\% of
-the maize genome \cite{Venter:2001p92,SanMiguel:1996wa}), TEs have
-proved an important source of data in numerous studies of genomic structure (e.g.
+Transposable Elements (TEs) are genomic sequences thathad the capacity
+to insert copies of themselves into other genomic locations, resulting
+in homologous families of sequences spread across the genome.  Present
+in almost every higher order genome (covering as much as 45\% of the
+human genome and 90\% of the maize genome
+\cite{Venter:2001p92,SanMiguel:1996wa}), TEs have proved an important
+source of data in numerous studies of genomic structure (e.g.
 \cite{Arndt:2005p279,Karro:2008p123,Mugal:2009p581,Hardison:2003p97}).
-But given their prevalence, it is important for those studying other aspect
-the genome to have TEs masked out -- their bases replaced by N to
-allow for easy identification and filtering.  Failure to filter can
-reek havoc with genomic analysis tools.  For example,
-TEs can trigger huge numbers of false positives automated gene finding
-tool \cite{Jiang:2013jt}, as well as inflate tool runtime.
+But given their prevalence, it is important for those studying other
+aspect the genome to have TEs masked out -- their bases replaced by N
+to allow for easy identification and filtering.  Failure to filter can
+reek havoc with genomic analysis tools.  For example, unfiltered TEs can trigger
+huge numbers of false positives in automated gene finding tool
+\cite{Jiang:2013jt}, as well as inflate tool runtime.
 
 The best tools for repeat identification are RepeatMasker and nHMMER
-\cite{RepeatMaskerOpen:XkNxXSd7,Wheeler:2013gj}, but both employ a
-library-based search strategy.  They require some pre-compiled
+\cite{RepeatMaskerOpen:XkNxXSd7,Wheeler:2013gj}, but both employ 
+library-based search strategies using a pre-compiled
 description of sequences in the family (e.g. a ancestral sequence for
-BLASTing, or a profile HMM) that is used as the basis for
-identification of new member of the family.  But, much like we ask how
+BLASTing, or a profile HMM).  But, much like we ask how
 the snow plow driver gets to work \cite{Pratchett:uw}, we must ask how
-these libraries are based are initially compiled.  Library-based tools
-are useless for the discovery of new families, and cannot be applied
-to a newly sequenced genome for which such a library has not been
-compiled.
+these libraries are compiled.  Library-based tools
+are useless for the discovery of new families, and this are
+challanging to use on newly sequenced genomes.
 

### 2015-09-19 testing_pipeline.py
Updating testing_pipeline.py

@@ -113,6 +113,7 @@ def parse_params():
     raider_argument.add_argument('--pre', '--pre_scan', action = 'store_true', help = "Use pre-scan version of raider", default = False)
     raider_argument.add_argument('--mem', action = 'store_true', help = "Use large memory-nodes", default = False);
     raider_argument.add_argument("--mn", '--max_nodes', dest = "max_nodes", action="store_true", help="Reserve all nodes of a processor for each tool (disabled by default).", default=False)
+    raider_argument.add_argument("--SC", "--suppress_consensus", dest = suppress_consensus, action="store_store", help="Suppress consensus generation", default = False)
     seed_group = raider_argument.add_mutually_exclusive_group(required = False)     
     seed_group.add_argument('-s', '--seed', dest = "seed", help = "Spaced seed string", default = "111111111111111111111111111111")    
     seed_group.add_argument('--sf', '--seed_file', dest = 'seed_file', help = 'File containing raider seeds', default = None)
@@ -336,30 +337,31 @@ def raider_pipeline(raider_exe, input_file, seed, f):
     
 
     # Step 2: Generate consensus sequence
-    cmd2 = consensus_cmd.format(python=Locations['python'], data_file=input_file,
-                                elements_file=elements_dir + "/elements",
-                                consensus_txt=consensus_txt,
-                                consensus_fa=consensus_fa)
-    title2 = "con." + title
-    p2 = launch_job(cmd=cmd2, title=title2, base_dir=elements_dir, depend=[p1], attrs = {'consensus':consensus_fa})
+    if not args.suppress_consensus:
+        cmd2 = consensus_cmd.format(python=Locations['python'], data_file=input_file,
+                                    elements_file=elements_dir + "/elements",
+                                    consensus_txt=consensus_txt,
+                                    consensus_fa=consensus_fa)
+        title2 = "con." + title
+        p2 = launch_job(cmd=cmd2, title=title2, base_dir=elements_dir, depend=[p1], attrs = {'consensus':consensus_fa})
 
-    # Step 3: Apply repeat masker consensus output
-    if (args.run_rm):
-        cmd3 = repeat_masker_cmd.format(RepeatMasker = Locations['RepeatMasker'],
-                                        library_file = consensus_fa, pa = args.pa,
-                                        output_dir = rm_dir,
[…]

### 2015-09-19 manuscript/JCB/manuscript.tex

@@
 
-We can now modif the previous definition of elementary repeats as
+We can now modify the previous definition of elementary repeats as
 follows:
@@
 
+We see in the attached table the pseudocode for the phRAIDER
+algorithm, used to find a sequence descriptor for each set of
+sequences that comprise an elementary repeat family with respoect to
+some fixed seed $s$ and minimum frequency threshold $f$.  We see see
+the core components of this algorithm as:
+\begin{itemize}
+\item
+
 \begin{algorithm}
-\caption{This is the algorthm caption.}
+\renewcommand\thealgorithm{}
+\caption{phRAIDER}
 \begin{algorithmic}
@@
 \end{algorithmic}
+\addtocounter{algorithm}{-1}
 \end{algorithm}

### 2015-09-19 testing_pipeline.py
Updating locations testing_pipeline

@@ -113,7 +113,7 @@ def parse_params():
     raider_argument.add_argument('--pre', '--pre_scan', action = 'store_true', help = "Use pre-scan version of raider", default = False)
     raider_argument.add_argument('--mem', action = 'store_true', help = "Use large memory-nodes", default = False);
     raider_argument.add_argument("--mn", '--max_nodes', dest = "max_nodes", action="store_true", help="Reserve all nodes of a processor for each tool (disabled by default).", default=False)
-    raider_argument.add_argument("--SC", "--suppress_consensus", dest = suppress_consensus, action="store_store", help="Suppress consensus generation", default = False)
+    raider_argument.add_argument("--SC", "--suppress_consensus", dest = "suppress_consensus", action="store_true", help="Suppress consensus generation", default = False)
     seed_group = raider_argument.add_mutually_exclusive_group(required = False)     
     seed_group.add_argument('-s', '--seed', dest = "seed", help = "Spaced seed string", default = "111111111111111111111111111111")    
     seed_group.add_argument('--sf', '--seed_file', dest = 'seed_file', help = 'File containing raider seeds', default = None)
@@ -294,7 +294,7 @@ def setup():
 
 
         
-raider_cmd = "mkdir {TMPDIR}; /usr/bin/time {raider} -q -c {f} -s {seed} {input_file} {TMPDIR}; ls {TMPDIR}; cp {TMPDIR}/elements {output_dir}/; rm -r -f {TMPDIR}"
+raider_cmd = "mkdir {TMPDIR}; {time} {raider} -q -c {f} -s {seed} {input_file} {TMPDIR}; ls {TMPDIR}; cp {TMPDIR}/elements {output_dir}/; rm -r -f {TMPDIR}"
 consensus_cmd = "{python} consensus_seq.py -s {data_file} -e {elements_file} {consensus_txt} {consensus_fa}"
[…]

### 2015-09-19 locations.py
Updating locations.py for time

@@ -37,7 +37,8 @@ RedhawkLocations = {'build_lmer_table':'./build_lmer_table',
                     'high_mem_arch' : 'redhawk',
                     'blast' : 'blastn',
                     'blast_modules' : ['blast+'],
-                    'time_cmd' : "/usr/bin/time -f \"User: %U\\nM: %M\\nK: %K\""}
+                    'time_cmd' : '/usr/bin/time -v'}
+                    #'time_cmd' : "/usr/bin/time -f \"User: %U\\nM: %M\\nK: %K\""}
 OakleyLocations = {'build_lmer_table':'./build_lmer_table',
                    'RptScout':'./RepeatScout',
                    'filter_stage-1':'./filter-stage-1.prl',

### 2015-09-20 collect_runtimes.py
Adding collect_runtimes.py

@@ -1,0 +1,29 @@
+import glob
+import re
+
+TD = {'RS':'RepeatScout', 'phRA':'phRAIDER', 'RA':'RAIDER', 'prephRA':'pre-phRAIDER'}
+
+D = {x:y for line in open("sizes.txt") for x,y in [re.split("\s+", line.rstrip())]}
+
+for file in glob.glob("*/*/*.stderr"):
+   if "add_files" in file:
+      continue
+   r = re.search("(\w+)\.(\w+)\.(chr\w+)\.s(\d+)\.f(\d+)\.stderr$", file);
+   if not r:
+      continue
+   tool, build, chr, seed, f = r.group(1,2,3,4,5)
+   if build + "." + chr not in D:
+      continue
+
+   time = "NA"
+   mem = "NA"
+   for line in open(file):
+      r = re.search("User time \(seconds\):\s+(\d+\.\d+)", line)
+      if r:
+         time=r.group(1)
+         continue
+      r = re.search("Maximum resident set size \(kbytes\):\s+(\d+)", line)
+      if r:
+         mem = int(r.group(1))/4
+
+   print("\t".join([TD[tool], build, chr, build + "." + chr, seed, D[build + "." + chr], str(time), str(mem)]))

### 2015-09-20 collect_runtimes.py
Updating collect_runtimes.py

@@ -1,10 +1,23 @@
 import glob
 import re
 
-TD = {'RS':'RepeatScout', 'phRA':'phRAIDER', 'RA':'RAIDER', 'prephRA':'pre-phRAIDER'}
+ToolsDic = {'RS':'RepeatScout', 'phRA':'phRAIDER', 'RA':'RAIDER', 'prephRA':'pre-phRAIDER'}
+SizeDic = {x:y for line in open("sizes.txt") for x,y in [re.split("\s+", line.rstrip())]}
 
-D = {x:y for line in open("sizes.txt") for x,y in [re.split("\s+", line.rstrip())]}
+def readFile(file):
+   for line in open(file):
+      r = re.search("User time \(seconds\):\s+(\d+\.\d+)", line)
+      if r:
+         time=int(r.group(1))
+         continue
+      r = re.search("Maximum resident set size \(kbytes\):\s+(\d+)", line)
+      if r:
+         mem = int(r.group(1))/4
+         return (time,mem)
+      return ('NA', 'NA')
 
+
+R = {}
 for file in glob.glob("*/*/*.stderr"):
    if "add_files" in file:
       continue
@@ -12,18 +25,24 @@ for file in glob.glob("*/*/*.stderr"):
    if not r:
       continue
    tool, build, chr, seed, f = r.group(1,2,3,4,5)
+   key = (tool,build,chr,seed,f)
+
+   
+   label = build + "." + chr + "." = 
    if build + "." + chr not in D:
       continue
 
-   time = "NA"
-   mem = "NA"
-   for line in open(file):
-      r = re.search("User time \(seconds\):\s+(\d+\.\d+)", line)
-      if r:
-         time=r.group(1)
-         continue
-      r = re.search("Maximum resident set size \(kbytes\):\s+(\d+)", line)
-      if r:
-         mem = int(r.group(1))/4
+   
+   time,mem = readFile(file)
+   if time == 'NA' or mem == 'NA':
+      next
 
-   print("\t".join([TD[tool], build, chr, build + "." + chr, seed, D[build + "." + chr], str(time), str(mem)]))
+   if key not in R:
+      R[key] = (time,mem)
+   else:
+      t,m = R[key]
+      R[key][0] += t
+      R[key][1] = min(R[key][1], m)
+
+for key in sorted(R.key):
+   print("\t".join([tool, build + "." + chr, SizeDic[build + "." + chr]

### 2015-09-20 collect_runtimes.py
Updating collect_runtimes.py

@@ -8,13 +8,13 @@ def readFile(file):
    for line in open(file):
       r = re.search("User time \(seconds\):\s+(\d+\.\d+)", line)
       if r:
-         time=int(r.group(1))
+         time= float(r.group(1))
          continue
       r = re.search("Maximum resident set size \(kbytes\):\s+(\d+)", line)
       if r:
-         mem = int(r.group(1))/4
+         mem = float(r.group(1))/4
          return (time,mem)
-      return ('NA', 'NA')
+   return ('NA', 'NA')
 
 
 R = {}
@@ -25,24 +25,21 @@ for file in glob.glob("*/*/*.stderr"):
    if not r:
       continue
    tool, build, chr, seed, f = r.group(1,2,3,4,5)
+   if build + "." + chr not in SizeDic:
+      continue
    key = (tool,build,chr,seed,f)
 
-   
-   label = build + "." + chr + "." = 
-   if build + "." + chr not in D:
-      continue
-
-   
+      
    time,mem = readFile(file)
    if time == 'NA' or mem == 'NA':
-      next
+      continue
 
    if key not in R:
       R[key] = (time,mem)
    else:
       t,m = R[key]
-      R[key][0] += t
-      R[key][1] = min(R[key][1], m)
+      R[key] = (t+time, max(m,mem))
 
-for key in sorted(R.key):
-   print("\t".join([tool, build + "." + chr, SizeDic[build + "." + chr]
+print("\t".join(["tool", "label", "seed", "f", "size", "time", "mem"]) + "\n")
+for key in sorted(R.keys()):
+   print("\t".join([ToolsDic[key[0]], key[1] + "." + key[2], str(key[3]), str(key[4]), SizeDic[key[1] + "." + key[2]], str(R[key][0]), str(R[key][1])]))

### 2015-09-21 analysis2.R
Updating analysis / runtimes

@@ -107,3 +107,69 @@ fixed_plots <- function() {
   plot.sen.by.ratio("ce10.chrV", stats6)
   plot.sen.by.ratio("ce10.chrV", stats7)
 }
+
+############################################
+plot.runtimes <- function() {
+  RT <- read.table("runtime.txt", header=TRUE)
+  RT$size = RT$size / 1000000
+  RT <- RT[RT$size > 0,]
+  plot(c(), c(), xlim = range(RT$size), ylim = range(RT$time, na.rm = TRUE), main="Runtime",
+       xlab = "Genome size (Mb)", ylab = "Tool runtime (seconds)")
+
+  #LMER = RT[RT$tool=="LMER",]
+  #points(LMER$size, LMER$time, col = 'green')  
+  
+  RS = RT[RT$tool=="RepeatScout",]
+  points(RS$size, RS$time, col='red')
+    
+  PHRA = RT[RT$tool=="phRAIDER" & RT$seed==1,]
+  points(PHRA$size, PHRA$time, col='blue')
+  
+  RA = RT[RT$tool=="RAIDER" & RT$seed==1,]
+  points(RA$size, RA$time, col='purple')
+  
+  PRE = RT[RT$tool=="pre-phRAIDER" & RT$seed==1,]
+  points(PRE$size, PRE$time, col='black')  
+  
+}
+
+
+plot.mem <- function() {
+  RT <- read.table("runtime.txt", header=TRUE)
+  RT$size = RT$size / 1000000
+  RT$mem = RT$mem / 1000000
+  RT <- RT[RT$size > 0,]
+  plot(c(), c(), xlim = range(RT$size), ylim = range(RT$mem, na.rm = TRUE), main="Memory",
+       xlab = "Genome size (Mb)", ylab = "Tool memory (GB)")
+  
+  #LMER = RT[RT$tool=="LMER",]
+  #points(LMER$size, LMER$mem, col = 'green')  
+  
+  RS = RT[RT$tool=="RepeatScout",]
+  points(RS$size, RS$mem, col='red')
+
+  PHRA = RT[RT$tool=="phRAIDER" & RT$seed==1,]
+  points(PHRA$size, PHRA$mem, col='blue')
+  
+  RA = RT[RT$tool=="RAIDER" & RT$seed==1,]
+  points(RA$size, RA$mem, col='purple')
+  
+  PRE = RT[RT$tool=="pre-phRAIDER" & RT$seed==1,]
[…]

### 2015-09-21 collect_runtimes.py

@@ -21,7 +21,7 @@ R = {}
 for file in glob.glob("*/*/*.stderr"):
    if "add_files" in file:
       continue
-   r = re.search("(\w+)\.(\w+)\.(chr\w+)\.s(\d+)\.f(\d+)\.stderr$", file);
+   r = re.search("(\w+(?:\.\w+))\.(\w+)\.(chr\w+)\.s(\d+)\.f(\d+)\.stderr$", file);
    if not r:
       continue
    tool, build, chr, seed, f = r.group(1,2,3,4,5)

### 2015-09-27 analysis2.R
Updating analysis2.txt

@@ -182,7 +182,7 @@ fig1Plot <- function() {
   RT$size <- RT$size / 1000000
   RT$mem <- RT$mem / 1000000
   #quartz()
-  par(mfrow=c(2,2))
+  par(mfrow=c(4,2))
   
   # Make RS table.  (Need to add lmer and RS times/mem.)
   RS = subset(RT, tool == "RepeatScout")
@@ -207,6 +207,7 @@ fig1Plot <- function() {
   with(PPH1, points(size, mem, col = 'green', pch = 3))
   legend("topleft", c("RepeatScout", "phRAIDER", "pre-phRAIDER"), col = c('red', 'blue', 'green'), pch = c(1,2,3))
   
+  
   # Third: runtime for phRAIDER on different seeds
   plot(c(), c(), xlim = range(RT$size), ylim = range(c(PH0$time, PH1$time), na.rm = TRUE), xlab = "Genome size (Mb)", ylab = "Runtime (seconds)", main = "Runtime for phRAIDER on different seeds")
   colors = c('blue', 'black')
@@ -215,11 +216,25 @@ fig1Plot <- function() {
   }
   legend("topleft", c("phRAIDER seed 1", "phRAIDER seed 2"), col = c('blue', 'black'), pch = c(1,2))
   
-  plot(c(), c(), xlim = range(RT$size), ylim = range(c(PH0$mem, PH1$mem), na.rm = TRUE), xlab = "Genome size Mb)", ylab = "Runtime (seconds)", main = "Memory Usage for phRAIDER on different seeds")
+  plot(c(), c(), xlim = range(RT$size), ylim = range(c(PH0$mem, PH1$mem), na.rm = TRUE), xlab = "Genome size Mb)", ylab = "Runtime (seconds)", main = "Memory usage for phRAIDER on different seeds")
   colors = c('blue', 'black')
   for (s in c(0,1)) {
     with(subset(RT, tool=='phRAIDER' & seed==s), points(size, mem, col = colors[[s+1]], pch = s+1))
   }
   legend("topleft", c("phRAIDER seed 1", "phRAIDER seed 2"), col = c('blue', 'black'), pch = c(1,2))
+  
+  RT <- read.table("runtime2.txt", header=TRUE)
+  T1 <- subset(RT, w==l)
[…]

### 2015-09-27 manuscript/JCB/manuscript.tex
Updating manuscript

--- 2015-09-18 manuscript/JCB/manuscript.tex
@@
-\documentclass[11pt]{article}
+\documentclass[11pt,letterpaper]{amsart}
 \usepackage{longtable}
@@
 \usepackage[noend]{algpseudocode}
-\usepackage{algorithmicx}
+%\usepackage{algorithmicx}
 \usepackage{color}
@@
 
-\usepackage[paper=letterpaper,margin=1in]{geometry}
+\usepackage[paper=letterpaper,margin=1.0in]{geometry}
+
+\usepackage[english]{babel}
+\usepackage[protrusion=true,expansion=true,final,babel]{microtype}
 
@@
 
-Transposable Elements (TEs) are genomic sequences thathad the capacity
+Transposable Elements (TEs) are genomic sequences that had (or have) the capacity
 to insert copies of themselves into other genomic locations, resulting
@@
 \cite{RepeatMaskerOpen:XkNxXSd7,Wheeler:2013gj}, but both employ
-library-based search strategies using a pre-compiled
-description of sequences in the family (e.g. a ancestral sequence for
-BLASTing, or a profile HMM).  But, much like we ask how
-the snow plow driver gets to work \cite{Pratchett:uw}, we must ask how
-these libraries are compiled.  Library-based tools
-are useless for the discovery of new families, and this are
-challanging to use on newly sequenced genomes.
-
-Within mammalian species we can largely rely on homology relationships
-to port libraries across species.  This does not hold
-so well in plants: in many cases TE composition of a given plant
-organism is species-specific.  For example, a rice-based TE library
-will only identify 25\% of the TEs in the maize genome
+library-based search strategies using a pre-compiled description of
+sequences in the family (e.g. an ancestral sequence for BLASTing, or a
+profile HMM).  But, much like we ask how the snow plow driver gets to
+work \cite{Pratchett:uw}, we must ask how these libraries are
+compiled.  Library-based tools are useless for the discovery of new
+families, making them difficult to use on newly sequenced or
+unannotated genomes. Within mammalian species we can largely rely on
+homology relationships to port libraries across species.  This does
+not hold so well in plants: in many cases TE composition of a given
+plant organism is species-specific.  For example, a rice-based TE
+library will only identify 25\% of the TEs in the maize genome
 \cite{Jiang:2013jt}.
@@
 for use on whole genomes (requiring 60 hours for 18Mb rice genome in a
-2013 study), while PILER achieves a good runtime with very low
-sensitivity .  ReAS and RepeatScout \cite{Price:2005p1247} are based
+2013 study \cite{Saha:2008dm}), while PILER achieves a good runtime with very low
+sensitivity.  ReAS and RepeatScout \cite{Price:2005p1247} are based
 on $k$-mer searches, with the earlier showing less sensitivity than
@@
 DeBrujin graphs, which allows for a decomposition of TE families into
-domains, but is very, very computationally expensive.  In Saha {\it et
+domains, but is computationally expensive.  In Saha {\it et
   al.} the authors perform an extensive comparison of the tools, and
@@
 an elementary repeat if: (1) It is of at least length $l$ (the length
-requirement); (2) there are at least $f$ copies of $r$ appear (the
+requirement); (2) at least $f$ copies of $r$ appear  n the genome (the
 frequency requirement) (3) there is no proper substring of $r$ of
@@
 algorithm that could handle variation through use of
-PatternHunter-like spaced seeds that we developed the prototype RIADER
+PatternHunter-like spaced seeds that we developed the prototype RAIDER
 \cite{Li:2004wl}.  A rough implementation was first presented in
-Figueroa {\it et al.}, with more details in the Figueroa masters
-thesis \cite{Figueroa:2014uk,Figueroa:2013cz}.  RAIDER was built along
+Figueroa {\it et al.} \cite{Figueroa:2014uk}, with more details in the Figueroa masters
+thesis \cite{Figueroa:2013cz}.  RAIDER was built along
 an alternate, but equivalent definition of elementary repeats based on
 $l$-mers (sequences of length $l$).  Specifically, it was observed
-that the minimality condition could be rewritten as: There is no
+that the minimality condition could be rewritten as: {\it There is no
 $l$-mer contained within elementary repeat $r$ that appears in the
-genome more times than $r$.  From there we make four core observations
+genome more times than $r$}.  From there four core observations
 that form the basis for the RAIDER algorithm: (1) An $l$-mer cannot
@@
 we mean ``is a substring of'' -- a definition we will be generalizing
-shortly.)  For discussion and proof, see the Figueroa Thesis
+shortly.)  For discussion and proof, see the Figueroa thesis
 \cite{Figueroa:2013cz}.
@@
 then break it down later if we discover violations of the minimality
-condition.  The algorithm is summarized in Figueroa {\it et al.}, and
-discussed in detail in the Figueroa Thesis \cite{Figueroa:2014uk,Figueroa:2013cz}.
+condition.
 
@@
 (2344 seconds to 192 seconds), while coverage of the RepBase
-\cite{Jurka:2005bl} ancestral sequence improved (77\% to 84\%), while
-on mouse chromosome 19 we saw the same speedup with a significant drop
-in coverage (53\% to 30\%).  On the full human genome RAIDER ran in
-$6.3$ hours, while RepeatScout was unable to complete it run.  For
-details, see Figueroa {\it et al.} \cite{Figueroa:2014uk}.
+\cite{Jurka:2005bl} ancestral sequence improved (77\% to 84\%).  On
+mouse chromosome 19 we saw the same speedup with a significant drop in
+coverage (53\% to 30\%).  On the full human genome RAIDER ran in $6.3$
+hours, while RepeatScout was unable to complete it run.
 
@@
 demonstrated that certain seed patterns can indunce significant
-imporvements in BLAST sensitivity with out time penalty, though what
+imporvements in BLAST sensitivity without time penalty, though what
 makes a good pattern is not well understood.
@@
 
-In the following we will present a new theoretical model for {\it
-  seeded elementary repeats}, followed by an algorithm for identifying
-them.  In the results section we will show the result of using that
-algorithm for masking transp
-
-In our Methods section
-we present the new model that allows us to extend RAIDER to correctly
-use spaced seeds, and briefly outline the algorithm (with more details
-provided in the supplementary materials), with a analysis of phRAIDER
-performance in our Results.
-
 \section*{Model}
@@
 Our goal is to redefine the concept of transposbile elements to
-accomodate a spaced seed strategy.  In the next section we will describe our
-identification algorithm, and then quantified its success in 
-masking transposable elements. But we will start here with a brief
+accomodate a spaced seed strategy.   We will start here with a brief
 outline of our theoretical model (with a more detailed description in
-the appendix).
+the appendix), then present the algorithm and results.
 
@@
 $r=AA*CCGTT$, the decomposition would be $\{AA*CC, CC*TT\}$.)  We say
-$s$ {\it covers} $r$ is every base in $r$ is contained in at least on
-string in the decompositio.  (Hence the previous $s$ does not cover $AAA*CC$, as
+$s$ {\it covers} $r$ if every base in $r$ is contained in at least on
+string in the decomposition.  (Hence the previous $s$ does not cover $AAA*CC$, as
 the first base in not in any of the strings of the decomposition.)
 
-We can now modif the previous definition of elementary repeats as
+We can now modify the previous definition of elementary repeats as
 follows:
@@
   decomposition of $r$ w.r.t $s$, the number of occurences of $t$ in
-  $G$ is equal to the number of occurneces of $r$ in the genome.
+  $G$ is equal to the number of occureneces of $r$ in the genome.
 \item \underline{Maximality requirement:} There is no sequence
   descriptor $r'$ of $r$ that contains $r$ as a proper substring and
-  satisfies conditions 1-1.
+  satisfies conditions the previous thre conditions.
 \end{itemize}
@@
 
+We have written out a proof of this in the appendix.
+
 \subsection*{phRAIDER Algorithm}
 
+We see in Algorithm Pseudocode on the next page a description of the
+phRAIDER algorithm, used to find a sequence descriptor for each set of
+sequences that comprise an elementary repeat family with respect to
+some fixed seed $s$ and minimum frequency threshold $f$. The core idea
+behind the algorothm is to find sets of $l$-mers that match w.r.t. the
+seed, and whose instances are fixed at a specific spacing internval
+between them.  The $H$ dictionary maps each encountered {\it seeded}
+$l$-mer (that is, the $l$-mer after removing all characters
+corresponding to a 0 in the seed) to a list of genomic locations.
+Thus the {\bf sameOffset($v_1$, $v_2$)} can, in constant time, tell us
+whterh instances of seed $l$-mers $v_1$ and $v_2$ have the same
+distance between their offsets. The loop at line
+\algref{phRaider}{mainLoop} takes us through the genome, looking at
+each $l$-mer to identify whether it should be grouped with a
+recently-seen seen $l$-mer set.  The $Q$ structure holds the candidate
+families that $v$ might belong to (any family that we have seen in the
+last $s$ bases).  When we look at the seeded $l$-mer $v$ we need to
+decide if (1) this is the second occurance of $v$, and hence is the
+start of a family, (2) this is the next $l$-mer belonging to a family
+we have recently see (seen in the last $|s$ bases, (3) this is the
+first $l$-mer of a family not recently seen, or (4) this is an
+out-of-place $l$-mer in the midle of the family, implying that family
+needs to be broken up.
+
 \begin{algorithm}
-\caption{This is the algorthm caption.}
-\begin{algorithmic}
+\renewcommand\thealgorithm{}
+\caption{phRAIDER}
+\label{phRaider}
+\begin{algorithmic}[1]
   \Function{sameOffset}{$l$-mer $v_1$, $l$-mer $v_2$}
@@
     \comment{of the last $|L|$ $l$-mers seen, sorted by the position of those last $l$-mers.}
-    \State Queue$<$Family$>$ $Q$;
+    \State Queue$<$Family$>$ $Q$; \label{Queue}
     \State
     \comment{Iterate over all $l$-mer positions in $G$}
-    \For {$i \leftarrow 1 \to |G|-|s|$}
+    \For {$i \leftarrow 1 \to |G|-|s|$}  \label{mainLoop}
 
        \comment{Remove any families whose last $l$-mer did not occur in the last $i$ bases.}
+       \comment{I {\it think} this will never remove more than one element.}
        \While{$i - $location($Q$, finish, last, end)$ > |s|$}
@@
           \comment{$v$ either is the beginning of a family, or should be combined with a}
-          \comment{previous $l$-mer to form a family.}
+          \comment{previous $l$-mer to form a family.  (I think we can
+            set this up to work in $O(1)$ time.}
           \State $F = \mbox{arg\_find}_{F \in Q}\{\mbox{sameOffset}(F.\mbox{start}(), v)\}$

### 2015-09-28 create_pra.py
Adding create_pra.py

@@ -1,0 +1,33 @@
+import sys
+import redhawk
+import glob
+import os
+
+reg = re.compile("^(.*)\//([^\/]()\.s(\d+)\.f(\d+)\.blast.6.txt(\.bz2)?$")
+
+def main(DIR):
+    for file in glob.glob("{DIR}/*/*/*.blast.6.txt*".format(DIR)):
+        r = reg.search(file)
+        if not r:
+            sys.stderr.write("Bad file: " + file + "\n")
+            exit(1);
+        D = r.group(1)
+        org = r.group(2)
+        s = int(r.group(3))
+        f = int(r.group(4))
+
+        pra_output =  D + "/" + "{org}.s{seed}.f{f}.pra.txt".format(org=org, seed=seed_num, f=f)
+
+        if os.file.exists(pra_output):
+            sys.stdout.write("Skipping: " + pra_output + "\n")
+            continue
+
+        if file.endswith(".bz2"):
+            cmd = "bzcat {blast_output} | ./pra_analysis2 {output}".format(blast_output=file, output=pra_output)
+        else:
+            cmd = "cat {blast_output} | ./pra_analysis2 {output}".format(blast_output=file, output=pra_output)
+
+        sys.stdout.write("CMD: " + cmd)
+        
+if __name__ == "__main__":
+    main(sys.arv[1])

### 2015-09-28 collect_stats.py
updating collect_stats.py, create_pra.py

@@ -78,6 +78,7 @@ def collectRaider(DIR, tool):
     for org in data_map.keys():
         for seed_num in seed_map.keys():
             for f in f_list:
+                print("File: " + org + " " + str(seed_num) + " " + str(f) + "\n")
                 RAIDER_job_file = DIR + "/../job_log/{prefix}.{org}.s{seed_num}.f{f}".format(prefix = tool_prefix[tool], org=org, seed_num=seed_num, f=f)        
                 RM_job_file = DIR + "/../job_log/rm.{prefix}.{org}.s{seed_num}.f{f}".format(prefix = tool_prefix[tool], org=org, seed_num=seed_num, f=f)
                 RM_dir = DIR + "/" + ("{org}.s{seed}.f{f}".format(org=org, seed=seed_num, f=f)).upper()
@@ -133,8 +134,10 @@ def collectRaider(DIR, tool):
 
                 
                 if os.path.exists(blast_file):
-                    cmd = "bzcat {blast_output} | ./pra_analysis2 {output}".format(blast_output=blast_file, output=pra_output)
-                    subprocess.call(cmd, shell=True)
+                    if not os.path.exists(pra_output):
+                        cmd = "bzcat {blast_output} | ./pra_analysis2 {output}".format(blast_output=blast_file, output=pra_output)
+                        print("cmd: " + cmd)
+                        #subprocess.call(cmd, shell=True)
                     query_cover, target_cover, Used = parse_pra_output.parse_pra_output(pra_output, "exclude.txt")
                     H['ConCoverage'], H['QuCoverage'] = query_cover, target_cover
 

### 2015-09-28 testing_pipeline.py
testing_pipeline.py: bug fix. Problem with blast and rm blast file creation.

@@ -356,7 +356,7 @@ def raider_pipeline(raider_exe, input_file, seed, f):
     # Step 3: Apply repeat masker consensus output
     if args.run_rm:
         cmd3 = repeat_masker_cmd.format(RepeatMasker = Locations['RepeatMasker'],
-                                        library_file = consensus_fa, pa = args.pa,
+                                        library_file = elements_dir + "/" + consensus_fa, pa = args.pa,
                                         output_dir = rm_dir,
                                         seq_file = input_file, seq_file_base = file_base(input_file), TMPDIR = tmp_dir())
         title3 = "rm." + title
@@ -365,7 +365,7 @@ def raider_pipeline(raider_exe, input_file, seed, f):
 
     # Step 4: Apply blast to consensus sequences:
     if args.run_blast:
-        cmd4 = blast_cmd.format(blast = Locations['blast'], blast_format = blast_format, blast_dir = blast_dir, blast_file = blast_file, consensus_file = consensus_fa,
+        cmd4 = blast_cmd.format(blast = Locations['blast'], blast_format = blast_format, blast_dir = blast_dir, blast_file = blast_file, consensus_file = elements_dir + "/" + consensus_fa,
                                 db_file = database_file, evalue = args.evalue, short = "-task blastn-short" if args.short else "", max_target = args.max_target,
                                 num_threads = args.num_threads, TMPDIR = tmp_dir())
         title4 = "bl." + title

### 2015-09-28 testing_pipeline.py
Update seeds15.stats; fix testing_pipeline.py bug (add to f.txt)

@@ -247,7 +247,10 @@ def setup():
     if not os.path.exists(job_log_dir):
         os.makedirs(job_log_dir)
 
-    open(args.results_dir + "/f.txt", "w").write("\t".join([str(x) for x in args.f]) + "\n")
+    if os.path.exists(args.results_dir + "/f.txt"):
+        S = {int(f.rstrip()) for f in open(args.results_dir + "/f.txt")}
+        args.f = list(S | set(args.f))
+    open(args.results_dir + "/f.txt", "w").write("\n".join([str(x) for x in args.f]) + "\n")
 
     global progress_fp
     progress_fp = open(debug_file, "w")

### 2015-09-29 analysis3.R
Adding analysis3.R

@@ -1,0 +1,207 @@
+orgList <- c("ce10.chrV", "hg38.chr22", "ce10.chrV", "hg38.chr18.sim")
+
+read.file <- function(file) {
+  T <- read.table(file, header = TRUE)
+  return(T)
+}
+
+read.stats <- function() {
+  seeds32 <<- read.table("seeds32.stats.txt", header = TRUE)
+  RptSct <<- read.table("RptSct.stats.txt", header = TRUE)
+  
+  TList <<- lapply(orgList, function(o) subset(seeds32, org == o))
+  names(TList) <<- orgList;
+}
+
+add.legend <- function(position) {
+  legend(position, c("RM", "PRA"), col = c('blue', 'green'), 
+                     lty = c('solid', 'solid', 'solid', 'solid'))
+}
+
+plot.gen <- function(org_name, x_col, y_col, xlab = NA, ylab = NA, main = NA, 
+                     xlim = NA, ylim = NA, legend_position = "topleft", new_plot = TRUE,
+                     reverse_x = FALSE, reverse_y = FALSE) {
+
+  PH = subset(seeds32, org == org_name, select = c(x_col,y_col))
+  RS = subset(RptSct, org == org_name, select = c(x_col,y_col))
+  
+  if (is.na(xlim)) {
+    xlim = range(PH[[x_col]], rm.na = TRUE)
+  }
+  if (is.na(ylim)) {
+    ylim = range(c(PH[[y_col]], RS[[y_col]]));
+  }
+  
+  if (reverse_x) {
+    xlim = rev(xlim)
+  }
+  if (reverse_y) {
+    ylim = rev(ylim)
+  }
+  
+  if (new_plot) {
+    plot(c(), c(), xlim = xlim, ylim = ylim, xlab = xlab, ylab = ylab, main = main)
+  }
+  abline(h = RS[1,y_col], col = 'red')
+  points(PH[[x_col]], PH[[y_col]], col = 'blue')
+}
+  
+  
+
+
+plot.test <- function(org_name, x_axis = "w.l", x_label = "w/l") {
+  par(mfrow=c(2,2))
+  
+  PH = subset(seeds32, org == org_name)
+  RS = subset(RptSct, org == org_name)
+  
+  # First the Masking-metric
+  xlim = range(PH$x_axis)
[…]
